---
title: "Clean Air Countdown Campaign"
author: "Aarsh"
date: '2022-12-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# get packages, global variables, datasets
```{r load-lib-helper-files-data-global-var}

# libraries
library(readr)
library(dplyr)
library(stringr)
library(magrittr)
library(ggplot2)
library(readr)
library(tidytext)
library(tidyr)
library(tidyverse)
library(sf)
library(usethis)
library(devtools)
#library(data.table)
library(stringr)
library(forcats) # for reordering bar graphs within a plot
library(cowplot) # for inset plots

# load and document to keep things updated
devtools::load_all()
devtools::document()

# read in latest color file
color_2020 <- read_csv("./june.2022/master.dataset/color_2020.csv")

# read in old available color datasets
color_2019 <- read_csv("./june.2022/master.dataset/color_2019.csv")
color_2016 <- read_csv("./june.2022/master.dataset/color_2016.csv")

# read in color map shape file
colormap_shp <- st_read("./june.2022/other.important.calculations.data/color.map.shape.file/colormap.shp")

# GBD final results file (resulting file that we get after applying the steps in the GBD technical appendix)
gbd_results <- read_csv("./june.2022/other.important.calculations.data/estimated_life_expectancy_differences_master_table_final.csv")

# Countries with Ambient Air Quality Standards
aaqs_places <- read_csv("./june.2022/other.important.calculations.data/countries_list_with_aaqs.csv")
 
# read in "county for aqli pm2.5 dataset" (US technical appendix output file)
county_for_aqli_pm2.5_dataset <- read_csv("./june.2022/other.important.calculations.data/county_pm25_foraqli_stats.csv")


# investments in NCAP data (not using)
# ncap_budget_tracker <- read_csv("./june.2022/clean.air.countdown.campaign.2023/input/ncap_budget_tracker_data.csv")

# NCAP city, districts, state list 
ncap_regions_list <- read_csv("./june.2022/clean.air.countdown.india.campaign.2023/input.data.shared.between.graphs/ncap_cities_district_state_long.csv")

# AQLI district level data classified as Rural/Urban using CSE's classification
aqli_2020_district_level_rural_urban_subset <- read_csv("./june.2022/clean.air.countdown.india.campaign.2023/input.data.shared.between.graphs/aqli_2020_district_level_rural_urban_subset.csv")

# india state
india_state <- st_read("./june.2022/clean.air.countdown.india.campaign.2023/input.data.shared.between.graphs/india_state.shp")

# color 2020 data for those Indian states, through which the TOC passes, this contains an identifier column that tells us, whether a given district is south/north of the toc
color_2020_india_states_toc_passes_with_toc_id <- read_csv("./june.2022/clean.air.countdown.india.campaign.2023/input.data.shared.between.graphs/color_2020_states_tropic_of_cancer_passes.csv")

color_2020_india_states_toc_passes_with_toc_id <- color_2020_india_states_toc_passes_with_toc_id %>%
  select(-dir_rel_to_toc, dir_rel_to_toc)



# global variables
who_guideline <- 5
le_constant <- 0.098
ncap_midpoint <- 25
nat_stan_india <- 40

# global operations
`%notin%` <- Negate(`%in%`)


```


# Figure 1: Five Indian states with greatest increase in pollution over the past five years

```{r fig1}

# five states with greatest increase in pollution over the past five years
five_states_with_greatest_increase <- color_2020 %>%
  filter(country == "India") %>%
  group_by(name_1) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2020_pop_weighted = pm2020*pop_weights, 
         pm2016_pop_weighted = pm2016*pop_weights) %>%
  summarise(avg_pm2.5_2016 = sum(pm2016_pop_weighted, na.rm = TRUE), 
            avg_pm2.5_2020 = sum(pm2020_pop_weighted, na.rm = TRUE)) %>%
  mutate(avg_pm2.5_2020_vs_2016 = avg_pm2.5_2020 - avg_pm2.5_2016, 
         avg_pm2.5_2020_vs_2016_prop_change = round((avg_pm2.5_2020_vs_2016/avg_pm2.5_2016)*100, 2)) %>%
    filter(name_1 %notin% c("Daman and Diu", "Dadra and Nagar Haveli")) %>%
  slice_max(avg_pm2.5_2020_vs_2016, n = 5) %>%
  select(name_1) %>%
  unlist() %>%
  as.vector()


# figure 1 data
figure_1_data <-  color_2020 %>%
  filter(country == "India") %>%
  group_by(name_1) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2020_pop_weighted = pm2020*pop_weights, 
         pm2019_pop_weighted = pm2019*pop_weights,
         pm2018_pop_weighted = pm2018*pop_weights,
         pm2017_pop_weighted = pm2017*pop_weights,
         pm2016_pop_weighted = pm2016*pop_weights) %>%
  summarise(avg_pm2.5_2016 = sum(pm2016_pop_weighted, na.rm = TRUE), 
            avg_pm2.5_2017 = sum(pm2017_pop_weighted, na.rm = TRUE), 
            avg_pm2.5_2018 = sum(pm2018_pop_weighted, na.rm = TRUE), 
            avg_pm2.5_2019 = sum(pm2019_pop_weighted, na.rm = TRUE),
            avg_pm2.5_2020 = sum(pm2020_pop_weighted, na.rm = TRUE)) %>%
  mutate(avg_pm2.5_2020_vs_2016 = avg_pm2.5_2020 - avg_pm2.5_2016, 
         avg_pm2.5_2020_vs_2016_prop_change = round((avg_pm2.5_2020_vs_2016/avg_pm2.5_2016)*100, 2)) %>%
  filter(name_1 %in% five_states_with_greatest_increase) %>%
  select(-avg_pm2.5_2020_vs_2016, -avg_pm2.5_2020_vs_2016_prop_change) %>%
  pivot_longer(avg_pm2.5_2016:avg_pm2.5_2020, names_to = "year", values_to = "pm2.5_pollution")%>% mutate(year = str_extract(str_extract(year, "_\\d+"), "\\d+")) %>%
  mutate(order_var = ifelse(name_1 == "Gujarat", "a", NA), 
         order_var = ifelse(name_1 == "Sikkim", "b", order_var), 
         order_var = ifelse(name_1 == "Madhya Pradesh", "c", order_var), 
         order_var = ifelse(name_1 == "Rajasthan", "d", order_var), 
         order_var = ifelse(name_1 == "Chhattisgarh", "e", order_var)) 

# coercing the "year" column to numeric
figure_1_data$year <- as.numeric(figure_1_data$year)
figure_1_data$name_1 <- factor(figure_1_data$name_1, levels = c("Gujarat","Sikkim", "Madhya Pradesh", 
         "Rajasthan", "Chhattisgarh"))

# creating a data frame that will contain different text for different facets 
figure_1_facet_text <- tibble(label = c("+21.1%", "+11.0%", "+11.4%", "+9.5%", "+10.4%"), 
                              label_2 = c(expression(paste("WHO ", PM[2.5], " ", "Guideline (last updated: 2021): 5 ", mu, "g","/", m^3, "")), "", "", "", ""),
                              name_1 = factor(c("Gujarat", "Sikkim", "Madhya Pradesh", 
                                         "Rajasthan", "Chhattisgarh"), levels = c("Gujarat","Sikkim", "Madhya Pradesh", 
         "Rajasthan", "Chhattisgarh")),
                              x = c(2017.7, 2018, 2018, 2018, 2018), 
                              y = c(36, 41.7, 51, 54, 59.7))

# figure 1 plot
figure_1_plot <- figure_1_data %>%
 #filter(year %in% c(2016, 2020)) %>%
  ggplot(mapping = aes(x = year, y = pm2.5_pollution, color = name_1, group = name_1)) +
  geom_point(size = 2.3) +
  geom_line(lwd = 1.5) +
  facet_grid(. ~ name_1) +
  ggplot2::geom_text(mapping = aes(x = x, y = y, label = label), data = figure_1_facet_text, fontface = "bold") +
  labs(color = "State", title = expression("Five Indian states with greatest increase in" ~ PM[2.5] ~ pollution ~ "from 2016 to 2020")) +
 # viridis::scale_color_viridis(discrete = TRUE) +
   scale_color_manual(values = c("Gujarat" = "wheat2", "Sikkim" = "gold1", "Madhya Pradesh" = "goldenrod3", "Rajasthan" = "red" , "Chhattisgarh" = "darkred")) +
  ggthemes::theme_stata() +
  labs(x = "Years", y = expression(Annual ~ average ~ PM[2.5] ~ concentration ~ "("*"in"*~mu*g*"/"*m^3*")")) +
  theme(axis.line.x = element_line(), 
        axis.line.y = element_line(), 
        strip.text.x = element_text(size = 12), 
        legend.position = "bottom", 
        plot.title = element_text(size = 15), 
        # legend.box.background = element_rect(linetype = 1, size = 0.5), 
        axis.title.y = element_text(size = 12), 
        axis.title.x = element_text(size = 12),
        legend.text = element_text(size = 12), 
        plot.caption = element_text(hjust = 0, size = 7), 
        plot.background = element_rect(fill = "white"), 
        strip.background = element_rect(fill = "darkseagreen3"), 
        axis.text.x = element_text(size = 8, margin = margin(r = 0)), 
        axis.text.y = element_text(size = 8), 
        legend.title = element_text(size = 15), 
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank()) +
  scale_y_continuous(breaks = seq(0, 70, 5)) +
    geom_hline(mapping = aes(yintercept = who_guideline), color = "black", linetype = 2) +
  scale_x_continuous(breaks = seq(2016, 2020, 1), expand = c(0.3, 0)) +
  geom_text(x = 2018.05, y = 6.5, mapping = aes(label = label_2), data = figure_1_facet_text, color = "black", size = 2.7, parse = T, fontface = "bold")

ggsave(filename = "./june.2022/clean.air.countdown.india.campaign.2023/chart.1.data.graph.code/chart1[final].png", plot = figure_1_plot, height = 7, width = 14)

figure_1_data %>%
  write_csv("./june.2022/clean.air.countdown.india.campaign.2023/chart.1.data.graph.code/chart1.data[final].csv")

svglite::svglite("final_fig1_cleanAirIndiaCountdown2023.svg")
ggsave("./final_fig1_cleanAirIndiaCountdown2023.svg", width = 14, height = 7)
```

# Figure 2: Five Indian states with greatest decrease in pollution over the past five years

```{r}

# five states with greatest decrease in pollution over the past five years
five_states_with_greatest_decrease <- color_2020 %>%
  filter(country == "India") %>%
  group_by(name_1) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2020_pop_weighted = pm2020*pop_weights, 
         pm2016_pop_weighted = pm2016*pop_weights) %>%
  summarise(avg_pm2.5_2016 = sum(pm2016_pop_weighted, na.rm = TRUE), 
            avg_pm2.5_2020 = sum(pm2020_pop_weighted, na.rm = TRUE)) %>%
  mutate(avg_pm2.5_2020_vs_2016 = avg_pm2.5_2020 - avg_pm2.5_2016, 
         avg_pm2.5_2020_vs_2016_prop_change = round((avg_pm2.5_2020_vs_2016/avg_pm2.5_2016)*100, 2)) %>%
  slice_min(avg_pm2.5_2020_vs_2016, n = 5) %>%
  select(name_1) %>%
  unlist() %>%
  as.vector()


# figure 2 data
figure_2_data <-  color_2020 %>%
  filter(country == "India") %>%
  group_by(name_1) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2020_pop_weighted = pm2020*pop_weights,
         pm2019_pop_weighted = pm2019*pop_weights,
         pm2018_pop_weighted = pm2018*pop_weights,
         pm2017_pop_weighted = pm2017*pop_weights,
         pm2016_pop_weighted = pm2016*pop_weights) %>%
  summarise(avg_pm2.5_2016 = sum(pm2016_pop_weighted, na.rm = TRUE), 
            avg_pm2.5_2017 = sum(pm2017_pop_weighted, na.rm = TRUE),
            avg_pm2.5_2018 = sum(pm2018_pop_weighted, na.rm = TRUE),
            avg_pm2.5_2019 = sum(pm2019_pop_weighted, na.rm = TRUE),
            avg_pm2.5_2020 = sum(pm2020_pop_weighted, na.rm = TRUE)) %>%
  mutate(avg_pm2.5_2020_vs_2016 = avg_pm2.5_2020 - avg_pm2.5_2016, 
         avg_pm2.5_2020_vs_2016_prop_change = round((avg_pm2.5_2020_vs_2016/avg_pm2.5_2016)*100, 2)) %>%
  filter(name_1 %in% five_states_with_greatest_decrease) %>%
  select(-avg_pm2.5_2020_vs_2016, -avg_pm2.5_2020_vs_2016_prop_change) %>%
  pivot_longer(avg_pm2.5_2016:avg_pm2.5_2020, names_to = "year", values_to = "pm2.5_pollution")%>% mutate(year = str_extract(str_extract(year, "_\\d+"), "\\d+")) %>%
  mutate(order_var = ifelse(name_1 == "Himachal Pradesh", "a", NA), 
         order_var = ifelse(name_1 == "Jammu and Kashmir", "b", order_var), 
         order_var = ifelse(name_1 == "Uttarakhand", "c", order_var), 
         order_var = ifelse(name_1 == "Punjab", "d", order_var), 
         order_var = ifelse(name_1 == "NCT of Delhi", "e", order_var)) 

# coercing the "year" column to numeric
figure_2_data$year <- as.numeric(figure_2_data$year)
figure_2_data$name_1 <- factor(figure_2_data$name_1, levels = c("Himachal Pradesh","Jammu and Kashmir", "Uttarakhand", 
         "Punjab", "NCT of Delhi"))

# creating a data frame that will contain different text for different facets 
figure_2_facet_text <- tibble(label = c("-15.0%", "-25.7%", "-13.9%", "-9.0%", "-5.3%"), 
                              label_2 = c(expression(paste("WHO ", PM[2.5], " ", "Guideline (last updated: 2021): 5 ", mu, "g","/", m^3, "")), "", "", "", ""),
                              name_1 = factor(c("Himachal Pradesh","Jammu and Kashmir", "Uttarakhand", 
         "Punjab", "NCT of Delhi"), levels = c("Himachal Pradesh","Jammu and Kashmir", "Uttarakhand", 
         "Punjab", "NCT of Delhi")),
                              x = c(rep(2018, 5)), 
                              y = c(34.7, 37.3, 44.7, 71.2, 116.4))


# figure 2 plot
figure_2_plot <- figure_2_data %>%
 # filter(year %in% c(2016, 2020)) %>%
  ggplot(mapping = aes(x = year, y = pm2.5_pollution, color = name_1, group = name_1)) +
  geom_point(size = 2.3) +
  geom_line(lwd = 1.5) +
  facet_grid(. ~ name_1) +
   ggplot2::geom_text(mapping = aes(x = x, y = y, label = label), data = figure_2_facet_text, fontface = "bold") +
  labs(color = "State", title = expression("Five Indian states with greatest decrease in" ~ PM[2.5] ~ pollution ~ "from 2016 to 2020")) +
 # viridis::scale_color_viridis(discrete = TRUE) +
   scale_color_manual(values = c("Himachal Pradesh" = "wheat2", "Jammu and Kashmir" = "gold1", "Uttarakhand" = "goldenrod3", "Punjab" = "red" , "NCT of Delhi" = "darkred")) +
  ggthemes::theme_stata() +
  labs(x = "Years", y = expression(Annual ~ average ~ PM[2.5] ~ concentration ~ "("*"in"*~mu*g*"/"*m^3*")")) +
  theme(axis.line.x = element_line(), 
        axis.line.y = element_line(), 
        strip.text.x = element_text(size = 12), 
        legend.position = "bottom", 
        plot.title = element_text(size = 15), 
        # legend.box.background = element_rect(linetype = 1, size = 0.5), 
        axis.title.y = element_text(size = 12), 
        axis.title.x = element_text(size = 12),
        legend.text = element_text(size = 12), 
        plot.caption = element_text(hjust = 0, size = 7), 
        plot.background = element_rect(fill = "white"), 
        strip.background = element_rect(fill = "darkseagreen3"),
        axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size = 8), 
        legend.title = element_text(size = 15), 
        axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank()) +
  scale_y_continuous(breaks = seq(0, 120, 10)) +
    geom_hline(mapping = aes(yintercept = who_guideline), color = "black", linetype = 2) +
scale_x_continuous(breaks = seq(2016, 2020, 1), expand = c(0.3, 0)) +
  geom_text(x = 2018.05, y = 8, mapping = aes(label = label_2), data = figure_2_facet_text, color = "black", size = 2.7, parse = T)

ggsave(filename = "./june.2022/clean.air.countdown.india.campaign.2023/chart.2.data.graph.code/chart2[final].png", plot = figure_2_plot, height = 7, width = 14)

figure_2_data %>%
  write_csv("./june.2022/clean.air.countdown.india.campaign.2023/chart.2.data.graph.code/chart2.data[final].csv")

svglite::svglite("final_fig2_cleanAirIndiaCountdown2023.svg")
ggsave("./final_fig2_cleanAirIndiaCountdown2023.svg", width = 14, height = 7)
  

```


# Figure 4: Pollution in the top 5 most populated districts not touched by NCAP

# Note:
```{r}

# ncap regions dataset
ncap_regions_list <- ncap_regions_list %>%
  dplyr::select(state_ut, city, district) %>%
  dplyr::filter(!is.na(district))

# ncap state and district list
ncap_state_list <- unique(ncap_regions_list$state_ut)
ncap_district_list <- unique(ncap_regions_list$district)

# pollution levels in the top 5 most populated districts not touched by NCAP
non_ncap_district_top5_most_populated_data <- color_2020 %>%
  filter(country == "India") %>%
  filter(name_2 %notin% ncap_district_list) %>%
  slice_max(population, n = 5) 

# top 5 most populated districts not touched by NCAP list 
non_ncap_district_top5_most_populated_list <- non_ncap_district_top5_most_populated_data %>% .$name_2

# states in which the top 5 most populated districts lie
states_of_interest <- non_ncap_district_top5_most_populated_data %>% .$name_1 %>% unique()

# states of interest shape file
states_of_focus_shp_file <- india_state %>%
  filter(NAME_1 %in% states_of_interest)

# districts of interest (within the above states of interest) shape file
districts_of_interst_shp_file <- colormap_shp %>%
  filter(NAME_0 == "India", NAME_1 %in% states_of_interest, NAME_2 %in% non_ncap_district_top5_most_populated_list)

# states of interest bounding box
states_of_interest_bb <- st_as_sfc(st_bbox(states_of_focus_shp_file))

# a state level India map with a bounding box on states of interest
figure_3_plot_part_1 <- india_state %>%
  ggplot() +
  geom_sf(fill = "snow2") +
  geom_sf(data = states_of_focus_shp_file, fill = "white", size = 0.6) +
  geom_sf(data = districts_of_interst_shp_file, fill = "red", size = 0.6) +
  geom_sf(data = states_of_interest_bb, fill = NA, color = "red", size = 0.9) +
  theme_map() 

# save shape file
 india_state %>%
  st_write("./june.2022/clean.air.countdown.india.campaign.2023/chart.3.data.graph.code/plotdata_part_1_india_state_wise_shape_file.shp")

svglite::svglite("final_fig3_part1_cleanAirIndiaCountdown2023.svg")
ggsave("./final_fig3_part1_cleanAirIndiaCountdown2023.svg", width = 7, height = 7)

figure_3_plot_part_2 <- states_of_focus_shp_file %>%
  ggplot() + 
    geom_sf(fill = "white", size = 0.8) +
   geom_sf(data = districts_of_interst_shp_file, fill = "red", size = 1) +
  theme_map() +
  labs(title = "", 
        caption = expression("*NCAP is India's National Clean Air Programme. Pollution values specifically refer to annual average" ~ PM[2.5] ~ pollution ~ "and are reported in µg/m³. Life years lost are relative to WHO" ~ PM[2.5] ~ pollution ~ "guideline"), subtitle = "2016 to 2020") +
  theme_map() +
  theme(plot.background = element_rect(fill = "azure2"), 
        plot.caption = element_text(hjust = 0, size = 9), plot.caption.position = "plot") +
   ggsflabel::geom_sf_label_repel(data = districts_of_interst_shp_file, mapping = aes(label = NAME_2), force = 100, nudge_x = -2, seed = 10, size = 7) 
  
 
states_of_focus_shp_file %>%
    st_write("./june.2022/clean.air.countdown.india.campaign.2023/chart.3.data.graph.code/plotdata_part_2_states_of_focus_shape_file.shp")

svglite::svglite("final_fig3_part2_cleanAirIndiaCountdown2023.svg")
ggsave("./final_fig3_part2_cleanAirIndiaCountdown2023.svg", width = 7, height = 7)
 
# combined inset map
figure_3_plot <- cowplot::ggdraw() +
  draw_plot(figure_3_plot_part_2, x = 0.1, y = 0.0001) +
  draw_plot(figure_3_plot_part_1, x = 0, y = 0.1, width = 0.4, height = 0.4)

# get the total population for each district that will go inside the text boxes, alongside pollution and life years lost data

non_ncap_district_top5_most_populated_data %>%
  mutate(pop_lakhs = round(population/100000, 1)) %>%
  select(country:name_2, population, pop_lakhs, everything()) %>%
    write_csv("./june.2022/clean.air.countdown.india.campaign.2023/chart.3.data.graph.code/plotdata_part_3_india_state_wise_population.csv")


  

```




# Figure 5: Calculating 2 average pollution and (corresponding life years lost) number for all regions north and south of the tropic of cancer respectively.

```{r}

# # # states north of the tropic of cancer
# states_north_of_toc <- c("Jammu and Kashmir", "Punjab", "Chandigarh", "Himachal Pradesh",
#                          "Uttarakhand", "Haryana", "NCT of Delhi", "Rajasthan",
#                          "Uttar Pradesh", "Bihar", "Sikkim", "Assam",
#                          "Arunachal Pradesh", "Nagaland", "Manipur",
#                          "Mizoram", "Tripura", "West Bengal",
#                          "Madhya Pradesh", "Jharkhand")

# states through which tropic of cancer passes
states_toc_passes <- c("Gujarat", "Rajasthan", "Madhya Pradesh", 
                       "Chhattisgarh", "Jharkhand", "West Bengal", 
                       "Tripura", "Mizoram")

# data for states through which tropic of cancer doesn't pass
color_2020_india_states_toc_not_passes <- color_2020 %>%
  filter(country == "India") %>%
  filter(name_1 %notin% states_toc_passes)

#  list of states through which tropic of cancer doesn't pass 
color_2020_india_states_toc_not_passes_list <- color_2020_india_states_toc_not_passes %>% 
  .$name_1 %>%
  unique()

# Of the states that do not lie on the tropic of the cancer, which ones are clear north of toc: list
color_2020_india_clear_north_of_toc <- c("Jammu and Kashmir", "Ladakh", 
                                            "Himachal Pradesh", "Punjab", 
                                            "Chandigarh", "Uttarakhand", "NCT of Delhi", 
                                            "Haryana", "Uttar Pradesh", "Bihar", "Sikkim", 
                                            "Arunachal Pradesh", "Assam", "Nagaland", 
                                            "Manipur", "Meghalaya")

# Of the states that do not lie on the tropic of the cancer, which ones are clear south of toc: list
color_2020_india_clear_south_of_toc <- color_2020 %>% 
  filter(country == "India") %>%
  filter(name_1 %notin% c(states_toc_passes, color_2020_india_clear_north_of_toc)) %>%
  .$name_1 %>% 
  unique()

# color 2020 data with toc identifier, for those states that lie to the clear north or south of the toc
color_2020_clear_ns_states_with_toc_id <- color_2020 %>%
  filter(country == "India", name_1 %notin% states_toc_passes) %>%
  mutate(dir_rel_to_toc = ifelse(name_1 %in% color_2020_india_clear_north_of_toc, "NorthTOC", "SouthTOC"))

# r binding state toc passes and state toc not passes datasets

final_toc_data <- rbind(color_2020_clear_ns_states_with_toc_id, color_2020_india_states_toc_passes_with_toc_id)

# final dataset for the graph (part-1)
final_toc_data %>%
  group_by(dir_rel_to_toc) %>%
  mutate(pop_weights = population/sum(population, na.rm =TRUE), 
         pm2020_pop_weighted = pm2020*pop_weights) %>%
  summarise(avg_pm2.5_2020_toc = round(sum(pm2020_pop_weighted, na.rm = TRUE), 1), 
            lyl_rel_to_who_2020 = round((avg_pm2.5_2020_toc - 5)*0.098, 1), 
            lyl_rel_to_who_2020 = ifelse(lyl_rel_to_who_2020 < 0, 0, lyl_rel_to_who_2020), 
            total_pop = sum(population, na.rm = TRUE), 
            total_pop_crores = round(total_pop/10000000, 1))

final_toc_data %>%
  write_csv("./june.2022/clean.air.countdown.india.campaign.2023/chart.4.data.graph.code/part1_chart4_data.csv")



# Plotting each state's labelled shapefile  
# state_n <- "Madhya Pradesh"
# 
# foo <- colormap_shp %>%
#   filter(NAME_0 == "India", NAME_1 == state_n) %>%
#   ggplot() +
#   geom_sf() + 
#   ggsflabel::geom_sf_label_repel(data = colormap_shp %>% filter(NAME_0 == "India", NAME_1 %in% c(state_n)), mapping = aes(label = NAME_2), force = 100, nudge_x = -2, seed = 10, size = 1.3) 

# 

# final data part 2
final_data_part_2 <- color_2020 %>%
  filter(country %in% c("India")) %>%
  left_join(colormap_shp, by = c("name_1" = "NAME_1",  "name_2" = "NAME_2")) %>%
  select(country, name_1, name_2, population, pm2020, llpp_who5_2020, geometry) %>%
  mutate(lyl_aqli_bucket = ifelse((llpp_who5_2020 >= 0) & (llpp_who5_2020 < 0.1), "0 - < 0.1 years", NA), 
         lyl_aqli_bucket = ifelse((llpp_who5_2020 >= 0.1) & (llpp_who5_2020 < 0.5), "0.1 - < 0.5", lyl_aqli_bucket), 
         lyl_aqli_bucket = ifelse((llpp_who5_2020 >= 0.5) & (llpp_who5_2020 < 1), "0.5 - < 1", lyl_aqli_bucket), 
         lyl_aqli_bucket = ifelse((llpp_who5_2020 >= 1) & (llpp_who5_2020 < 2), "1 - < 2", lyl_aqli_bucket), 
         lyl_aqli_bucket = ifelse((llpp_who5_2020 >= 2) & (llpp_who5_2020 < 3), "2 - < 3", lyl_aqli_bucket), 
         lyl_aqli_bucket = ifelse((llpp_who5_2020 >= 3) & (llpp_who5_2020 < 4), "3 - < 4", lyl_aqli_bucket), 
         lyl_aqli_bucket = ifelse((llpp_who5_2020 >= 4) & (llpp_who5_2020 < 5), "4 - < 5", lyl_aqli_bucket), 
         lyl_aqli_bucket = ifelse((llpp_who5_2020 >= 5) & (llpp_who5_2020 < 6), "5 - < 6", lyl_aqli_bucket), 
         lyl_aqli_bucket = ifelse((llpp_who5_2020 >= 6), ">= 6 years", lyl_aqli_bucket)) %>%
  mutate(order_var = ifelse(lyl_aqli_bucket == "0 - < 0.1 years", 1, NA), 
         order_var = ifelse(lyl_aqli_bucket == "0.1 - < 0.5", 2, order_var), 
         order_var = ifelse(lyl_aqli_bucket == "0.5 - < 1", 3, order_var), 
         order_var = ifelse(lyl_aqli_bucket == "1 - < 2", 4, order_var), 
         order_var = ifelse(lyl_aqli_bucket == "2 - < 3", 5, order_var), 
         order_var = ifelse(lyl_aqli_bucket == "3 - < 4", 6, order_var), 
         order_var = ifelse(lyl_aqli_bucket == "4 - < 5", 7, order_var), 
         order_var = ifelse(lyl_aqli_bucket == "5 - < 6", 8, order_var), 
         order_var = ifelse(lyl_aqli_bucket == ">= 6 years", 9, order_var))
  
final_data_part_2 %>%
   write_sf("./june.2022/clean.air.countdown.india.campaign.2023/chart.4.data.graph.code/part2_chart4_data.shp")

# plot final graph
 final_data_part_2_plot <- final_data_part_2 %>%
   select(-geometry, geometry) %>%
   st_as_sf() %>%
  ggplot() +
  geom_sf(mapping = aes(fill = lyl_aqli_bucket), color = "transparent", size = 0.01) +
   scale_fill_manual(values = c("0 - < 0.1 years" = "#FFFFFF", "0.1 - < 0.5" = "#FFE6B3", "0.5 - < 1" = "#FFD25D", 
                                "1 - < 2" = "#FFBA00", "2 - < 3" = "#FF9600", "3 - < 4" = "#FF6908", 
                                "4 - < 5" = "#E63D23", "5 - < 6" = "#BD251C", ">= 6 years" = "#8C130E")) +
   ggthemes::theme_map() +
  geom_hline(mapping = aes(yintercept = 23.5), lwd = 1.1) +
  geom_sf(data = india_state, fill = "transparent", color = "black", size = 0.5) +
  theme(legend.position = "bottom", 
        legend.justification = c(0.5, 3), 
        legend.background = element_rect(color = "black"), 
        legend.text = element_text(size = 12), 
        legend.title = element_text(size = 14)) +
   labs(fill = "Life years lost") +
  guides(fill = guide_legend(nrow = 1)) 

svglite::svglite("test_india_lyl_toc.svg")

ggsave("./test_india_lyl_toc.svg", height = 12, width = 12)
```





# [Not included in campaign, will probably use later, commented for now]: Figure 5: generate a life years lost graph grouped by population, for India as a whole, Rural India and Urban India 
```{r}

# # global aqli file grouped by pop_buckets
#   aqli_2020_district_level_pop_buckets <- aqli_2020_district_level_rural_urban_subset %>%
#   mutate(pop_bucket = ifelse(lyl_rel_to_who_2020 < 1, "< 1 year", lyl_rel_to_who_2020), 
#          pop_bucket = ifelse(((pop_bucket >= 1) & (pop_bucket < 3)), "1 - < 3", pop_bucket), 
#          pop_bucket = ifelse(((pop_bucket >= 3) & (pop_bucket < 5)), "3 - < 5", pop_bucket), 
#          pop_bucket = ifelse(((pop_bucket >= 5) & (pop_bucket <= 7)), "5 - 7", pop_bucket), 
#          pop_bucket = ifelse(pop_bucket > 7, "> 7 years", pop_bucket)) %>%
#   mutate(order_pop_buckets = ifelse(pop_bucket == "< 1 year", 1, 0), 
#          order_pop_buckets = ifelse(pop_bucket == "1 - < 3", 2, order_pop_buckets), 
#          order_pop_buckets = ifelse(pop_bucket == "3 - < 5", 3, order_pop_buckets), 
#          order_pop_buckets = ifelse(pop_bucket == "5 - 7", 4, order_pop_buckets), 
#          order_pop_buckets = ifelse(pop_bucket == "> 7 years", 5, order_pop_buckets)) 
# 
# # overall summary
# aqli_2020_district_level_pop_buckets_summary <- aqli_2020_district_level_pop_buckets %>%
#   group_by(pop_bucket) %>%
#   summarise(tot_pop = sum(population, na.rm = TRUE), 
#             order_var = order_pop_buckets[1]) %>%
#   arrange(order_var) %>%
#   ungroup() %>%
#      mutate(tot_fin_pop = sum(tot_pop, na.rm = TRUE), 
#             prop_pop = round((tot_pop/tot_fin_pop)*100, 2)) %>%
#     ungroup() %>%
#     select(pop_bucket, prop_pop, order_var) %>%
#   mutate(order_var_2 = 1)
# 
# # summary for the rural region
# aqli_2020_district_level_pop_buckets_summary_r <- aqli_2020_district_level_pop_buckets %>%
#   filter(region_type == "Predominantly Rural") %>%
#   group_by(pop_bucket) %>%
#   summarise(tot_pop = sum(population, na.rm = TRUE), 
#             order_var = order_pop_buckets[1]) %>%
#   arrange(order_var) %>%
#   ungroup() %>%
#      mutate(tot_fin_pop = sum(tot_pop, na.rm = TRUE), 
#             prop_pop = round((tot_pop/tot_fin_pop)*100, 2)) %>%
#     ungroup() %>%
#     select(pop_bucket, prop_pop, order_var) %>%
#   mutate(order_var_2 = 1)
# 
# # summary for the urban data
# aqli_2020_district_level_pop_buckets_summary_u <- aqli_2020_district_level_pop_buckets %>%
#   filter(region_type == "Predominantly Urban") %>%
#   group_by(pop_bucket) %>%
#   summarise(tot_pop = sum(population, na.rm = TRUE), 
#             order_var = order_pop_buckets[1]) %>%
#   arrange(order_var) %>%
#   ungroup() %>%
#      mutate(tot_fin_pop = sum(tot_pop, na.rm = TRUE), 
#             prop_pop = round((tot_pop/tot_fin_pop)*100, 2)) %>%
#     ungroup() %>%
#     select(pop_bucket, prop_pop, order_var) %>%
#   mutate(order_var_2 = 1)
# 
# 
# # v1 of the graph (not stacked)
# aqli_2020_district_level_pop_buckets_summary %>%
#   ggplot() +
#   geom_col(mapping = aes(x = forcats::fct_reorder(pop_bucket, order_var), y = prop_pop, fill = pop_bucket)) +
#      scale_fill_manual(values = c("> 7 years" = "black", "5 - 7" = "darkred", "3 - < 5" = "red", "1 - < 3" = "orange", "< 1 year" = "burlywood1")) +
#   labs(x = "Life Years Lost", y = "% population", title = "Proportion population in different life years lost buckets", fill = "Life Years Lost") +
#   theme(axis.title = element_text(size = 11) ,
#         axis.text.x = element_text(size = 10), 
#         axis.text.y = element_text(size = 10)) +
#   coord_flip() +
#   ggthemes::theme_hc() + 
#   theme(plot.title = element_text(size = 14)) 
# 
# # v2.1 of the graph (stacked)  
# foo <- aqli_2020_district_level_pop_buckets_summary %>% 
#   ggplot() + 
#   geom_bar(mapping = aes(x = order_var_2, y = prop_pop, fill = forcats::fct_reorder(pop_bucket, order_var, .desc = TRUE)), position = "stack", stat = "identity", height = 1, color = "black", size = 1) +
#    scale_fill_manual(values = c("< 1 year" = "burlywood1", "1 - < 3" = "orange", "3 - < 5" = "red", "5 - 7" = "darkred", "> 7 years" = "black")) +
#   labs(x = "", y = "", title = "Proportion of Indian population in different pollution life years lost buckets*", fill = "Life Years Lost", caption = "*Rural/Urban classification of India was performed using the Center for Science and Environment's classification criteria.") +
#   coord_flip() +
#   hrbrthemes::theme_ipsum() +
#   theme(plot.title = element_text(size = 15, hjust = 0.5), 
#         axis.text.y = element_blank(), 
#         axis.text.x = element_blank(),
#         axis.ticks.x = element_blank(), 
#         axis.title = element_text(size = 11), 
#         legend.position = "bottom", 
#         legend.background = element_rect(), 
#         axis.line.x = element_blank(), 
#         axis.line.y = element_blank(), 
#         axis.ticks = element_blank(), 
#         plot.background = element_rect(fill = "white"), 
#         panel.grid.major = element_blank(), 
#         panel.grid.minor = element_blank(), 
#         plot.caption = element_text(hjust = 0)) 
# 
# # v2.2 (stacked, rural, urban division)
# foo2 <- aqli_2020_district_level_pop_buckets_summary_r %>% 
#   ggplot() + 
#   geom_bar(mapping = aes(x = order_var_2, y = prop_pop, fill = forcats::fct_reorder(pop_bucket, order_var, .desc = TRUE)), position = "stack", stat = "identity", height = 1, color = "black", size = 1) +
#    scale_fill_manual(values = c("<1 year" = "burlywood1", "1 - <3" = "orange", "3 - <5" = "red", "5 - 7" = "darkred", ">7 years" = "black")) +
#   labs(x = "", y = "", title = "Proportion population in different life years lost buckets", fill = "Life Years Lost") +
#   coord_flip() +
#   ggthemes::theme_clean() +
#   theme(plot.title = element_text(size = 15), 
#         axis.text.y = element_blank(), 
#         axis.text.x = element_blank(),
#         axis.ticks.x = element_blank(), 
#         axis.title = element_text(size = 11), 
#         legend.position = "bottom", 
#         legend.background = element_rect(), 
#         axis.line.x = element_blank(), 
#         axis.line.y = element_blank(), 
#         axis.ticks = element_blank()) 
# 
# 
# # v2.3 (stacked, rural, urban division)
# foo3 <- aqli_2020_district_level_pop_buckets_summary_u %>% 
#   ggplot() + 
#   geom_bar(mapping = aes(x = order_var_2, y = prop_pop, fill = forcats::fct_reorder(pop_bucket, order_var, .desc = TRUE)), position = "stack", stat = "identity", height = 1, color = "black", size = 1) +
#    scale_fill_manual(values = c("<1 year" = "burlywood1", "1 - <3" = "orange", "3 - <5" = "red", "5 - 7" = "darkred", ">7 years" = "black")) +
#   labs(x = "", y = "", title = "Proportion population in different life years lost buckets", fill = "Life Years Lost") +
#   coord_flip() +
#   ggthemes::theme_clean() +
#   theme(plot.title = element_text(size = 15), 
#         axis.text.y = element_blank(), 
#         axis.text.x = element_blank(),
#         axis.ticks.x = element_blank(), 
#         axis.title = element_text(size = 11), 
#         legend.position = "bottom", 
#         legend.background = element_rect(), 
#         axis.line.x = element_blank(), 
#         axis.line.y = element_blank(), 
#         axis.ticks = element_blank()) 
# 
# 
# 
# # v3 (waffle package, waffle plot), # problem is that we have to round up to a whole number
# x <- round(c(0.87, 30, 25.7, 21.2, 22.2))
# waffle(x, rows = 10)
# 
# # v4 (lessR, donut chart):this will not work
# PieChart(pop_bucket, data = aqli_2020_district_level_pop_buckets)
# 
# # v5 (treemapify)
# ggplot(data = aqli_2020_district_level_pop_buckets_summary, aes(area = prop_pop, fill = prop_pop)) +
#   geom_treemap() +
#   viridis::scale_fill_viridis(direction = -1, option = "rocket")
# 

```






# Figure 3: Pollution in top 5 most polluted districts not touched by NCAP
```{r}
# lyl in the top 5 most polluted districts not touched by NCAP
top5_most_pol_dist_non_ncap_india <- color_2020 %>%
  filter(country == "India") %>%
  filter(name_2 %notin% ncap_regions_list$district) %>%
  slice_max(pm2020, n = 5) %>%
  mutate(name_2_1 = str_c(name_2, " ", "(", name_1, ")"), 
         lyl_rel_who = round((pm2020 - 5)*0.098, 1), 
         lyl_rel_who = ifelse(lyl_rel_who < 0, 0, lyl_rel_who)) %>% 
  select(country:name_2, name_2_1, population, pm2020, lyl_rel_who)

top5_most_pol_dist_non_ncap_india %>%
  write_csv("./june.2022/clean.air.countdown.india.campaign.2023/chart.6.data.graph.code/plot6_data_part1.csv")
  
  
top5_most_pol_dist_non_ncap_india_part1_plot <- top5_most_pol_dist_non_ncap_india %>%
    ggplot(mapping = aes(x = forcats::fct_reorder(name_2_1, lyl_rel_who), y = lyl_rel_who, fill = lyl_rel_who)) +
  geom_col(width = 0.2, fill = "darkred") +
  labs(fill = "District (State)", title = "Life years lost to pollution in top 5 most polluted Indian districts, not touched by NCAP*", x = "District (State)", y = "Life years lost", 
       caption = expression("*NCAP is India's National Clean Air Programme. Pollution values specifically refer to annual average" ~ PM[2.5] ~ pollution ~ "and are reported in µg/m³. Life years lost are reported relative to WHO" ~ PM[2.5] ~ pollution ~ "guideline.")) +
    ggthemes::theme_stata() +
      theme(axis.line.x = element_line(), 
        axis.line.y = element_line(), 
        legend.position = "bottom", 
        plot.title = element_text(size = 21), 
        legend.box.background = element_rect(linetype = 1, size = 0.5), 
        axis.title.y = element_text(size = 16), 
        axis.title.x = element_text(size = 16),
        legend.text = element_text(size = 12), 
        plot.background = element_rect(fill = "white"), 
        plot.caption = element_text(hjust = 0, size = 14), plot.caption.position = "plot", 
        axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 13), 
        axis.ticks.x = element_blank(), 
        axis.ticks.y = element_blank()) +
   geom_text(mapping = aes(label = lyl_rel_who), position = position_dodge(width = 1), vjust = -0.25, size = 5) +
   scale_y_continuous(breaks = seq(0, 11, 1))



svglite::svglite("final_fig6_part1_cleanAirIndiaCountdown2023.svg")
ggsave("./final_fig6_part1_cleanAirIndiaCountdown2023.svg", width = 16, height = 7)

#> creating part 2 of the graph: an inset map

# states of interest
states_of_interest_ch6 <- unique(top5_most_pol_dist_non_ncap_india$name_1)

# states of interest shape file
states_of_focus_shp_file_ch6 <- india_state %>%
  filter(NAME_1 %in% states_of_interest_ch6)

# districts of interest (within the above states of interest) shape file
districts_of_interst_shp_file_ch6 <- colormap_shp %>%
  filter(NAME_0 == "India", NAME_1 %in% states_of_interest_ch6, NAME_2 %in% top5_most_pol_dist_non_ncap_india$name_2) %>% 
  left_join(color_2020 %>% filter(country == "India"), by = c("NAME_0" = "country", "NAME_1" = "name_1", "NAME_2" = "name_2")) %>%
  select(NAME_0, NAME_1, NAME_2, population, geometry) %>% 
  mutate(pop_lakhs = round(population/100000, 1),
    graph_label = str_c(NAME_2, " (", NAME_1, "), ", pop_lakhs, " lakh people")) %>%
  select(-geometry, geometry)

# states of interest bounding box-
states_of_interest_bb_ch6 <- st_as_sfc(st_bbox(states_of_focus_shp_file_ch6))

# a state level India map with a bounding box on states of interest
figure_6_plot_part_2 <- india_state %>%
  ggplot() +
  geom_sf(fill = "snow2") +
  geom_sf(data = states_of_focus_shp_file_ch6, fill = "white", size = 0.9) +
  geom_sf(data = districts_of_interst_shp_file_ch6, fill = "red", size = 0.8) +
  geom_sf(data = states_of_interest_bb_ch6, fill = NA, color = "red", size = 0.8) +
  theme_map() 

india_state %>% st_write("./june.2022/clean.air.countdown.india.campaign.2023/chart.6.data.graph.code/plot6_data_part2.shp")

svglite::svglite("final_fig6_part2_cleanAirIndiaCountdown2023.svg")
ggsave("./final_fig6_part2_cleanAirIndiaCountdown2023.svg", width = 7, height = 7)

figure_6_plot_part_3 <- states_of_focus_shp_file_ch6 %>%
  ggplot() + 
    geom_sf(fill = "white", size = 0.8, color = "black") +
   geom_sf(data = districts_of_interst_shp_file_ch6, fill = "red", size = 1.2) +
  theme_map() +
  labs(title = "") +
  theme_map() +
    theme(plot.background = element_rect(fill = "azure2"), 
        plot.caption = element_text(hjust = 0, size = 6), plot.caption.position = "plot") 

+
   ggsflabel::geom_sf_label_repel(data = districts_of_interst_shp_file_ch6, mapping = aes(label = graph_label), force = 90, nudge_x = -2, seed = 10, size = 6) 
  
states_of_focus_shp_file %>%
  st_write("./june.2022/clean.air.countdown.india.campaign.2023/chart.6.data.graph.code/plot6_data_part3.shp")

 svglite::svglite("final_fig6_part3_cleanAirIndiaCountdown2023.svg")
ggsave("./final_fig6_part3_cleanAirIndiaCountdown2023.svg", width = 9, height = 7)
 

# combined inset map
figure_3_plot <- cowplot::ggdraw() +
  draw_plot(figure_3_plot_part_2, x = 0.1, y = 0.0001) +
  draw_plot(figure_3_plot_part_1, x = 0, y = 0.1, width = 0.4, height = 0.4)

  
 



```




############# Other graphs ##########################




# Figure 6:  5 Indian districts with the greatest increase in pollution over the past five years
```{r}

# five datasets with greatest increase
five_districts_with_greatest_increase <- color_2020 %>%
  filter(country == "India") %>%
  select(country, name_1, name_2, pm2016, pm2020) %>%
  mutate(district_state = str_c(name_2, " ", "(", name_1, ")"), 
         pol_2020_vs_2016  = pm2020 - pm2016) %>%
  select(country, name_1, name_2, district_state, everything()) %>%
   slice_max(pol_2020_vs_2016, n = 5) %>%
  select(district_state) %>%
  unlist() %>%
  as.vector()

# figure 1 data
figure_1_data <- color_2020 %>%
  filter(country == "India") %>%
  select(country, name_1, name_2, pm2016, pm2020) %>%
  mutate(district_state = str_c(name_2, " ", "(", name_1, ")"), 
         pol_2020_vs_2016  = pm2020 - pm2016) %>%
  select(country, name_1, name_2, district_state, everything()) %>%
  filter(district_state %in% five_districts_with_greatest_increase) %>%
  select(-pol_2020_vs_2016) %>%
  pivot_longer(pm2016:pm2020, names_to = "year", values_to = "pm2.5_pollution") %>%
  mutate(year = str_extract(year, "\\d+"))
  
# figure 1 plot  
figure_1_plot <- figure_1_data %>%
  ggplot() +
  geom_col(mapping = aes(x = year, y = pm2.5_pollution, fill = district_state), width = 0.5) +
  facet_wrap(~district_state, ncol = 5) +
  theme(legend.position = "bottom") +
  labs(fill = "District (State)", title = "5 Indian districts with the greatest increase in pollution over the past five years") +
   scale_fill_brewer(palette = "Accent") +
  ggthemes::theme_hc()
 

figure_1_plot


```

# Figure 7: 5 Indian cities with the greatest reduction in pollution over the past five years

```{r}

# five districts with greatest decrease in pollution
five_districts_with_greatest_decrease <- color_2020 %>%
  filter(country == "India") %>%
  select(country, name_1, name_2, pm2016, pm2020) %>%
  mutate(district_state = str_c(name_2, " ", "(", name_1, ")"), 
         pol_2020_vs_2016  = pm2020 - pm2016) %>%
  select(country, name_1, name_2, district_state, everything()) %>%
   slice_min(pol_2020_vs_2016, n = 5) %>%
  select(district_state) %>%
  unlist() %>%
  as.vector()


figure_2_data <- color_2020 %>%
  filter(country == "India") %>%
  select(country, name_1, name_2, pm2016, pm2020) %>%
  mutate(district_state = str_c(name_2, " ", "(", name_1, ")"), 
         pol_2020_vs_2016  = pm2020 - pm2016) %>%
  select(country, name_1, name_2, district_state, everything()) %>%
  filter(district_state %in% five_districts_with_greatest_decrease) %>%
  select(-pol_2020_vs_2016) %>%
  pivot_longer(pm2016:pm2020, names_to = "year", values_to = "pm2.5_pollution") %>%
  mutate(year = str_extract(year, "\\d+"))
  
  
figure_2_plot <- figure_2_data %>%
  ggplot() +
  geom_col(mapping = aes(x = year, y = pm2.5_pollution, fill = district_state), width = 0.5) +
  facet_wrap(~district_state, ncol = 5) +
  theme(legend.position = "bottom") +
  labs(fill = "District (State)", title = "5 Indian districts with the greatest reduction in pollution over the past five years") +
   scale_fill_brewer(palette = "Accent") +
  ggthemes::theme_hc()
 


```



# Figure 8: Pollution levels in the top 5 most polluted districts in India not touched by NCAP (bar graph)

```{r}


# figure 8 plot
figure_5_plot <- non_ncap_district_top5_most_populated_data  %>% # data from figure 3
  ggplot() +
  geom_col(mapping = aes(x = fct_reorder(name_2, pm2020), y = pm2020, fill = name_2), width = 0.3) +
  labs(x = "District", y = "PM2.5 pollution", fill = "District", 
       title = "Pollution levels in the top 5 most 'polluted' districts in India not touched by NCAP") +
  scale_fill_brewer(palette = "Accent") +
  scale_y_continuous(breaks = seq(0, 110, 15)) +
  ggthemes::theme_hc()
  

```

# Figure 9: Pollution levels in the top 5 most populated districts in India not touched by NCAP

```{r}

# figure 5 data
figure_6_data <- color_2020 %>%
  filter(country == "India") %>%
  filter(name_2 %notin% ncap_district_list) %>%
  slice_max(population, n = 5) 

# figure 5 plot
figure_6_plot <- figure_6_data %>%
  ggplot() +
  geom_col(mapping = aes(x = fct_reorder(name_2, pm2020), y = pm2020, fill = name_2), width = 0.3) +
  labs(x = "District", y = "PM2.5 pollution", fill = "District", 
       title = "Pollution levels in the top 5 most 'populated' districts in India not touched by NCAP") +
  scale_fill_brewer(palette = "Accent") +
  scale_y_continuous(breaks = seq(0, 110, 15)) +
  ggthemes::theme_hc()
```


# Figure 10: Top 5 NCAP districts that have seen a rise in pollution from 2019 to 2020

```{r}

districts_2019_2020_top5 <- color_2020 %>%
  filter(country == "India") %>%
  select(country, name_1, name_2, pm2019, pm2020) %>%
  mutate(district_state = str_c(name_2, " ", "(", name_1, ")"), 
         pol_2020_vs_2019  = pm2020 - pm2019) %>%
  select(country, name_1, name_2, district_state, everything()) %>%
  filter(name_2 %in% ncap_district_list) %>%
   slice_max(pol_2020_vs_2019, n = 5) %>%
  select(district_state) %>%
  unlist() %>%
  as.vector()
  
# figure 7 data
figure_7_data <- color_2020 %>%
  filter(country == "India") %>%
  select(country, name_1, name_2, pm2019, pm2020) %>%
  mutate(district_state = str_c(name_2, " ", "(", name_1, ")"), 
         pol_2020_vs_2019  = pm2020 - pm2019) %>%
  select(country, name_1, name_2, district_state, everything()) %>%
  filter(district_state %in% districts_2019_2020_top5) %>%
  select(-pol_2020_vs_2019) %>%
  pivot_longer(pm2019:pm2020, names_to = "year", values_to = "pm2.5_pollution") %>%
  mutate(year = str_extract(year, "\\d+"))
  
# figure 7 plot  
figure_7_plot <- figure_7_data %>%
  ggplot() +
  geom_col(mapping = aes(x = year, y = pm2.5_pollution, fill = district_state), size = 0.3) +
  facet_wrap(~district_state, ncol = 5) +
  theme(legend.position = "bottom") +
  labs(fill = "District (State)", title = "Top 5 NCAP districts that have seen a rise in pollution from 2019 to 2020") +
   scale_fill_brewer(palette = "Accent") +
  ggthemes::theme_hc()



```

# Figure 11: Top 5 NCAP districts that have seen a decline in pollution from 2019 to 2020

```{r}

districts_2019_2020_top5_dec <- color_2020 %>%
  filter(country == "India") %>%
  select(country, name_1, name_2, pm2019, pm2020) %>%
  mutate(district_state = str_c(name_2, " ", "(", name_1, ")"), 
         pol_2020_vs_2019  = pm2020 - pm2019) %>%
  select(country, name_1, name_2, district_state, everything()) %>%
  filter(name_2 %in% ncap_district_list) %>%
   slice_min(pol_2020_vs_2019, n = 5) %>%
  select(district_state) %>%
  unlist() %>%
  as.vector()
  
# figure 8 data
figure_8_data <- color_2020 %>%
  filter(country == "India") %>%
  select(country, name_1, name_2, pm2019, pm2020) %>%
  mutate(district_state = str_c(name_2, " ", "(", name_1, ")"), 
         pol_2020_vs_2019  = pm2020 - pm2019) %>%
  select(country, name_1, name_2, district_state, everything()) %>%
  filter(district_state %in% districts_2019_2020_top5_dec) %>%
  select(-pol_2020_vs_2019) %>%
  pivot_longer(pm2019:pm2020, names_to = "year", values_to = "pm2.5_pollution") %>%
  mutate(year = str_extract(year, "\\d+"))
  
# figure 8 plot  
figure_8_plot <- figure_8_data %>%
  ggplot() +
  geom_col(mapping = aes(x = year, y = pm2.5_pollution, fill = district_state), size = 0.3) +
  facet_wrap(~district_state, ncol = 5) +
  theme(legend.position = "bottom") +
  labs(fill = "District (State)", title = "Top 5 NCAP districts that have seen a decline in pollution from 2019 to 2020") +
   scale_fill_brewer(palette = "Accent") +
  ggthemes::theme_hc()



```


# Figure 12: Gujarat, trenlines pollution last 20 years.

```{r}
figure_9_data <- color_2020 %>%
  filter(country == "India") %>%
  group_by(name_1) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         mutate(across(starts_with("pm"), ~.x*pop_weights, .names = "{col}_weighted"))) %>%
  summarise(across(ends_with("weighted"), sum)) %>%
  pivot_longer(cols = pm1998_weighted:pm2020_weighted, names_to = "years", 
               values_to = "pop_weighted_avg_pm2.5") %>%
  mutate(years = as.integer(unlist(str_extract(years, "\\d+")))) %>%
  select(years, name_1, pop_weighted_avg_pm2.5)


  plt <- figure_9_data %>% 
    filter(name_1 == "Gujarat") %>%
    ggplot() +
  geom_line(mapping = ggplot2::aes(x = years, y = pop_weighted_avg_pm2.5), lwd = 1.1, 
            color = "darkred") +
    geom_hline(mapping = aes(yintercept = 5), lwd = 0.8) +
  scale_y_continuous(breaks = seq(0, 40, 5), limits = c(0, 40)) +
  scale_x_continuous(breaks = seq(1998, 2020, 2)) +
  ggthemes::theme_clean() +
  labs(x = "Years", 
       y = expression(paste("Average PM2.5 concentration ( ", mu, "g", "/", m^3, " )")), 
       title = "Average PM2.5 pollution in the state of Gujarat, India", 
       subtitle = "1998 to 2020") +
  theme(legend.position = "bottom", legend.title = element_blank(), 
        legend.text = element_text(size = 7), 
        axis.title.y = element_text(size = 9), 
        axis.title.x = element_text(size = 9)) + 
     geom_text(x = 2002.5, y = 7.2, label = expression(paste("WHO PM2.5 Guideline (last updated: 2021): 5 ", mu, "g","/", m^3, "")))

```

# Figure 13, Surat pollution 1998 to 2020

```{r}
figure_10_data <- color_2020 %>%
  filter(country == "India", name_1 == "Gujarat", name_2 == "Surat") %>%
  select(country, name_1, name_2, starts_with("pm")) %>%
  pivot_longer(cols = pm1998:pm2020, names_to = "years", 
               values_to = "avg_pm2.5") %>%
  mutate(years = as.integer(unlist(str_extract(years, "\\d+")))) 


  figure_10_plot <- figure_10_data %>% 
    ggplot() +
  geom_line(mapping = ggplot2::aes(x = years, y = avg_pm2.5), lwd = 1.1, 
            color = "darkred") +
    geom_hline(mapping = aes(yintercept = 5), lwd = 0.8) +
  scale_y_continuous(breaks = seq(0, 40, 5), limits = c(0, 40)) +
  scale_x_continuous(breaks = seq(1998, 2020, 2)) +
  ggthemes::theme_clean() +
  labs(x = "Years", 
       y = expression(paste("Average PM2.5 concentration ( ", mu, "g", "/", m^3, " )")), 
       title = "Average PM2.5 pollution in Surat district of Gujarat, India", 
       subtitle = "1998 to 2020") +
  theme(legend.position = "bottom", legend.title = element_blank(), 
        legend.text = element_text(size = 7), 
        axis.title.y = element_text(size = 9), 
        axis.title.x = element_text(size = 9)) + 
     geom_text(x = 2002.5, y = 7.2, label = expression(paste("WHO PM2.5 Guideline (last updated: 2021): 5 ", mu, "g","/", m^3, "")))

```


# Random
```{r}
# 
 color_2020 %>%
  filter(country == "India") %>%
  group_by(name_1) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2020_pop_weighted = pm2020*pop_weights, 
         pm2016_pop_weighted = pm2016*pop_weights) %>%
  summarise(avg_pm2.5_2016 = sum(pm2016_pop_weighted, na.rm = TRUE), 
            avg_pm2.5_2020 = sum(pm2020_pop_weighted, na.rm = TRUE)) %>%
  mutate(avg_pm2.5_2020_vs_2016 = avg_pm2.5_2020 - avg_pm2.5_2016, 
         avg_pm2.5_2020_vs_2016_prop_change = round((avg_pm2.5_2020_vs_2016/avg_pm2.5_2016)*100, 2)) %>%
  filter(avg_pm2.5_2020_vs_2016_prop_change <= -10) %>%
  nrow()


color_2020 %>%
  filter(country == "India") %>%
  mutate(pol_change = pm2020 - pm2016, 
         prop_change = (pol_change/pm2016)*100) %>%
  filter(prop_change < -10) %>%
  nrow()

# Michael graph (top 5 largest NCAP cities, pollution trend graph 2017 to 2020) and their corresponding pollution reduction targets horizontal lines.

ds_trendlines <- color_2020 %>%
  filter(name_2 %in% ncap_regions_list$district) %>%
  slice_max(population, n = 5) %>%
  select(country, name_1, name_2, population, starts_with("pm")) %>%
  pivot_longer(pm1998:pm2020, names_to = "year", values_to = "pm2.5_pollution") %>%
  mutate(year = as.numeric(str_extract(year, "\\d+"))) %>%
  filter(year >= 2017) 

city_wise_reduction_target <- color_2020 %>%
  filter(country == "India", name_2 %in% c("Bangalore", "Mumbai Suburban", "NCT of Delhi", "North 24 Parganas", "Pune")) %>%
  select(country, name_1, name_2, pm2017) %>%
  mutate(pm2026_target = round(pm2017*0.6, 2))

foo <- ds_trendlines %>%
  ggplot() +
  geom_line(mapping = ggplot2::aes(x = year, y = pm2.5_pollution, color = name_2), lwd = 1.2) +
    geom_hline(mapping = aes(yintercept = 23.6), lwd = 1.1, linetype = "dashed", color = "olivedrab1") +
  geom_hline(mapping = aes(yintercept = 22.2), lwd = 1.1, linetype = "dashed", color = "darkslateblue") +
  geom_hline(mapping = aes(yintercept = 24.9), lwd = 1.1, linetype = "dashed", color = "orangered4") +
  geom_hline(mapping = aes(yintercept = 65.6), lwd = 1.1, linetype = "dashed", color = "magenta1") +
  geom_hline(mapping = aes(yintercept = 34.3), lwd = 1.1, linetype = "dashed", color = "lightsteelblue4") +
  geom_vline(mapping = aes(xintercept = 2019), lwd = 0.8) +
  scale_y_continuous(breaks = seq(0, 130, 20), limits = c(0, 130)) +
  scale_x_continuous(breaks = seq(2017, 2020, 1)) +
  labs(x = "Years", 
       y = expression(paste("Average PM2.5 concentration ( ", mu, "g", "/", m^3, ")")), 
       title = "Average PM2.5 pollution in 5 largest districts touched by NCAP, with 2026 reduction targets", 
       subtitle = "2017 to 2020", caption = "* 'largest NCAP districts' means, largest in terms of population and dashed lines show city wise reduction targets", 
       color = "District") +
    theme_classic() +
  theme(legend.position = "bottom", legend.title = element_blank(), 
        legend.text = element_text(size = 10), 
        axis.title.y = element_text(size = 11), 
        axis.title.x = element_text(size = 11),
        axis.text.x = element_text(size = 13), 
        axis.text.y = element_text(size = 13), 
        plot.caption = element_text(hjust = 0)) +
  scale_color_manual(values = c("Bangalore" = "olivedrab1", "Mumbai Suburban" = "darkslateblue", "NCT of Delhi" = "magenta1", "North 24 Parganas" = "lightsteelblue4", "Pune" = "orangered4"))


foo1 <- ds_trendlines %>%
  ggplot() +
  geom_line(mapping = ggplot2::aes(x = year, y = pm2.5_pollution, color = name_2), lwd = 1.1) +
    geom_hline(mapping = aes(yintercept = 23.6), lwd = 0.3, color = "orange4") +
  geom_hline(mapping = aes(yintercept = 22.2), lwd = 0.3,  color = "lightblue4") +
  geom_hline(mapping = aes(yintercept = 24.9), lwd = 0.3,  color = "mediumpurple4") +
  geom_hline(mapping = aes(yintercept = 65.6), lwd = 0.3, color = "darkred") +
  geom_hline(mapping = aes(yintercept = 34.3), lwd = 0.3, color = "navyblue") +
  geom_vline(mapping = aes(xintercept = 2019), lwd = 0.8) +
  scale_y_continuous(breaks = seq(0, 130, 20), limits = c(0, 130)) +
  scale_x_continuous(breaks = seq(2017, 2020, 1)) +
  labs(x = "Years", 
       y = expression(paste("Average PM2.5 concentration ( ", mu, "g", "/", m^3, ")")), 
       title = "Average PM2.5 pollution in 5 largest districts touched by NCAP", 
       subtitle = "2017 to 2020", caption = "* 'largest NCAP districts' means, largest in terms of population", 
       color = "District") +
    theme_classic() +
  theme(legend.position = "bottom", legend.title = element_blank(), 
        legend.text = element_text(size = 10), 
        axis.title.y = element_text(size = 11), 
        axis.title.x = element_text(size = 11),
        axis.text.x = element_text(size = 13), 
        axis.text.y = element_text(size = 13), 
        plot.caption = element_text(hjust = 0)) +
  scale_color_manual(values = c("Bangalore" = "orange4", "Mumbai Suburban" = "blue4", "NCT of Delhi" = "darkred", "North 24 Parganas" = "navyblue", "Pune" = "mediumpurple4"))  


foo2 <- ds_trendlines %>%
  ggplot() +
  geom_line(mapping = ggplot2::aes(x = year, y = pm2.5_pollution, color = name_2), lwd = 1.5) +
  geom_hline(data = city_wise_reduction_target, mapping = aes(yintercept = pm2026_target, color = name_2), lwd = 0.8, linetype = "dashed",  color = "lightblue4") +
  geom_vline(mapping = aes(xintercept = 2019), lwd = 0.8) +
  scale_y_continuous(breaks = seq(0, 130, 20), limits = c(0, 130)) +
  scale_x_continuous(breaks = seq(2017, 2020, 1)) +
  labs(x = "Years", 
       y = expression(paste("Average PM2.5 concentration ( ", mu, "g", "/", m^3, ")")), 
       title = "Average PM2.5 pollution in 5 largest districts touched by NCAP, with 2026 reduction targets", 
       subtitle = "2017 to 2020, NCAP begins in 2019 (vertical line)", caption = "* 'largest NCAP districts' means, largest in terms of population and dashed lines show city wise reduction targets", 
       color = "District") +
    ggthemes::theme_clean() +
  theme(legend.position = "bottom", legend.title = element_blank(), 
        legend.text = element_text(size = 10), 
        axis.title.y = element_text(size = 11), 
        axis.title.x = element_text(size = 11),
        axis.text.x = element_text(size = 9), 
        axis.text.y = element_text(size = 9), 
        plot.caption = element_text(hjust = 0), 
        plot.title = element_text(size = 13), 
        plot.subtitle = element_text(size = 11)) +
  facet_wrap(~name_2, ncol = 5) 

+
  scale_color_manual(values = c("Bangalore" = "orange4", "Mumbai Suburban" = "blue4", "NCT of Delhi" = "darkred", "North 24 Parganas" = "navyblue", "Pune" = "mediumpurple4"))  

```


```{r}

bbox_state_of_interest <- st_as_sfc(st_bbox(colormap_shp %>% filter(NAME_0 == "India", NAME_1 %in% c("Uttar Pradesh", "Haryana", "NCT of Delhi"))))

india_state %>%
  filter(NAME_1 %notin% c("Andaman and Nicobar")) %>%
  ggplot() +
  geom_sf(fill = "white", color = "black", size = 1)  +
  ggthemes::theme_map() +
  geom_sf(data = india_state %>% filter(NAME_1 %in% c("Uttar Pradesh", "Haryana", "NCT of Delhi")), size = 1.5, fill = "white", color = "black") +
  geom_sf(data = bbox_state_of_interest, fill = NA, color = "red", size = 1)


districts_of_interest <- ggplot() +
  geom_sf(data = colormap_shp %>% filter(NAME_0 == "India", NAME_1 %in% c("Uttar Pradesh", "Haryana", "NCT of Delhi"), NAME_2 %in% c("Ghaziabad", "Gautam Buddha Nagar", "Muzaffarnagar", "Gurugram", "Panipat", "Faridabad")), color = "orange")




foo1 <- ggplot() + 
   geom_sf(data = india_state, fill = "white", color = "black", size = 1.2) +
  geom_sf(data = india_state %>% filter(NAME_1 %in% c("Uttar Pradesh", "NCT of Delhi", "Haryana")), color = "black", size = 1, fill = "white") +
  geom_sf(data = colormap_shp %>% filter(NAME_0 == "India", NAME_1 %in% c("Uttar Pradesh", "Haryana", "NCT of Delhi"), NAME_2 %in% c("Ghaziabad", "Gautam Buddha Nagar", "Muzaffarnagar", "Gurugram", "Panipat", "Faridabad", "NCT of Delhi")), color = "orange") +
   geom_sf(data = bbox_state_of_interest, fill = "NA", color = "red", size = 1.2) +
  theme_map()

 

foo2 <- ggplot() +
  geom_sf(data = india_state %>% filter(NAME_1 %in% c("Uttar Pradesh", "NCT of Delhi", "Haryana")), color = "black", size = 1, fill = "white") +
   geom_sf(data = colormap_shp %>% filter(NAME_0 == "India", NAME_1 %in% c("Uttar Pradesh", "Haryana", "NCT of Delhi"), NAME_2 %in% c("Ghaziabad", "Gautam Buddha Nagar", "Muzaffarnagar", "Gurugram", "Panipat", "Faridabad", "NCT of Delhi")), color = "black", fill = "orange", size = 1.2) +
  ggsflabel::geom_sf_label_repel(data = colormap_shp %>% filter(NAME_0 == "India", NAME_1 %in% c("Uttar Pradesh", "Haryana", "NCT of Delhi"), NAME_2 %in% c("Ghaziabad", "Gautam Buddha Nagar", "Muzaffarnagar", "Gurugram", "Panipat", "Faridabad", "NCT of Delhi")), mapping = aes(label = NAME_2), force = 100, nudge_x = -2, seed = 10, size = 7) +
  theme_map()
 
 
 
 inset_map_wb_with_5_most_pop_non_ncap_districts <- cowplot::ggdraw() +
  draw_plot(foo2) +
  draw_plot(foo1, x = 0, y = 0.1, width = 0.4, height = 0.4)
 
 
```


```{r}

foo <- read_csv("./foo.csv")

foo %>%
  pivot_longer(raasman:safar, names_to = "type", values_to = "pol") %>%
  ggplot() +
  geom_line(mapping = aes(x = Date, y = pol, color = type, size = 0.2))
  

```





























































