---
title: "Factsheets 2022"
author: "Aarsh"
date: '2022-07-26'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# get packages, global variables and datasets
```{r load-lib-helper-files-data-global-var}

# read in the helper file
source("./appPublic/aqli.data.explorer.helper.script.R")

```

# This script contains code for graphs and some calculations that were not computed by the
# AQLI Data Explorer dashboard: https://aarsh.shinyapps.io/aqli-epic-data-explorer/.

#========================================================================================================


## India Factsheet 

```{r india_fs_2023}

#> Since [year], [x%] of the world's increase in pollution camer from [country]-----------------

country_select <- "India"

# Since 2013 all gadm2 places where pollution has increased - > average pm2.5 in 2013, 2021
 avg_pm2.5_global <- gadm2_aqli_2021 %>% 
  mutate(diff_pm2021_pm2013 = pm2021 - pm2013) %>%
  filter(diff_pm2021_pm2013 > 0) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2013_pop_weighted = pm2013*pop_weights, 
         pm2021_pop_weighted = pm2021*pop_weights) %>%
  summarise(avg_pm2.5_2013 = sum(pm2013_pop_weighted, na.rm = TRUE), 
            avg_pm2.5_2021 = sum(pm2021_pop_weighted, na.rm = TRUE))

 actual_diff_global <- avg_pm2.5_global$avg_pm2.5_2021 - avg_pm2.5_global$avg_pm2.5_2013
 
# Assuming that of the places (gadm2 regions) where pollution increased, if amongst those places, 
# all Indian gadm2 regions had the same pollution in 2021 as it had in 2013, how would the global
# average PM2.5 look in 2021.
 
avg_pm2.5_global_adjusted <- gadm2_aqli_2021 %>% 
  mutate(diff_pm2021_pm2013 = pm2021 - pm2013) %>%
  filter(diff_pm2021_pm2013 > 0) %>%
   mutate(pm2021 = ifelse(country == country_select, pm2013, pm2021)) %>%
     mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2013_pop_weighted = pm2013*pop_weights, 
         pm2021_pop_weighted = pm2021*pop_weights) %>%
  summarise(avg_pm2.5_2013 = sum(pm2013_pop_weighted, na.rm = TRUE), 
            avg_pm2.5_2021 = sum(pm2021_pop_weighted, na.rm = TRUE))

# difference between adjusted and actual global average, i.e. India's contribution
adjusted_diff_global <- avg_pm2.5_global_adjusted$avg_pm2.5_2021 - avg_pm2.5_global_adjusted$avg_pm2.5_2013
 
# final answer
final_answer <- round(((actual_diff_global - adjusted_diff_global)/actual_diff_global)*100, 1)

#> Indo Gangetic Plain, population and average PM2.5----------------------------------

# average pm2.5 and total population in Indo-Gangetic plains in 2021
summary_indo_gangetic_plain <- gadm2_aqli_2021 %>%
  filter(country == "India", name_1 %in% indo_gangetic_plains_states) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2021_pop_weighted = pm2021*pop_weights) %>%
  summarise(avg_pm2.5_2021 = sum(pm2021_pop_weighted, na.rm = TRUE), 
            total_population = sum(population, na.rm = TRUE))

#> Fig 1: Potential gain in Life Expectancy from permanently reducing PM2.5 from 2021 levels to the-------- 
# WHO PM2.5 guideline.

# separately getting Gilgit Baltistan and Azad Kashmir
azad_kashmir_gilgit_baltistan <- gadm2_aqli_2021 %>%
  filter(country == "Pakistan", name_1 %in% c("Azad Kashmir", "Gilgit Baltistan"))

# plot 3 data
india_fs_fig1_data <- gadm2_aqli_2021 %>%
  filter(country == "India") %>%
  bind_rows(azad_kashmir_gilgit_baltistan) %>%
  left_join(gadm2_aqli_2021_shp, by = c("objectid_gadm2" = "obidgadm2")) %>%
  mutate(lyl_bucket = ifelse((llpp_who_2021 >= 0) & (llpp_who_2021 < 0.1), "0 - < 0.1", NA), 
         lyl_bucket = ifelse((llpp_who_2021 >= 0.1) & (llpp_who_2021 <= 0.5), "0.1 - 0.5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 0.5) & (llpp_who_2021 <= 1), "> 0.5 - 1", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 1) & (llpp_who_2021 <= 2), "> 1 - 2", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 2) & (llpp_who_2021 <= 3), "> 2 - 3", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 3) & (llpp_who_2021 <= 4), "> 3 - 4", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 4) & (llpp_who_2021 <= 5), "> 4 - 5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 5) & (llpp_who_2021 < 6), "> 5 - < 6", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 >= 6), ">= 6", lyl_bucket)) %>%
  mutate(order_lyl_bucket = ifelse(lyl_bucket == "0 - < 0.1", 1, NA), 
         order_lyl_bucket = ifelse(lyl_bucket == "0.1 - 0.5", 2, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 0.5 - 1", 3, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 1 - 2", 4, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 2 - 3", 5, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 3 - 4", 6, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 4 - 5", 7, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 5 - < 6", 8, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == ">= 6", 9, order_lyl_bucket)) %>%
  select(-geometry, geometry) %>%
  st_as_sf()

# change lyl_bucket and order_lyl_bucket columns to factor
# india_fs_fig1_data$lyl_bucket <- factor(india_fs_fig1_data$lyl_bucket)
# india_fs_fig1_data$order_lyl_bucket <- factor(india_fs_fig1_data$order_lyl_bucket)

# india factsheet figure 1
india_fs_fig1 <- india_fs_fig1_data %>%
  ggplot() +
  geom_sf(mapping = aes(fill = forcats::fct_reorder(lyl_bucket, order_lyl_bucket)), color = "lightgrey", lwd = 0.05) +
  geom_sf(data = india_state, color = "black", fill = "transparent", lwd = 0.5) +
  ggthemes::theme_map() + 
 scale_fill_manual(values = c("0 - < 0.1 years" = "#FFFFFF", 
                               "0.1 - 0.5" = "#FFE6B3", 
                               "> 0.5 - 1" = "#FFD25D", 
                               "> 1 - 2" = "#FFBA00", 
                               "> 2 - 3" = "#FF9600", 
                               "> 3 - 4" = "#FF6908", 
                               "> 4 - 5" = "#E63D23", 
                               "> 5 - < 6" = "#BD251C", 
                               ">= 6" = "#8C130E")) +
  ggthemes::theme_map() +
  labs(fill = "Gain in years of life expectancy  ", title = expression("Potential Gains in Life Expectancy (India) through Permanently Reducing" ~ PM[2.5] ~ "from 2021 concentrations to the WHO guideline")) + 
 theme(legend.position = "bottom", 
        legend.justification = c(0.5, 3), 
        legend.background = element_rect(color = "black"), 
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15), 
        plot.title = element_text(hjust = 0.5, size = 15), 
       # legend.key = element_rect(color = "black"),
       legend.box.margin = margin(b = 1, unit = "cm"),
        plot.subtitle = element_text(hjust = 0.5, size = 7), 
        plot.caption = element_text(hjust = 0.7, size = 9, face = "italic"), 
       legend.key = element_rect(color = "black"), 
       legend.box.spacing = unit(0, "cm"), 
        legend.direction = "horizontal", 
       plot.background = element_rect(fill = "white", color = "white")) +
  guides(fill = guide_legend(nrow = 1))


svglite("india_fs_fig1")
ggsave("./september.2023/factsheets/india/india_fs_fig1.svg", width = 15, height = 10)

india_fs_fig1_data %>% write_sf("./september.2023/factsheets/india/india_fs_fig1.shp")




# Figure 1.1: state wise figures (uses figure 1's data)-------------------------------------------

# separately getting Gilgit Baltistan and Azad Kashmir
azad_kashmir_gilgit_baltistan <- gadm2_aqli_2021 %>%
  filter(country == "Pakistan", name_1 %in% c("Azad Kashmir", "Gilgit Baltistan"))

# plot 3 data
india_fs_fig1.1_data <- gadm2_aqli_2021 %>%
  filter(country == "India") %>%
  bind_rows(azad_kashmir_gilgit_baltistan) %>%
  left_join(gadm2_aqli_2021_shp, by = c("objectid_gadm2" = "obidgadm2")) %>%
  mutate(lyl_bucket = ifelse((llpp_who_2021 >= 0) & (llpp_who_2021 < 0.1), "0 - < 0.1", NA), 
         lyl_bucket = ifelse((llpp_who_2021 >= 0.1) & (llpp_who_2021 <= 0.5), "0.1 - 0.5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 0.5) & (llpp_who_2021 <= 1), "> 0.5 - 1", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 1) & (llpp_who_2021 <= 2), "> 1 - 2", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 2) & (llpp_who_2021 <= 3), "> 2 - 3", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 3) & (llpp_who_2021 <= 4), "> 3 - 4", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 4) & (llpp_who_2021 <= 5), "> 4 - 5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 5) & (llpp_who_2021 < 6), "> 5 - < 6", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 >= 6), ">= 6", lyl_bucket)) %>%
  mutate(order_lyl_bucket = ifelse(lyl_bucket == "0 - < 0.1", 1, NA), 
         order_lyl_bucket = ifelse(lyl_bucket == "0.1 - 0.5", 2, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 0.5 - 1", 3, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 1 - 2", 4, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 2 - 3", 5, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 3 - 4", 6, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 4 - 5", 7, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 5 - < 6", 8, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == ">= 6", 9, order_lyl_bucket)) %>%
  select(-geometry, geometry) %>%
  st_as_sf()



states_to_plot <- c("Jammu and Kashmir", "Haryana", "Andhra Pradesh", "Assam", "West Bengal", "Maharashtra", 
                "Tamil Nadu", "Azad Kashmir")

# india factsheet figure 1.1 (state wise)
india_fs_fig1.1 <- india_fs_fig1.1_data %>%
  filter(name_1 %in% states_to_plot) %>%
  ggplot() +
  geom_sf(mapping = aes(fill = forcats::fct_reorder(lyl_bucket, order_lyl_bucket)), color = "transparent") +
  geom_sf(data = india_state, color = "grey", fill = "transparent", lwd = 0.1) +
  geom_sf(data = india_state %>% filter(NAME_1 %in% states_to_plot), color = "black", fill = "transparent", lwd = 0.8) +
    geom_sf(data = india_fs_fig1.1_data %>% filter(name_2 == "Chennai"), color = "black", lwd = 1) +
   ggthemes::theme_map() +
    ggsflabel::geom_sf_label_repel(data = india_state %>% filter(NAME_1 %in% states_to_plot), mapping = aes(label = NAME_1), force = 50, nudge_x = -10, nudge_y = -1.6, seed = 10, size = 3.5) +
   ggsflabel::geom_sf_label_repel(data = india_fs_fig1.1_data %>% filter(name_2 == "Chennai"), mapping = aes(label = name_2), force = 50, nudge_x = 3, seed = 10, size = 2.8) +
  ggthemes::theme_map() + 
 scale_fill_manual(values = c("0 - < 0.1 years" = "#FFFFFF", 
                               "0.1 - 0.5" = "#FFE6B3", 
                               "> 0.5 - 1" = "#FFD25D", 
                               "> 1 - 2" = "#FFBA00", 
                               "> 2 - 3" = "#FF9600", 
                               "> 3 - 4" = "#FF6908", 
                               "> 4 - 5" = "#E63D23", 
                               "> 5 - < 6" = "#BD251C", 
                               ">= 6" = "#8C130E")) +
  labs(fill = "Life years lost  ", title = expression("Potential Gains in Life Expectancy through Permanently Reducing" ~ PM[2.5] ~ "from 2021 concentrations to the WHO guideline"), 
       subtitle = "") + 
 theme(legend.position = "bottom", 
        legend.justification = c(0.5, 3), 
        legend.background = element_rect(color = "black"), 
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15), 
        plot.title = element_text(hjust = 0.5, size = 15), 
       plot.subtitle = element_text(hjust = 0.5, size = 9, margin = margin(b = 0.5, unit = "cm")),
       # legend.key = element_rect(color = "black"),
       legend.box.margin = margin(b = 1, unit = "cm"),
        plot.caption = element_text(hjust = 0.7, size = 9, face = "italic"), 
       legend.key = element_rect(color = "black"), 
       legend.box.spacing = unit(0, "cm"), 
        legend.direction = "horizontal") +
  guides(fill = guide_legend(nrow = 1)) 
   
   
#> Fig 1.2: individual state wise maps

state <- "Jammu and Kashmir"
two_states <- c("Jammu and Kashmir", "Azad Kashmir")

india_fs_fig1.2 <- india_fs_fig1.1_data %>%
  filter(name_1 %in% two_states) %>%
  ggplot() +
  geom_sf(mapping = aes(fill = forcats::fct_reorder(lyl_bucket, order_lyl_bucket)), color = "black", lwd = 0.1) +
      #geom_sf(data = india_fs_fig1.1_data %>% filter(name_2 == "Chennai"), color = "black", lwd = 1, fill = "transparent") +
     #ggsflabel::geom_sf_label_repel(data = india_fs_fig1.1_data %>% filter(name_2 == "Chennai"), mapping = aes(label = name_2), force = 100, nudge_x = -1, nudge_y = 0.5, seed = 10, size = 2.8) +
  #geom_sf(data = india_state, color = "lightgrey", fill = "transparent", lwd = 0.1) +
  #geom_sf(data = india_state %>% filter(NAME_1 %in% state), color = "black", fill = "transparent", lwd = 0.8) +
   ggthemes::theme_map() +
    #ggsflabel::geom_sf_label_repel(data = india_state %>% filter(NAME_1 %in% states_to_plot), mapping = aes(label = NAME_1), force = 50, nudge_x = -10, nudge_y = -1.6, seed = 10, size = 4) +
  ggthemes::theme_map() + 
 scale_fill_manual(values = c("0 - < 0.1 years" = "#FFFFFF", 
                               "0.1 - 0.5" = "#FFE6B3", 
                               "> 0.5 - 1" = "#FFD25D", 
                               "> 1 - 2" = "#FFBA00", 
                               "> 2 - 3" = "#FF9600", 
                               "> 3 - 4" = "#FF6908", 
                               "> 4 - 5" = "#E63D23", 
                               "> 5 - < 6" = "#BD251C", 
                               ">= 6" = "#8C130E")) +
  labs(fill = "Life years lost  ", title = expression("Potential Gains in Life Expectancy through Permanently Reducing" ~ PM[2.5] ~ "from 2021 concentrations to the WHO guideline"), 
       subtitle = str_c("(", state, ")")) + 
 theme(legend.position = "bottom", 
        legend.justification = c(0.5, 3), 
        legend.background = element_rect(color = "black"), 
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15), 
        plot.title = element_text(hjust = 0.5, size = 15, margin = margin(b = 0.3, unit = "cm")), 
       plot.subtitle = element_text(hjust = 0.5, size = 14, margin = margin(b = 0.8, unit = "cm")),
       # legend.key = element_rect(color = "black"),
       legend.box.margin = margin(b = 1, unit = "cm"),
        plot.caption = element_text(hjust = 0.7, size = 9, face = "italic"), 
       legend.key = element_rect(color = "black"), 
       legend.box.spacing = unit(0, "cm"), 
        legend.direction = "horizontal") +
  guides(fill = guide_legend(nrow = 1)) 


#> Fig 2: Potential gain in life expectancy in top 10 most populous states in India----------

india_fs_fig2_data <- gadm_level_summary(gadm2_aqli_2021, c("country", "name_1"), c(2021),  10) %>%
  filter(country == "India") %>% 
  slice_max(population, n = 10) %>%
  mutate(lyl_bucket = ifelse((llpp_who_2021 >= 0) & (llpp_who_2021 < 0.1), "0 - < 0.1 years", NA), 
         lyl_bucket = ifelse((llpp_who_2021 >= 0.1) & (llpp_who_2021 <= 0.5), "0.1 - 0.5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 0.5) & (llpp_who_2021 <= 1), "> 0.5 - 1", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 1) & (llpp_who_2021 <= 2), "> 1 - 2", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 2) & (llpp_who_2021 <= 3), "> 2 - 3", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 3) & (llpp_who_2021 <= 4), "> 3 - 4", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 4) & (llpp_who_2021 <= 5), "> 4 - 5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 5) & (llpp_who_2021 < 6), "> 5 - < 6", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 >= 6), ">= 6 years", lyl_bucket)) %>%
  mutate(order_lyl_bucket = ifelse(lyl_bucket == "0 - < 0.1 years", 1, NA), 
         order_lyl_bucket = ifelse(lyl_bucket == "0.1 - 0.5", 2, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 0.5 - 1", 3, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 1 - 2", 4, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 2 - 3", 5, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 3 - 4", 6, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 4 - 5", 7, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 5 - < 6", 8, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == ">= 6 years", 9, order_lyl_bucket))

india_fs_fig2_data$lyl_bucket <- as.factor(india_fs_fig2_data$lyl_bucket)

india_fs_fig2_plt <- india_fs_fig2_data %>%
  ggplot() +
   geom_col(mapping = aes(x = forcats::fct_reorder(name_1, llpp_who_2021), y = llpp_who_2021, fill = lyl_bucket), width = 0.5) +
  labs(x = "State", y = "Average Life Expectancy Gains (Years)",
       title = "State wise Average Life Expectancy Gains",
       subtitle = expression("(Reducing" ~ PM[2.5] ~ "pollution from 2021 levels to the WHO guideline in ten most populous states of India)"), 
       caption = "* Top 10 most populated states are ordered from most populated (Uttar Pradesh) to least populated (Tamil Nadu).") +
  scale_y_continuous(breaks = seq(0, 9, 1), limits = c(0, 9)) +
    scale_fill_manual(values = c("0 - < 0.1 years" = "#FFFFFF", 
                               "0.1 - 0.5" = "#FFE6B3", 
                               "> 0.5 - 1" = "#FFD25D", 
                               "> 1 - 2" = "#FFBA00", 
                               "> 2 - 3" = "#FF9600", 
                               "> 3 - 4" = "#FF6908", 
                               "> 4 - 5" = "#E63D23", 
                               "> 5 - < 6" = "#BD251C", 
                               ">= 6 years" = "#8C130E")) +
  coord_flip() +
  ggthemes::theme_tufte() +
  theme(legend.position = "none", 
        plot.title = element_text(hjust = 0.5, size = 16), 
        plot.subtitle = element_text(hjust =  0.5, size = 10, face = "italic", margin = margin(b = 0.8, unit = "cm")), 
        plot.caption = element_text(size = 8, hjust = 0, face = "italic"),
        axis.line = element_line(), 
        axis.text = element_text(size = 11), 
        axis.title = element_text(size = 13), 
        axis.title.y = element_text(margin = margin(r = 0.7, unit = "cm")), 
        axis.title.x = element_text(margin = margin(t = 0.6, b = 0.6, unit = "cm")), 
        axis.ticks = element_blank()) 


svglite("india_fs_fig2")
ggsave("./september.2023/factsheets/india/india_fs_fig2.svg", width = 12, height = 9)

india_fs_fig2_data %>%
  write_csv("./september.2023/factsheets/india/fig2/india_fs_fig2_data.csv")


#> GBD India graph--------------------------------------------

# top 5 most deadly diseases
india_fs_fig3_dataset <- gbd_results_master_2021 %>%
  filter(country == "India") %>%
  slice_max(lyl, n = 5)

colnames(india_fs_fig3_dataset)[3] <- c("llpp_who_2021")

india_fs_fig3_dataset <- india_fs_fig3_dataset %>%
    mutate(lyl_bucket = ifelse((llpp_who_2021 >= 0) & (llpp_who_2021 < 0.1), "0 - < 0.1 years", NA), 
         lyl_bucket = ifelse((llpp_who_2021 >= 0.1) & (llpp_who_2021 <= 0.5), "0.1 - 0.5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 0.5) & (llpp_who_2021 <= 1), "> 0.5 - 1", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 1) & (llpp_who_2021 <= 2), "> 1 - 2", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 2) & (llpp_who_2021 <= 3), "> 2 - 3", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 3) & (llpp_who_2021 <= 4), "> 3 - 4", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 4) & (llpp_who_2021 <= 5), "> 4 - 5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 5) & (llpp_who_2021 < 6), "> 5 - < 6", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 >= 6), ">= 6", lyl_bucket)) %>%
  mutate(order_lyl_bucket = ifelse(lyl_bucket == "0 - < 0.1 years", 1, NA), 
         order_lyl_bucket = ifelse(lyl_bucket == "0.1 - 0.5", 2, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 0.5 - 1", 3, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 1 - 2", 4, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 2 - 3", 5, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 3 - 4", 6, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 4 - 5", 7, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 5 - 6", 8, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == ">= 6", 9, order_lyl_bucket))


# getting disease (ordered from most deadly to least deadly) in a vector (to be used later in the
# scale_x_discrete function).

# cause_of_death_ordered <- india_fs_fig3_dataset %>%
#   arrange(llpp_who_2021) %>%
#   pull(cause_of_death) %>%
#   unlist()





#> figure 3 plot

# figure 3 caption removed: Source: Global Burden of Disease (https://vizhub.healthdata.org/gbd-results/) causes and risks data and WHO Life Tables (https://apps.who.int/gho/data/node.main.LIFECOUNTRY?lang=en) were used combined with the Life table method to arrive at these results. 'PM2.5 relative to WHO Guideline' bar displays life years lost relative to the WHO guideline as calculated by latest AQLI (2021) data, which will be published in September 2023.

india_fs_fig3 <- india_fs_fig3_dataset %>%
  ggplot() + 
  geom_col(mapping = aes(x = forcats::fct_reorder(cause_of_death, llpp_who_2021), y = llpp_who_2021, fill = forcats::fct_reorder(lyl_bucket, order_lyl_bucket)), width = 0.4, color = "black") +
  labs(x = "Cause of Death", y = "Life years lost", fill = "Life years lost", 
       title = expression("Top 5 causes of reduction in life expectancy in India")) +
  coord_flip() + 
  ggthemes::theme_tufte() +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 14),
        axis.title.y = element_text(size = 16, margin = margin(r = 0.6, unit = "cm")),
        axis.title.x = element_text(size = 16, margin = margin(t = 0.6, b = 0.6, unit = "cm")),
        plot.caption = element_text(hjust = 0, size = 8, margin = margin(t = 0.8, unit = "cm")),
        plot.title = element_text(hjust = 0.5, size = 20, margin = margin(b = 0.8, unit = "cm")),
        plot.subtitle = element_text(hjust = 0.5, size = 10, margin = margin(b = 0.8, unit = "cm")),
        legend.box.background = element_rect(color = "black"),
        plot.background = element_rect(color = "white"),
        axis.line = element_line(), 
        legend.text = element_text(size = 11), 
        legend.title = element_text(size = 14),
        panel.grid.major.y = element_blank()) +
  scale_y_continuous(breaks = seq(0, 6, 0.5)) +
 # scale_x_discrete(limits = cause_of_death_ordered[seq(1, length(cause_of_death_ordered), by = 2)]) +
  scale_fill_manual(values = c("0 - < 0.1 years" = "#FFFFFF", 
                               "0.1 - 0.5" = "#FFE6B3", 
                               "> 0.5 - 1" = "#FFD25D", 
                               "> 1 - 2" = "#FFBA00", 
                               "> 2 - 3" = "#FF9600", 
                               "> 3 - 4" = "#FF6908", 
                               "> 4 - 5" = "#E63D23", 
                               "> 5 - < 6" = "#BD251C", 
                               ">= 6" = "#8C130E")) + 
  guides(fill = guide_legend(nrow = 1)) +
  themes_aqli_base

svglite("india_fs_fig3")
ggsave("./september.2023/factsheets/india/fig3/india_fs_fig3.svg", width = 15, height = 12)

india_fs_fig3_dataset %>%
   write_csv("./september.2023/factsheets/india/fig3/india_fs_fig3_data.csv")


#> Figure 4: Average PM2.5 concentration in India from 1998 to 2021------------------------

# adding a north India/rest of india region column (specifically Indo gangetic plain)
north_india <- c("West Bengal", "Uttar Pradesh", "Punjab", "Haryana", 
                 "Chandigarh", "Bihar", "NCT of Delhi")

india_fs_fig4_data_part1 <- gadm2_aqli_2021 %>%
  filter(country == "India") %>%
  mutate(region = ifelse(name_1 %in% north_india, "North India", "All other regions (excluding North India)"))


# creating India's region wise average PM2.5 data from 1998 to 2021 
india_fs_fig4_data_part1 <- india_fs_fig4_data_part1 %>%
  group_by(region) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         mutate(across(starts_with("pm"), ~.x*pop_weights, .names = "{col}_weighted"))) %>%
  summarise(across(ends_with("weighted"), sum)) %>%
  pivot_longer(cols = pm1998_weighted:pm2021_weighted, names_to = "years", 
               values_to = "pop_weighted_avg_pm2.5") %>%
  mutate(years = as.integer(unlist(str_extract(years, "\\d+")))) %>%
  select(years, region, pop_weighted_avg_pm2.5)

# creating a national average PM2.5 trendlines data from 1998 to 2021
india_fs_fig4_data_part2 <- gadm2_aqli_2021 %>%
  filter(country == "India") %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         mutate(across(starts_with("pm"), ~.x*pop_weights, .names = "{col}_weighted"))) %>%
  summarise(across(ends_with("weighted"), sum)) %>%
  pivot_longer(cols = pm1998_weighted:pm2021_weighted, names_to = "years", 
               values_to = "pop_weighted_avg_pm2.5") %>%
  mutate(years = as.integer(unlist(str_extract(years, "\\d+"))), 
         region = "National Average") %>% 
  select(years, region, pop_weighted_avg_pm2.5)

# India factsheet figure 4 dataset
india_fs_fig4_data <- rbind(india_fs_fig4_data_part1, india_fs_fig4_data_part2)

india_fs_fig4_data %>%
   write_csv("./september.2023/factsheets/india/fig4/india_fs_fig4_data.csv")

# India factsheet figure 4
# fig4 caption removed: "* We define the Indo-Gangetic plain region as containing the following seven states and union territories: Bihar, Chandigarh, NCT of Delhi, Haryana, Punjab, Uttar Pradesh, and West Bengal."

india_fs_fig4_data$region <- factor(india_fs_fig4_data$region, levels = c("North India", "National Average", "All other regions (excluding North India)"))

india_fs_fig4 <- ggplot(india_fs_fig4_data) +
  geom_line(mapping = ggplot2::aes(x = years, y = pop_weighted_avg_pm2.5, color = region), lwd = 1.3) +
    geom_hline(mapping = aes(yintercept = 5), lwd = 0.8, linetype = "dotted") +
  scale_y_continuous(breaks = seq(0, 100, 10), limits = c(0, 100)) +
  scale_x_continuous(breaks = c(seq(1998, 2019, 3), 2021)) +
    scale_color_manual(values = c("North India" = "darkred", "National Average" = "darkgrey", "All other regions (excluding North India)" = "orange")) +
  ggthemes::theme_tufte() +
  labs(x = "Year", 
       y = expression("Annual Average   " ~ PM[2.5] ~ "(in µg/m³)"), 
       title = expression("Annual Average" ~ PM[2.5] ~ "concentrations in India, 1998 to 2021")) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 11),
        legend.title = element_blank(),
        axis.title.y = element_text(size = 13, margin = margin(r = 0.6, unit = "cm")), 
        axis.title.x = element_text(size = 13, margin = margin(t = 0.6, b = 0.6, unit = "cm")), 
        axis.line = element_line(), 
        legend.box.background = element_rect(color = "black"),
        plot.title = element_text(hjust = 0.5, size = 16, margin = margin(b = 0.7, unit = "cm")), 
        plot.subtitle = element_text(hjust = 0.5, size = 8, face = "italic"), 
        plot.caption = element_text(size = 7, margin = margin(t = 0.8, unit = "cm"), hjust = 0, face = "italic"),
        axis.text = element_text(size = 12), 
        plot.background = element_rect(color = "white"), 
        axis.ticks = element_blank()) +
     geom_text(x = 2000.5, y = 7.9, label = expression("WHO" ~ PM[2.5] ~ "Guideline (last updated: 2021): 5 µg/m³"), size = 4.5) 


svglite("india_fs_fig4")
ggsave("./september.2023/factsheets/india/india_fs_fig4.svg", width = 12, height = 9)


# Appendix table

appendix_table_raw_india <- gadm_level_summary(gadm2_aqli_2021, c("country", "name_1"), c(2021), 10) %>%
filter(country == "India") %>%
   mutate(population_lakhs = round(population/100000, 1)) %>%
  select(name_1, population_lakhs, pm2021, llpp_who_2021, llpp_nat_2021) %>%
  rename("State/UT" = name_1, 
         "Population (Lakhs)" = population_lakhs, 
         "Annual Average 2021 PM2.5 Concentration (µg/m³)" = pm2021, 
         "Life Expectancy Gains from reducing PM2.5 from 2021 concentration to WHO PM2.5 guideline of 5 µg/m³" = llpp_who_2021, 
         "Life Expectancy Gains from reducing PM2.5 from 2021 concentration to National PM2.5 guideline of 40 µg/m³" = llpp_nat_2021)

# Experimental figure



```


# South Korea factsheet

```{r}

#> quick stats

# South Korea ranking amongst other countries in terms of pollution
gadm_level_summary(gadm2_aqli_2021, c("country"), c(2021), 10) %>%
  slice_max(pm2021, n = 50) %>%
  View()

# most populated region
gadm_level_summary(gadm2_aqli_2021, c("country", "name_1"), c(2021), 10) %>%
  filter(country == "South Korea") %>%
  slice_max(population, n = 10)

#> plot 1: map screenshot for South Korea

# south korea figure1 dataset
south_korea_fs_fig1_data <- gadm2_aqli_2021 %>%
  filter(country == "South Korea") %>%
  left_join(gadm2_aqli_2021_shp, by = c("objectid_gadm2" = "obidgadm2")) %>%
  mutate(lyl_bucket = ifelse((llpp_who_2021 >= 0) & (llpp_who_2021 < 0.1), "0 - < 0.1", NA), 
         lyl_bucket = ifelse((llpp_who_2021 >= 0.1) & (llpp_who_2021 <= 0.5), "0.1 - 0.5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 0.5) & (llpp_who_2021 <= 1), "> 0.5 - 1", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 1) & (llpp_who_2021 <= 2), "> 1 - 2", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 2) & (llpp_who_2021 <= 3), "> 2 - 3", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 3) & (llpp_who_2021 <= 4), "> 3 - 4", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 4) & (llpp_who_2021 <= 5), "> 4 - 5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 5) & (llpp_who_2021 < 6), "> 5 - < 6", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 >= 6), ">= 6", lyl_bucket)) %>%
  mutate(order_lyl_bucket = ifelse(lyl_bucket == "0 - < 0.1", 1, NA), 
         order_lyl_bucket = ifelse(lyl_bucket == "0.1 - 0.5", 2, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 0.5 - 1", 3, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 1 - 2", 4, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 2 - 3", 5, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 3 - 4", 6, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 4 - 5", 7, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 5 - < 6", 8, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == ">= 6", 9, order_lyl_bucket)) %>%
  select(-geometry, geometry) %>%
  st_as_sf()

south_korea_fs_fig1_data %>%
  sf::write_sf("./september.2023/factsheets/southkorea/fig1/sk_fs_fig1.shp")


# south korea figure 1
south_korea_fs_fig1 <- south_korea_fs_fig1_data %>%
  ggplot() +
  geom_sf(mapping = aes(fill = forcats::fct_reorder(lyl_bucket, order_lyl_bucket)), color = "lightgrey", lwd = 0.05) +
  geom_sf(data = gadm1_aqli_2021_shp %>% filter(name0 == "South Korea"), color = "black", fill = "transparent", lwd = 0.5) +
  ggthemes::theme_map() + 
 scale_fill_manual(values = c("0 - < 0.1 years" = "#FFFFFF", 
                               "0.1 - 0.5" = "#FFE6B3", 
                               "> 0.5 - 1" = "#FFD25D", 
                               "> 1 - 2" = "#FFBA00", 
                               "> 2 - 3" = "#FF9600", 
                               "> 3 - 4" = "#FF6908", 
                               "> 4 - 5" = "#E63D23", 
                               "> 5 - < 6" = "#BD251C", 
                               ">= 6" = "#8C130E")) +
  ggthemes::theme_map() +
  labs(fill = "Gain in years of life expectancy ", title = expression("Potential Gains in Life Expectancy (South Korea) through Permanently Reducing" ~ PM[2.5] ~ "from 2021 concentrations to the WHO guideline")) + 
 theme(legend.position = "bottom", 
        legend.justification = c(0.5, 3), 
        legend.background = element_rect(color = "black"), 
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15), 
        plot.title = element_text(hjust = 0.5, size = 15), 
       # legend.key = element_rect(color = "black"),
       legend.box.margin = margin(b = 1, unit = "cm"),
        plot.subtitle = element_text(hjust = 0.5, size = 7), 
        plot.caption = element_text(hjust = 0.7, size = 9, face = "italic"), 
       legend.key = element_rect(color = "black"), 
       legend.box.spacing = unit(0, "cm"), 
        legend.direction = "horizontal", 
       plot.background = element_rect(fill = "white", color = "white")) +
  guides(fill = guide_legend(nrow = 1))

ggsave("./september.2023/factsheets/southkorea/sk_fs_fig1.svg", width = 15, height = 10)

#> plot 2: Gain in LE in reducing 2021 pollution to WHO guideline in top 10 most populous regions

# fig2 dataset
southkorea_fs_fig2_dataset <- gadm2_aqli_2021 %>%
  filter(country == "South Korea") %>%
  add_aqli_color_scale_buckets(scale_type = "lyl", col_name = "llpp_who_2021") %>%
  slice_max(population, n = 10)

southkorea_fs_fig2_dataset %>%
   write_csv("./september.2023/factsheets/southkorea/fig2/sk_fs_fig2_data.csv")


# fig 2

south_korea_fs_fig2 <- southkorea_fs_fig2_dataset %>%
  aqli_bar(scale_type = "lyl", x_var = "name_2", y_var = "llpp_who_2021", title = expression("Potential Gain in Life Expectancy from Reducing 2021 " ~ PM[2.5] ~ "concentrations to the WHO Guideline"), subtitle = "(Ten most populous regions of South Korea)", x_label = "Region", y_label = "Average Life Expectancy Gains (Years)", legend_title = "Life Expectancy Gains (Years)", caption = "")
 
# plot 3: Trendlines: 1998 to 2021

# southkorea factsheet figure 3 dataset
southkorea_fs_fig3_dataset <- gadm2_aqli_2021 %>%
  filter(country == "South Korea") %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         mutate(across(starts_with("pm"), ~.x*pop_weights, .names = "{col}_weighted"))) %>%
  summarise(across(ends_with("weighted"), sum)) %>%
  pivot_longer(cols = pm1998_weighted:pm2021_weighted, names_to = "years", 
               values_to = "pop_weighted_avg_pm2.5") %>%
  mutate(years = as.integer(unlist(str_extract(years, "\\d+"))), 
         region = "National Average") %>% 
  select(years, region, pop_weighted_avg_pm2.5)


southkorea_fs_fig3_dataset %>%
   write_csv("./september.2023/factsheets/southkorea/fig3/sk_fs_fig3_data.csv")


  southkorea_fs_fig3 <- southkorea_fs_fig3_dataset %>%
    ggplot() +
  geom_line(mapping = ggplot2::aes(x = years, y = pop_weighted_avg_pm2.5), lwd = 1.1, 
            color = "darkred") +
    geom_hline(mapping = aes(yintercept = 5), lwd = 0.8, linetype = "dashed") +
    geom_hline(mapping = aes(yintercept = 15), lwd = 0.8, linetype = "dashed") + 
  scale_y_continuous(breaks = seq(0, 35, 5), limits = c(0, 35)) +
  scale_x_continuous(breaks = c(seq(1998, 2019, 3), 2021)) +
    scale_color_manual(values = c("National Average" = "darkgrey")) +
  ggthemes::theme_clean() +
  labs(x = "Year", 
       y = expression("Annual Average  " ~ PM[2.5] ~ "concentrations (in µg/m³)"), 
       title = expression("Annual Average" ~ PM[2.5] ~ " concentrations in South Korea, 1998 to 2021")) +
  theme(legend.position = "bottom", legend.title = element_blank(), 
        legend.text = element_text(size = 7), 
        axis.title.y = element_text(size = 9), 
        axis.title.x = element_text(size = 9)) +  
    geom_text(x = 2001.8, y = 6.1, label =  expression("WHO " ~ PM[2.5] ~ "Guideline (last updated in 2021): 5 µg/m³"), size = 5) +
    geom_text(x = 2001.8, y = 16.1, label = expression("South Korea National" ~ PM[2.5] ~ "Guideline: 15 µg/m³"), size = 5) +
    themes_aqli_base
  
  
  ggsave("./september.2023/factsheets/southkorea/fig3/sk_fs_fig3.svg", width = 15, height = 10)



  # appendix table (gadm level 2)
  gadm2_aqli_2021 %>%
    filter(country == "South Korea") %>%
    select(name_2, population, pm2021, llpp_who_2021, llpp_nat_2021) %>%
    mutate(pop_millions = round(population/1000000, 1), 
           pm2021 = round(pm2021, 1), 
           llpp_who_2021 = round(llpp_who_2021, 1), 
           llpp_nat_2021 = round(llpp_nat_2021, 1)) %>%
    slice_max(population, n = 25) %>% 
    select(name_2, pop_millions, pm2021, llpp_who_2021, llpp_nat_2021) %>%
    rename(`Population (Millions)` = pop_millions, 
           "Region" = name_2, 
           "PM2.5 concentration 2021 (in µg/m³)" = pm2021, 
           "Life Expectancy Gains from reducing PM2.5 from 2021 Concentrations to the National Guideline of 15 µg/m³" = llpp_nat_2021, 
           "Life Expectancy Gains from reducing PM2.5 from 2021 Concentrations to the WHO Guideline of 5 µg/m³" = llpp_who_2021) %>% 
    write_csv("./[updatedOnJune072023]South Korea Factsheet 2023 appendix table.csv")
    

```

# China factsheet

```{r}

# sanity checks

# total life years lost for regions out of compliance with China's national standard
tmp_data <- gadm_level_summary(gadm2_aqli_2021, c("country", "name_1", "name_2"), c(2021), 10) %>%
  filter(country == "China") %>%
  filter(pm2021 > 5) 

tmp_data %>%
  gadm_level_summary(c("country"), c(2021), 10) %>%
  View()

#> plot 1: map screenshot--------------------

# china figure1 dataset
china_fs_fig1_data <- gadm2_aqli_2021 %>%
  filter(country == "China") %>%
  left_join(gadm2_aqli_2021_shp, by = c("objectid_gadm2" = "obidgadm2")) %>%
  mutate(lyl_bucket = ifelse((llpp_who_2021 >= 0) & (llpp_who_2021 < 0.1), "0 - < 0.1", NA), 
         lyl_bucket = ifelse((llpp_who_2021 >= 0.1) & (llpp_who_2021 <= 0.5), "0.1 - 0.5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 0.5) & (llpp_who_2021 <= 1), "> 0.5 - 1", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 1) & (llpp_who_2021 <= 2), "> 1 - 2", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 2) & (llpp_who_2021 <= 3), "> 2 - 3", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 3) & (llpp_who_2021 <= 4), "> 3 - 4", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 4) & (llpp_who_2021 <= 5), "> 4 - 5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 5) & (llpp_who_2021 < 6), "> 5 - < 6", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 >= 6), ">= 6", lyl_bucket)) %>%
  mutate(order_lyl_bucket = ifelse(lyl_bucket == "0 - < 0.1", 1, NA), 
         order_lyl_bucket = ifelse(lyl_bucket == "0.1 - 0.5", 2, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 0.5 - 1", 3, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 1 - 2", 4, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 2 - 3", 5, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 3 - 4", 6, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 4 - 5", 7, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 5 - < 6", 8, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == ">= 6", 9, order_lyl_bucket)) %>%
  select(-geometry, geometry) %>%
  st_as_sf()

# china_fs_fig1_data %>%
#   sf::write_sf("./september.2023/factsheets/southkorea/fig1/sk_fs_fig1.shp")


# china figure 1
china_fs_fig1 <- china_fs_fig1_data %>%
  ggplot() +
  geom_sf(mapping = aes(fill = forcats::fct_reorder(lyl_bucket, order_lyl_bucket)), color = "lightgrey", lwd = 0.05) +
 geom_sf(data = gadm1_aqli_2021_shp %>% filter(name0 == "China"), color = "black", fill = "transparent", lwd = 0.5) +
  ggthemes::theme_map() + 
 scale_fill_manual(values = c("0 - < 0.1" = "#FFFFFF", 
                               "0.1 - 0.5" = "#FFE6B3", 
                               "> 0.5 - 1" = "#FFD25D", 
                               "> 1 - 2" = "#FFBA00", 
                               "> 2 - 3" = "#FF9600", 
                               "> 3 - 4" = "#FF6908", 
                               "> 4 - 5" = "#E63D23", 
                               "> 5 - < 6" = "#BD251C", 
                               ">= 6" = "#8C130E")) +
  ggthemes::theme_map() +
  labs(fill = "Gain in years of life expectancy  ", title = expression("Potential Gains in Life Expectancy (China) through Permanently Reducing" ~ PM[2.5] ~ "from 2021 concentrations to the WHO guideline")) + 
 theme(legend.position = "bottom", 
        legend.justification = c(0.5, 3), 
        legend.background = element_rect(color = "black"), 
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15), 
        plot.title = element_text(hjust = 0.5, size = 15), 
       # legend.key = element_rect(color = "black"),
       legend.box.margin = margin(b = 1, unit = "cm"),
        plot.subtitle = element_text(hjust = 0.5, size = 7), 
        plot.caption = element_text(hjust = 0.7, size = 9, face = "italic"), 
       legend.key = element_rect(color = "black"), 
       legend.box.spacing = unit(0, "cm"), 
        legend.direction = "horizontal", 
       plot.background = element_rect(fill = "white", color = "white")) +
  guides(fill = guide_legend(nrow = 1))






#> plot 2: Potential Gain in Life Expectancy from Reducing PM2.5 from 2021 to the WHO Guideline in 10 most populous regions of China -----------------------------------------

china_fs_fig2_dataset <- gadm2_aqli_2021 %>%
  filter(country == "China") %>%
  add_aqli_color_scale_buckets(scale_type = "lyl", col_name = "llpp_who_2021") %>%
  slice_max(population, n = 10)

china_fs_fig2 <- china_fs_fig2_dataset %>%
  aqli_bar(scale_type = "lyl", x_var = "name_2", y_var = "llpp_who_2021", title = expression("Potential Gain in Life Expectancy from Reducing 2021 " ~ PM[2.5] ~ "concentrations to the WHO Guideline"), subtitle = "(Ten most populous prefectures of China)", x_label = "Prefecture", y_label = "Average Life Expectancy Gains (Years)", legend_title = "LE gains (Years)", caption = "")


#> plot 3: PM2.5 Concentrations in Major Regions in Mainland China Over Time (1998 to 2021)-----------------

# filtering the color file for China
color_2021_china <- gadm2_aqli_2021 %>%
  filter(country == "China")

# BTH region (name_1)
bth_region <- c("Beijing", "Tianjin", "Hebei")

# YRD (name_1)
yrd_region <- c("Shanghai", "Jiangsu", "Zhejiang")

# PRD (name_1, name_2)

prd_region_name_2_guandong_regions <- c("Dongguan", "Foshan", "Guangzhou", 
                                        "Huizhou", "Jiangmen", "Shenzhen", 
                                        "Zhaoqing", "Zhongshan", "Zhuhai")

prd_region_name_1 <- c("Hong Kong", "Macau")

prd <- c(prd_region_name_2_guandong_regions, prd_region_name_1)


# add region column
color_2021_china  <- color_2021_china %>%
  mutate(region = ifelse(name_1 %in% bth_region, "BTH", "others"), 
         region = ifelse(name_1 %in% yrd_region,  "YRD", region), 
         region = ifelse(name_1 %in% prd_region_name_1, "PRD", region), 
         region = ifelse(name_2 %in% prd_region_name_2_guandong_regions, "PRD", region)) 


# China factsheet figure 3 (part-1): trend lines figure region wise (BTH, YRD, PRD) dataset
trendlines_china_region_wise_df <- color_2021_china %>%
  group_by(region) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         mutate(across(starts_with("pm"), ~.x*pop_weights, .names = "{col}_weighted"))) %>%
  summarise(across(ends_with("weighted"), sum)) %>%
  pivot_longer(cols = pm1998_weighted:pm2021_weighted, names_to = "years", 
               values_to = "pop_weighted_avg_pm2.5") %>%
  mutate(years = as.integer(unlist(str_extract(years, "\\d+")))) %>%
  select(years, region, pop_weighted_avg_pm2.5) %>%
  filter(region %in% c("YRD", "BTH", "PRD"))

# China factsheet figure 3 (part-2): trend line national average dataset
trendline_national_avg_df <- color_2021_china %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         mutate(across(starts_with("pm"), ~.x*pop_weights, .names = "{col}_weighted"))) %>%
  summarise(across(ends_with("weighted"), sum)) %>%
  pivot_longer(cols = pm1998_weighted:pm2021_weighted, names_to = "years", 
               values_to = "pop_weighted_avg_pm2.5") %>%
  mutate(years = as.integer(unlist(str_extract(years, "\\d+"))), 
         region = "China") %>% 
  select(years, region, pop_weighted_avg_pm2.5)

# China factsheet figure 3 dataset final
china_fs_fig3_dataset <- rbind(trendlines_china_region_wise_df, trendline_national_avg_df)

china_fs_fig3_dataset <-  china_fs_fig3_dataset %>%
    mutate(region_order = ifelse(region == "China", 1, NA), 
         region_order = ifelse(region == "PRD", 2, region_order), 
         region_order = ifelse(region == "YRD", 3, region_order), 
         region_order = ifelse(region == "BTH", 4, NA))

# save China factsheet figure 3 dataset
# china_fs_fig3_dataset %>%
#   write_csv("./june.2022/factsheets/china/graph.wise.datasets/china_fs_fig3.csv")

china_fs_fig3_dataset$region <- factor(china_fs_fig3_dataset$region, levels = c("BTH", "China", "YRD", "PRD"))

# plot china factsheet figure 3
china_fs_fig3 <- china_fs_fig3_dataset %>%
  ggplot() +
  geom_line(mapping = ggplot2::aes(x = years, y = pop_weighted_avg_pm2.5, color = forcats::fct_reorder(region, region_order)), lwd = 1.1) +
    geom_hline(mapping = aes(yintercept = 5), lwd = 0.5, linetype = "dashed") +
    geom_vline(mapping = aes(xintercept = 2014), lwd = 0.5, linetype = "dashed") +
  scale_y_continuous(breaks = seq(0, 100, 10), limits = c(0, 100)) +
  scale_x_continuous(breaks = c(seq(1998, 2019, 3), 2021)) +
    scale_color_manual(values = c("PRD" = "burlywood1", "China" = "darkgrey", "BTH" = "darkred", "YRD" = "darkorange")) +
  ggthemes::theme_clean() +
  labs(x = "Year", 
       y = expression("Annual average  " ~ PM[2.5] ~ " concentrations (in µg/m³)"), 
       title = expression("Annual average" ~ PM[2.5] ~ "Concentrations in Major Regions in Mainland China Over Time"), 
    subtitle = "(1998 to 2021)", 
    color = "Region") +
  theme(legend.position = "bottom", legend.title = element_blank(), 
        legend.text = element_text(size = 6.5), 
        axis.title.y = element_text(size = 9), 
        axis.title.x = element_text(size = 9)) + 
    geom_text(x = 2001, y = 8.2, label = expression("WHO" ~ PM[2.5] ~ "Guideline (last updated: 2021): 5 µg/m³")) + 
    geom_text(x = 2015.3, y = 93, label = str_wrap("China announces war on pollution", width = 18)) +
  themes_aqli_base
  





#> plot 4: GBD----------------------------------------- 

# top 5 most deadly diseases
china_fs_fig4_dataset <- gbd_results_master_2021 %>%
  filter(country == "China") %>%
  slice_max(lyl, n = 5)

colnames(china_fs_fig4_dataset)[3] <- c("llpp_who_2021")

china_fs_fig4_dataset <- china_fs_fig4_dataset %>%
    mutate(lyl_bucket = ifelse((llpp_who_2021 >= 0) & (llpp_who_2021 < 0.1), "0 - < 0.1 years", NA), 
         lyl_bucket = ifelse((llpp_who_2021 >= 0.1) & (llpp_who_2021 <= 0.5), "0.1 - 0.5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 0.5) & (llpp_who_2021 <= 1), "> 0.5 - 1", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 1) & (llpp_who_2021 <= 2), "> 1 - 2", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 2) & (llpp_who_2021 <= 3), "> 2 - 3", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 3) & (llpp_who_2021 <= 4), "> 3 - 4", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 4) & (llpp_who_2021 <= 5), "> 4 - 5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 5) & (llpp_who_2021 < 6), "> 5 - < 6", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 >= 6), ">= 6", lyl_bucket)) %>%
  mutate(order_lyl_bucket = ifelse(lyl_bucket == "0 - < 0.1 years", 1, NA), 
         order_lyl_bucket = ifelse(lyl_bucket == "0.1 - 0.5", 2, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 0.5 - 1", 3, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 1 - 2", 4, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 2 - 3", 5, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 3 - 4", 6, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 4 - 5", 7, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 5 - 6", 8, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == ">= 6", 9, order_lyl_bucket))


# getting disease (ordered from most deadly to least deadly) in a vector (to be used later in the
# scale_x_discrete function).

# cause_of_death_ordered <- india_fs_fig3_dataset %>%
#   arrange(llpp_who_2021) %>%
#   pull(cause_of_death) %>%
#   unlist()





#> figure 3 plot

# figure 3 caption removed: Source: Global Burden of Disease (https://vizhub.healthdata.org/gbd-results/) causes and risks data and WHO Life Tables (https://apps.who.int/gho/data/node.main.LIFECOUNTRY?lang=en) were used combined with the Life table method to arrive at these results. 'PM2.5 relative to WHO Guideline' bar displays life years lost relative to the WHO guideline as calculated by latest AQLI (2021) data, which will be published in September 2023.

china_fs_fig4 <- china_fs_fig4_dataset %>%
  ggplot() + 
  geom_col(mapping = aes(x = forcats::fct_reorder(cause_of_death, llpp_who_2021), y = llpp_who_2021, fill = forcats::fct_reorder(lyl_bucket, order_lyl_bucket)), width = 0.4, color = "black") +
  labs(x = "Cause of Death", y = "Life years lost", fill = "Life years lost", 
       title = "Top 5 causes of reduction in life expectancy in China") +
  coord_flip() + 
  ggthemes::theme_tufte() +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 14),
        axis.title.y = element_text(size = 16, margin = margin(r = 0.6, unit = "cm")),
        axis.title.x = element_text(size = 16, margin = margin(t = 0.6, b = 0.6, unit = "cm")),
        plot.caption = element_text(hjust = 0, size = 8, margin = margin(t = 0.8, unit = "cm")),
        plot.title = element_text(hjust = 0.5, size = 20, margin = margin(b = 0.8, unit = "cm")),
        plot.subtitle = element_text(hjust = 0.5, size = 10, margin = margin(b = 0.8, unit = "cm")),
        legend.box.background = element_rect(color = "black"),
        plot.background = element_rect(color = "white"),
        axis.line = element_line(), 
        legend.text = element_text(size = 11), 
        legend.title = element_text(size = 14),
        panel.grid.major.y = element_blank()) +
  scale_y_continuous(breaks = seq(0, 6, 0.5)) +
 # scale_x_discrete(limits = cause_of_death_ordered[seq(1, length(cause_of_death_ordered), by = 2)]) +
  scale_fill_manual(values = c("0 - < 0.1 years" = "#FFFFFF", 
                               "0.1 - 0.5" = "#FFE6B3", 
                               "> 0.5 - 1" = "#FFD25D", 
                               "> 1 - 2" = "#FFBA00", 
                               "> 2 - 3" = "#FF9600", 
                               "> 3 - 4" = "#FF6908", 
                               "> 4 - 5" = "#E63D23", 
                               "> 5 - < 6" = "#BD251C", 
                               ">= 6" = "#8C130E")) + 
  guides(fill = guide_legend(nrow = 1)) +
  themes_aqli_base


#> appendix table-------------------------------

china_fs_appendix_table <- gadm2_aqli_2021 %>%
gadm_level_summary(c("country", "name_1", "name_2"), c(2014, 2021), 10) %>%
  filter(country == "China") %>%
  slice_max(population, n = 25) %>%
  mutate(pop_millions = round(population/1000000, 1), 
         le_gain_2014_to_2021 = round((pm2014 - pm2021)*0.098, 1)) %>%
  select(name_2, pop_millions, pm2021, le_gain_2014_to_2021,
         llpp_who_2021, llpp_nat_2021) %>%
  rename(Prefecture = name_2, `Population (Millions)` = pop_millions, 
         `PM2.5 concentration 2021 (µg/m³)` = pm2021, 
         `Life Expectancy Gains from reducing PM2.5 from 2014 concentrations to 2021 concentrations` = le_gain_2014_to_2021,
         `Life Expectancy Gains from reducing PM2.5 concentrations from 2021 Concentration to the WHO Guideline of 5 µg/m³` = llpp_who_2021, 
         `Life Expectancy Gains from reducing PM2.5 concentrations from 2021 Concentration to the National Standard of 35 µg/m³` = llpp_nat_2021)

china_fs_appendix_table %>% 
  write_csv("[June082023]china_fs_appendix_table.csv")


```


# North India Fact Sheet 2023

```{r}

# Filtering color dataset for India
color_2021_india <- gadm2_aqli_2021 %>%
  filter(country == "India") 

# adding a north India/rest of india region column (specifically Indo gangetic plain)
north_india <- c("West Bengal", "Uttar Pradesh", "Punjab", "Haryana", 
                 "Chandigarh", "Bihar", "NCT of Delhi")

# adding North India column
color_2021_india <- color_2021_india %>%
  mutate(region = ifelse(name_1 %in% north_india, "North India", "All other regions (excluding North India)"))

# quick stats
color_2021_india %>%
  filter(region == "North India") %>%
  gadm_level_summary(c("country"), c(2021, 1998), 10)

# north india
gadm2_aqli_2021 %>%
  slice_max(pm2021, n = 90)
  filter(region == "North India") %>%
  slice_max(pm2021, n = 10)


# plot 1: 2 regions map (North India, All other regions)
plt <- india_state %>%
    mutate(region = ifelse(NAME_1 %in% north_india, "North India", "Other States and UTs")) %>%
    select(-geometry, geometry) %>%
    st_as_sf() %>%
    ggplot() +
    geom_sf(mapping = aes(fill = region), lwd = 0.8) +
    scale_fill_manual(values = c("North India" = "slategray2", "Other States and UTs" = "lightgrey")) +
      labs(fill = "") + 
  ggthemes::theme_tufte() +
  theme(legend.position = "bottom", 
        axis.line = element_blank(), 
        axis.ticks = element_blank(), 
        axis.text = element_blank())
  

   
#> plot 2: AQLI map screenshot (I have limited this to the Indo-Gangetic plain as opposed to the current screenshot that shows it for all of India)------------------


#> plot 2: map screenshot--------------------

# north india figure1 dataset
north_india_fs_fig1_data <- gadm2_aqli_2021 %>%
  filter(country == "India", name_1 %in% north_india) %>%
  left_join(gadm2_aqli_2021_shp, by = c("objectid_gadm2" = "obidgadm2")) %>%
  mutate(lyl_bucket = ifelse((llpp_who_2021 >= 0) & (llpp_who_2021 < 0.1), "0 - < 0.1", NA), 
         lyl_bucket = ifelse((llpp_who_2021 >= 0.1) & (llpp_who_2021 <= 0.5), "0.1 - 0.5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 0.5) & (llpp_who_2021 <= 1), "> 0.5 - 1", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 1) & (llpp_who_2021 <= 2), "> 1 - 2", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 2) & (llpp_who_2021 <= 3), "> 2 - 3", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 3) & (llpp_who_2021 <= 4), "> 3 - 4", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 4) & (llpp_who_2021 <= 5), "> 4 - 5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 5) & (llpp_who_2021 < 6), "> 5 - < 6", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 >= 6), ">= 6", lyl_bucket)) %>%
  mutate(order_lyl_bucket = ifelse(lyl_bucket == "0 - < 0.1", 1, NA), 
         order_lyl_bucket = ifelse(lyl_bucket == "0.1 - 0.5", 2, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 0.5 - 1", 3, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 1 - 2", 4, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 2 - 3", 5, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 3 - 4", 6, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 4 - 5", 7, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 5 - < 6", 8, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == ">= 6", 9, order_lyl_bucket)) %>%
  select(-geometry, geometry) %>%
  st_as_sf()



# north india figure 2
north_india_fs_fig1 <- north_india_fs_fig1_data %>%
  ggplot() +
  geom_sf(mapping = aes(fill = forcats::fct_reorder(lyl_bucket, order_lyl_bucket)), color = "lightgrey", lwd = 0.05) +
 geom_sf(data = gadm1_aqli_2021_shp %>% filter(name0 == "India", name1 %in% north_india), color = "black", fill = "transparent", lwd = 0.5) +
  ggthemes::theme_map() + 
 scale_fill_manual(values = c("0 - < 0.1" = "#FFFFFF", 
                               "0.1 - 0.5" = "#FFE6B3", 
                               "> 0.5 - 1" = "#FFD25D", 
                               "> 1 - 2" = "#FFBA00", 
                               "> 2 - 3" = "#FF9600", 
                               "> 3 - 4" = "#FF6908", 
                               "> 4 - 5" = "#E63D23", 
                               "> 5 - < 6" = "#BD251C", 
                               ">= 6" = "#8C130E")) +
  ggthemes::theme_map() +
  labs(fill = "Gain in years of life expectancy  ", title = expression("Potential Gains in Life Expectancy (North India) through Permanently Reducing" ~ PM[2.5] ~ "from 2021 concentrations to the WHO guideline")) + 
 theme(legend.position = "bottom", 
        legend.justification = c(0.5, 3), 
        legend.background = element_rect(color = "black"), 
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15), 
        plot.title = element_text(hjust = 0.5, size = 15), 
       # legend.key = element_rect(color = "black"),
       legend.box.margin = margin(b = 1, unit = "cm"),
        plot.subtitle = element_text(hjust = 0.5, size = 7), 
        plot.caption = element_text(hjust = 0.7, size = 9, face = "italic"), 
       legend.key = element_rect(color = "black"), 
       legend.box.spacing = unit(0, "cm"), 
        legend.direction = "horizontal", 
       plot.background = element_rect(fill = "white", color = "white")) +
  guides(fill = guide_legend(nrow = 1)) +
   ggsflabel::geom_sf_label_repel(data = gadm1_aqli_2021_shp %>% filter(name0 == "India", name1 %in% north_india), mapping = aes(label = name1), force = 50, nudge_x = 3, seed = 10, size = 4)

  
  
  
  
#> plot 3: potential gain in life expectancy in 10 most populous districts of the Indo-Gangetic plains----------------

# fig3 dataset
north_india_fs_fig3_dataset <- gadm2_aqli_2021 %>%
  filter(country == "India", name_1 %in% north_india) %>%
  add_aqli_color_scale_buckets(scale_type = "lyl", col_name = "llpp_who_2021") %>%
  slice_max(population, n = 10)


# fig 3

north_india_fs_fig3 <- north_india_fs_fig3_dataset %>%
  aqli_bar(scale_type = "lyl", x_var = "name_2", y_var = "llpp_who_2021", title = expression("Potential Gain in Life Expectancy from Reducing 2021 " ~ PM[2.5] ~ "concentrations to the WHO Guideline"), subtitle = "(Ten most populous districts of North India)", x_label = "District", y_label = "Average Life Expectancy Gains (Years)", legend_title = "Life expectancy gains (Years)", caption = "") +
  scale_y_continuous(breaks = seq(0, 12, 2))
 

  
  

#> plot 4: Trend lines graph comparing Indo-Gangetic plains and all other regions in India-------------

# adding a north India/rest of india region column (specifically Indo gangetic plain)
north_india <- c("West Bengal", "Uttar Pradesh", "Punjab", "Haryana", 
                 "Chandigarh", "Bihar", "NCT of Delhi")

north_india_fs_fig4_data_part1 <- gadm2_aqli_2021 %>%
  filter(country == "India") %>%
  mutate(region = ifelse(name_1 %in% north_india, "North India", "All other regions (excluding North India)"))


# creating India's region wise average PM2.5 data from 1998 to 2021 
north_india_fs_fig4_data_part1 <- north_india_fs_fig4_data_part1 %>%
  group_by(region) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         mutate(across(starts_with("pm"), ~.x*pop_weights, .names = "{col}_weighted"))) %>%
  summarise(across(ends_with("weighted"), sum)) %>%
  pivot_longer(cols = pm1998_weighted:pm2021_weighted, names_to = "years", 
               values_to = "pop_weighted_avg_pm2.5") %>%
  mutate(years = as.integer(unlist(str_extract(years, "\\d+")))) %>%
  select(years, region, pop_weighted_avg_pm2.5)

# creating a national average PM2.5 trendlines data from 1998 to 2021
north_india_fs_fig4_data_part2 <- gadm2_aqli_2021 %>%
  filter(country == "India") %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         mutate(across(starts_with("pm"), ~.x*pop_weights, .names = "{col}_weighted"))) %>%
  summarise(across(ends_with("weighted"), sum)) %>%
  pivot_longer(cols = pm1998_weighted:pm2021_weighted, names_to = "years", 
               values_to = "pop_weighted_avg_pm2.5") %>%
  mutate(years = as.integer(unlist(str_extract(years, "\\d+"))), 
         region = "National Average") %>% 
  select(years, region, pop_weighted_avg_pm2.5)

# India factsheet figure 4 dataset
north_india_fs_fig4_data <- rbind(north_india_fs_fig4_data_part1, north_india_fs_fig4_data_part2)


# India factsheet figure 4
# fig4 caption removed: "* We define the Indo-Gangetic plain region as containing the following seven states and union territories: Bihar, Chandigarh, NCT of Delhi, Haryana, Punjab, Uttar Pradesh, and West Bengal."

north_india_fs_fig4_data$region <- factor(north_india_fs_fig4_data$region, levels = c("North India", "National Average", "All other regions (excluding North India)"))

north_india_fs_fig4 <- ggplot(north_india_fs_fig4_data) +
  geom_line(mapping = ggplot2::aes(x = years, y = pop_weighted_avg_pm2.5, color = region), lwd = 1.3) +
    geom_hline(mapping = aes(yintercept = 5), lwd = 0.8, linetype = "dotted") +
  scale_y_continuous(breaks = seq(0, 100, 10), limits = c(0, 100)) +
  scale_x_continuous(breaks = c(seq(1998, 2019, 3), 2021)) +
    scale_color_manual(values = c("North India" = "darkred", "National Average" = "darkgrey", "All other regions (excluding North India)" = "orange")) +
  ggthemes::theme_tufte() +
  labs(x = "Year", 
       y = expression("Annual Average   " ~ PM[2.5] ~ " concentrations (in µg/m³)"), 
       title = expression("Annual Average" ~ PM[2.5] ~ "concentrations in North India and other regions of India, 1998 to 2021")) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 11),
        legend.title = element_blank(),
        axis.title.y = element_text(size = 13, margin = margin(r = 0.6, unit = "cm")), 
        axis.title.x = element_text(size = 13, margin = margin(t = 0.6, b = 0.6, unit = "cm")), 
        axis.line = element_line(), 
        legend.box.background = element_rect(color = "black"),
        plot.title = element_text(hjust = 0.5, size = 16, margin = margin(b = 0.7, unit = "cm")), 
        plot.subtitle = element_text(hjust = 0.5, size = 8, face = "italic"), 
        plot.caption = element_text(size = 7, margin = margin(t = 0.8, unit = "cm"), hjust = 0, face = "italic"),
        axis.text = element_text(size = 12), 
        plot.background = element_rect(color = "white"), 
        axis.ticks = element_blank()) +
     geom_text(x = 2000.5, y = 7.9, label = expression("WHO" ~ PM[2.5] ~ "Guideline (last updated: 2021): 5 µg/m³"), size = 4.5) 

# appendix table

foo <- gadm2_aqli_2021 %>%
  filter(country == "India", name_1 %in% north_india) %>%
  gadm_level_summary(c("country", "name_1"), c(2021), perc_red_by = 10) %>%
  mutate(pop_millions = round(population/1000000, 1)) %>%
  select(name_1, pop_millions, pm2021, llpp_who_2021, llpp_nat_2021) %>%
  rename(State = name_1, `Population (in millions)` = pop_millions, 
         "Annual Average PM2.5 in 2021 (in µg/m³)" = pm2021, 
         "Life Expectancy Gains from reducing PM2.5 concentrations from 2021 Concentration to WHO Guideline of 5 µg/m³" = llpp_who_2021, 
         "Life Expectancy Gains from reducing PM2.5 concentrations from 2021 Concentration to National Guideline of 40 µg/m³" = llpp_nat_2021) %>% View()
  write_csv("./[June052023]north_india_fs_appendix_table.csv")



```


# Bangladesh factsheet 2023

```{r}

#> stats check/update

# by what multiple does particulate pollution exceed the WHO guideline in each of Bangladesh's districts
gadm2_aqli_2021 %>%
  filter(country == "Bangladesh") %>%
  mutate(mult_of_who = pm2021/5, 
         mult_of_nat = pm2021/15) %>%
  select(name_1, name_2, mult_of_who, mult_of_nat) %>% 
  slice_min(mult_of_who, n = 5) %>%
  View()


gadm2_aqli_2021 %>%
  filter(country == "Bangladesh") %>%
  slice_max(population, n = 5)

gadm2_aqli_2021 %>%
  gadm_level_summary(c("country"), c(2021), 10) %>%
  filter(country == "Bangladesh")


#> plot 1: map screenshot--------------------------------


# bangladesh figure1 dataset
bangladesh_fs_fig1_data <- gadm2_aqli_2021 %>%
  filter(country == "Bangladesh") %>%
  left_join(gadm2_aqli_2021_shp, by = c("objectid_gadm2" = "obidgadm2")) %>%
  mutate(lyl_bucket = ifelse((llpp_who_2021 >= 0) & (llpp_who_2021 < 0.1), "0 - < 0.1", NA), 
         lyl_bucket = ifelse((llpp_who_2021 >= 0.1) & (llpp_who_2021 <= 0.5), "0.1 - 0.5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 0.5) & (llpp_who_2021 <= 1), "> 0.5 - 1", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 1) & (llpp_who_2021 <= 2), "> 1 - 2", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 2) & (llpp_who_2021 <= 3), "> 2 - 3", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 3) & (llpp_who_2021 <= 4), "> 3 - 4", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 4) & (llpp_who_2021 <= 5), "> 4 - 5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 5) & (llpp_who_2021 < 6), "> 5 - < 6", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 >= 6), ">= 6", lyl_bucket)) %>%
  mutate(order_lyl_bucket = ifelse(lyl_bucket == "0 - < 0.1", 1, NA), 
         order_lyl_bucket = ifelse(lyl_bucket == "0.1 - 0.5", 2, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 0.5 - 1", 3, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 1 - 2", 4, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 2 - 3", 5, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 3 - 4", 6, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 4 - 5", 7, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 5 - < 6", 8, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == ">= 6", 9, order_lyl_bucket)) %>%
  select(-geometry, geometry) %>%
  st_as_sf()

# china_fs_fig1_data %>%
#   sf::write_sf("./september.2023/factsheets/southkorea/fig1/sk_fs_fig1.shp")


# bangladesh figure 1
bangladesh_fs_fig1 <- bangladesh_fs_fig1_data %>%
  ggplot() +
  geom_sf(mapping = aes(fill = forcats::fct_reorder(lyl_bucket, order_lyl_bucket)), color = "lightgrey", lwd = 0.05) +
 geom_sf(data = gadm1_aqli_2021_shp %>% filter(name0 == "Bangladesh"), color = "black", fill = "transparent", lwd = 0.5) +
  ggthemes::theme_map() + 
 scale_fill_manual(values = c("0 - < 0.1" = "#FFFFFF", 
                               "0.1 - 0.5" = "#FFE6B3", 
                               "> 0.5 - 1" = "#FFD25D", 
                               "> 1 - 2" = "#FFBA00", 
                               "> 2 - 3" = "#FF9600", 
                               "> 3 - 4" = "#FF6908", 
                               "> 4 - 5" = "#E63D23", 
                               "> 5 - < 6" = "#BD251C", 
                               ">= 6" = "#8C130E")) +
  ggthemes::theme_map() +
  labs(fill = "Gain in years of life expectancy  ", title = expression("Potential Gains in Life Expectancy (Bangladesh) through Permanently Reducing" ~ PM[2.5] ~ "from 2021 concentrations to the WHO guideline")) + 
 theme(legend.position = "bottom", 
        legend.justification = c(0.5, 3), 
        legend.background = element_rect(color = "black"), 
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15), 
        plot.title = element_text(hjust = 0.5, size = 15), 
       # legend.key = element_rect(color = "black"),
       legend.box.margin = margin(b = 1, unit = "cm"),
        plot.subtitle = element_text(hjust = 0.5, size = 7), 
        plot.caption = element_text(hjust = 0.7, size = 9, face = "italic"), 
       legend.key = element_rect(color = "black"), 
       legend.box.spacing = unit(0, "cm"), 
        legend.direction = "horizontal", 
       plot.background = element_rect(fill = "white", color = "white")) +
  guides(fill = guide_legend(nrow = 1))


#> plot 2: Potential Gain in Life Expectancy from Reducing PM2.5 from 2021 to the WHO Guideline in 10 most populous regions of Bangladesh -----------------------------------------

bangladesh_fs_fig2_dataset <- gadm2_aqli_2021 %>%
  filter(country == "Bangladesh") %>%
  add_aqli_color_scale_buckets(scale_type = "lyl", col_name = "llpp_who_2021") %>%
  slice_max(population, n = 10)

bangladesh_fs_fig2 <- bangladesh_fs_fig2_dataset %>%
  aqli_bar(scale_type = "lyl", x_var = "name_2", y_var = "llpp_who_2021", title = expression("Potential Gain in Life Expectancy from Reducing 2021 " ~ PM[2.5] ~ "concentrations to the WHO Guideline"), subtitle = "(Ten most populous districts of Bangladesh)", x_label = "District", y_label = "Average Life Expectancy Gains (Years)", legend_title = "LE gains (Years)", caption = "")


#> plot 3: GBD-------------------------------------------------------

# top 10 most deadly diseases
bangladesh_fs_fig3_dataset <- gbd_results_master_2021 %>%
  filter(country == "Bangladesh") %>%
  slice_max(lyl, n = 10)

colnames(bangladesh_fs_fig3_dataset)[3] <- c("llpp_who_2021")

bangladesh_fs_fig3_dataset <- bangladesh_fs_fig3_dataset %>%
    mutate(lyl_bucket = ifelse((llpp_who_2021 >= 0) & (llpp_who_2021 < 0.1), "0 - < 0.1 years", NA), 
         lyl_bucket = ifelse((llpp_who_2021 >= 0.1) & (llpp_who_2021 <= 0.5), "0.1 - 0.5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 0.5) & (llpp_who_2021 <= 1), "> 0.5 - 1", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 1) & (llpp_who_2021 <= 2), "> 1 - 2", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 2) & (llpp_who_2021 <= 3), "> 2 - 3", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 3) & (llpp_who_2021 <= 4), "> 3 - 4", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 4) & (llpp_who_2021 <= 5), "> 4 - 5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 5) & (llpp_who_2021 < 6), "> 5 - < 6", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 >= 6), ">= 6", lyl_bucket)) %>%
  mutate(order_lyl_bucket = ifelse(lyl_bucket == "0 - < 0.1 years", 1, NA), 
         order_lyl_bucket = ifelse(lyl_bucket == "0.1 - 0.5", 2, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 0.5 - 1", 3, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 1 - 2", 4, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 2 - 3", 5, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 3 - 4", 6, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 4 - 5", 7, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 5 - 6", 8, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == ">= 6", 9, order_lyl_bucket))


#> figure 3 plot

# figure 3 caption removed: Source: Global Burden of Disease (https://vizhub.healthdata.org/gbd-results/) causes and risks data and WHO Life Tables (https://apps.who.int/gho/data/node.main.LIFECOUNTRY?lang=en) were used combined with the Life table method to arrive at these results. 'PM2.5 relative to WHO Guideline' bar displays life years lost relative to the WHO guideline as calculated by latest AQLI (2021) data, which will be published in September 2023.

bangladesh_fs_fig3 <- bangladesh_fs_fig3_dataset %>%
  ggplot() + 
  geom_col(mapping = aes(x = forcats::fct_reorder(cause_of_death, llpp_who_2021), y = llpp_who_2021, fill = forcats::fct_reorder(lyl_bucket, order_lyl_bucket)), width = 0.4, color = "black") +
  labs(x = "Cause of Death", y = "Life years lost", fill = "Life years lost", 
       title = expression("Top 10 causes of reductions in life expectancy in Bangladesh")) +
  coord_flip() + 
  ggthemes::theme_tufte() +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 14),
        axis.title.y = element_text(size = 16, margin = margin(r = 0.6, unit = "cm")),
        axis.title.x = element_text(size = 16, margin = margin(t = 0.6, b = 0.6, unit = "cm")),
        plot.caption = element_text(hjust = 0, size = 8, margin = margin(t = 0.8, unit = "cm")),
        plot.title = element_text(hjust = 0.5, size = 20, margin = margin(b = 0.8, unit = "cm")),
        plot.subtitle = element_text(hjust = 0.5, size = 10, margin = margin(b = 0.8, unit = "cm")),
        legend.box.background = element_rect(color = "black"),
        plot.background = element_rect(color = "white"),
        axis.line = element_line(), 
        legend.text = element_text(size = 11), 
        legend.title = element_text(size = 14),
        panel.grid.major.y = element_blank()) +
  scale_y_continuous(breaks = seq(0, 6, 0.5)) +
 # scale_x_discrete(limits = cause_of_death_ordered[seq(1, length(cause_of_death_ordered), by = 2)]) +
  scale_fill_manual(values = c("0 - < 0.1 years" = "#FFFFFF", 
                               "0.1 - 0.5" = "#FFE6B3", 
                               "> 0.5 - 1" = "#FFD25D", 
                               "> 1 - 2" = "#FFBA00", 
                               "> 2 - 3" = "#FF9600", 
                               "> 3 - 4" = "#FF6908", 
                               "> 4 - 5" = "#E63D23", 
                               "> 5 - < 6" = "#BD251C", 
                               ">= 6" = "#8C130E")) + 
  guides(fill = guide_legend(nrow = 1)) +
  themes_aqli_base




#> plot 4: Trendlines: 1998 to 2021------------------------------------

# bangladesh factsheet figure 3 dataset
bangladesh_fs_fig4_dataset <- gadm2_aqli_2021 %>%
  filter(country == "Bangladesh") %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         mutate(across(starts_with("pm"), ~.x*pop_weights, .names = "{col}_weighted"))) %>%
  summarise(across(ends_with("weighted"), sum)) %>%
  pivot_longer(cols = pm1998_weighted:pm2021_weighted, names_to = "years", 
               values_to = "pop_weighted_avg_pm2.5") %>%
  mutate(years = as.integer(unlist(str_extract(years, "\\d+"))), 
         region = "National Average") %>% 
  select(years, region, pop_weighted_avg_pm2.5)




  bangladesh_fs_fig4 <- bangladesh_fs_fig4_dataset %>%
    ggplot() +
  geom_line(mapping = ggplot2::aes(x = years, y = pop_weighted_avg_pm2.5), lwd = 1.1, 
            color = "darkred") +
    geom_hline(mapping = aes(yintercept = 5), lwd = 0.8, linetype = "dashed") +
    geom_hline(mapping = aes(yintercept = 15), lwd = 0.8, linetype = "dashed") + 
  scale_y_continuous(breaks = seq(0, 80, 10), limits = c(0, 80)) +
  scale_x_continuous(breaks = c(seq(1998, 2019, 3), 2021)) +
    scale_color_manual(values = c("National Average" = "darkgrey")) +
  ggthemes::theme_clean() +
  labs(x = "Year", 
       y = expression("Annual Average  " ~ PM[2.5] ~ " concentrations (in µg/m³)"), 
       title = expression("Annual Average " ~ PM[2.5] ~ "concentrations in Bangladesh, 1998 to 2021")) +
  theme(legend.position = "bottom", legend.title = element_blank(), 
        legend.text = element_text(size = 7), 
        axis.title.y = element_text(size = 9), 
        axis.title.x = element_text(size = 9)) +  
    geom_text(x = 2001.8, y = 6.5, label =  expression("WHO " ~ PM[2.5] ~ " Guideline (last updated in 2021): 5 µg/m³"), size = 5) +
    geom_text(x = 2001.8, y = 16.5, label = expression("Bangladesh National" ~ PM[2.5] ~ "Guideline: 15 µg/m³"), size = 5) +
    themes_aqli_base
  




# appendix table (gadm level 2)
gadm2_aqli_2021 %>%
    filter(country == "Bangladesh") %>%
    select(name_1, name_2, population, pm2021, llpp_who_2021, llpp_nat_2021) %>%
    mutate(pop_millions = round(population/1000000, 1), 
           pm2021 = round(pm2021, 1), 
           llpp_who_2021 = round(llpp_who_2021, 1), 
           llpp_nat_2021 = round(llpp_nat_2021, 1)) %>%
  mutate(lyg_red_pm2.5_30perc_2021 = round((pm2021*0.3)*0.098, 1)) %>%
    slice_max(population, n = 25) %>% 
    select(name_1, name_2, pop_millions, pm2021, llpp_who_2021, llpp_nat_2021, lyg_red_pm2.5_30perc_2021) %>%
    rename(`Population (Millions)` = pop_millions, 
           "Division" = name_1, 
           "District" = name_2, 
           "PM2.5 concentration 2021 (in µg/m³)" = pm2021, 
           "Life Expectancy Gains from reducing PM2.5 from 2021 Concentrations to the National Guideline of 15 µg/m³" = llpp_nat_2021, 
           "Life Expectancy Gains from reducing PM2.5 from 2021 Concentrations to the WHO Guideline of 5 µg/m³" = llpp_who_2021, 
           "Life Expectancy Gains from reducing PM2.5 from 2021 Concentrations by 30 percent" = lyg_red_pm2.5_30perc_2021) %>% View()
    write_csv("./[updatedOnJune082023]Bangladesh Factsheet 2023 appendix table.csv")
 



```


# Central and West Africa factsheet

```{r}

# central africa definition
central_african_countries <- c("Angola", "Burundi", "Cameroon", 
                                        "Central African Republic", "Chad", 
                                        "Republic of the Congo",
                                        "Democratic Republic of the Congo", 
                                        "Equatorial Guinea", "Gabon",
                                        "São Tomé and Príncipe", 
                                        "Rwanda")
# west africa definition
west_african_countries <- c("Benin", "Burkina Faso", "Cabo Verde", 
                            "Gambia", "Ghana", "Guinea", "Guinea-Bissau", 
                            "Côte d'Ivoire", "Liberia", "Mali", "Mauritania", 
                            "Niger", "Nigeria", "Senegal", "Sierra Leone", 
                            "Togo")

# central and west africa countries definition, combine in a single vector
central_and_west_african_countries <- c(central_african_countries, west_african_countries)

# sanity check: check if our dataset has all of the above countries (yes, it does)
gadm2_aqli_2021 %>%
  filter(country %in% central_and_west_african_countries) %>%
  select(country) %>%
  unique() 

# central and west african countries
gadm2_aqli_2021_cw_african <- gadm2_aqli_2021 %>%
  filter(country %in% central_and_west_african_countries) %>%
  mutate(name_1_country = str_c(name_1, " (", country, ")", sep = ""), 
         region = ifelse(country %in% central_african_countries, "Central African", 
                         "West African")) 

sum(gadm2_aqli_2021_cw_african$population, na.rm = TRUE)


#> stats check----------------------------------

# c/w africa region wide average stats
gadm2_aqli_2021_cw_african %>%
  mutate(country = "foo") %>%
  gadm_level_summary(c("country"), c(2021), 10) 

# top polluting countries
gadm2_aqli_2021_cw_african %>%
  gadm_level_summary(c("country"), c(2021), 10) 




#> plot 1: map screenshot--------------------------------


# cw africa figure1 dataset
cw_africa_fs_fig1_data <- gadm2_aqli_2021_cw_african %>%
  left_join(gadm2_aqli_2021_shp, by = c("objectid_gadm2" = "obidgadm2")) %>%
  mutate(lyl_bucket = ifelse((llpp_who_2021 >= 0) & (llpp_who_2021 < 0.1), "0 - < 0.1", NA), 
         lyl_bucket = ifelse((llpp_who_2021 >= 0.1) & (llpp_who_2021 <= 0.5), "0.1 - 0.5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 0.5) & (llpp_who_2021 <= 1), "> 0.5 - 1", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 1) & (llpp_who_2021 <= 2), "> 1 - 2", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 2) & (llpp_who_2021 <= 3), "> 2 - 3", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 3) & (llpp_who_2021 <= 4), "> 3 - 4", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 4) & (llpp_who_2021 <= 5), "> 4 - 5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 5) & (llpp_who_2021 < 6), "> 5 - < 6", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 >= 6), ">= 6", lyl_bucket)) %>%
  mutate(order_lyl_bucket = ifelse(lyl_bucket == "0 - < 0.1", 1, NA), 
         order_lyl_bucket = ifelse(lyl_bucket == "0.1 - 0.5", 2, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 0.5 - 1", 3, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 1 - 2", 4, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 2 - 3", 5, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 3 - 4", 6, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 4 - 5", 7, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 5 - < 6", 8, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == ">= 6", 9, order_lyl_bucket)) %>%
  select(-geometry, geometry) %>%
  st_as_sf()

# china_fs_fig1_data %>%
#   sf::write_sf("./september.2023/factsheets/southkorea/fig1/sk_fs_fig1.shp")


# cw africa fig1
cw_africa_fig1 <- cw_africa_fs_fig1_data %>%
  ggplot() +
  geom_sf(mapping = aes(fill = forcats::fct_reorder(lyl_bucket, order_lyl_bucket)), color = "lightgrey", lwd = 0.05) +
   geom_sf(data = gadm1_aqli_2021_shp %>% filter(name0 %in% central_and_west_african_countries), color = "black", fill = "transparent", lwd = 0.1) +
 geom_sf(data = gadm0_aqli_2021_shp %>% filter(name0 %in% central_and_west_african_countries), color = "black", fill = "transparent", lwd = 0.6) +
  ggthemes::theme_map() + 
 scale_fill_manual(values = c("0 - < 0.1" = "#FFFFFF", 
                               "0.1 - 0.5" = "#FFE6B3", 
                               "> 0.5 - 1" = "#FFD25D", 
                               "> 1 - 2" = "#FFBA00", 
                               "> 2 - 3" = "#FF9600", 
                               "> 3 - 4" = "#FF6908", 
                               "> 4 - 5" = "#E63D23", 
                               "> 5 - < 6" = "#BD251C", 
                               ">= 6" = "#8C130E")) +
  ggthemes::theme_map() +
  labs(fill = "Gain in years of life expectancy  ", title = expression("Potential Gains in Life Expectancy through Permanently Reducing" ~ PM[2.5] ~ "from 2021 concentrations to the WHO guideline"), 
       subtitle = "(Central and West Africa)") + 
 theme(legend.position = "bottom", 
        legend.justification = c(0.5, 3), 
        legend.background = element_rect(color = "black"), 
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15), 
        plot.title = element_text(hjust = 0.5, size = 15), 
       # legend.key = element_rect(color = "black"),
       legend.box.margin = margin(b = 1, unit = "cm"),
        plot.subtitle = element_text(hjust = 0.5, size = 12), 
        plot.caption = element_text(hjust = 0.7, size = 9, face = "italic"), 
       legend.key = element_rect(color = "black"), 
       legend.box.spacing = unit(0, "cm"), 
        legend.direction = "horizontal", 
       plot.background = element_rect(fill = "white", color = "white")) +
  guides(fill = guide_legend(nrow = 1))



#> cw africa factsheet plot 2-------------------------------------


#> cwafrica fig2

# GBD results filtered for relevant cause of death and countries 
gbd_results_cwafrica_fig2 <- gbd_results_master_2021 %>%
  filter(cause_of_death %in% c("PM2.5 relative to WHO guideline",
                               "Neglected tropical diseases and malaria", "Unsafe water, sanitation, and handwashing", 
                               "HIV/AIDS and sexually transmitted infections"), country %in% c("Nigeria", "Democratic Republic of the Congo", "Angola", "Ghana", "Cameroon"))

# making the 'location' column of type factor
gbd_results_cwafrica_fig2$country <- factor(gbd_results_cwafrica_fig2$country, 
                                        levels = c("Nigeria", "Democratic Republic of the Congo", "Angola", "Ghana", "Cameroon"))

# Rename Democratic Republic of the Congo to DRC
gbd_results_cwafrica_fig2$country <-  str_replace(gbd_results_cwafrica_fig2$country, 
            "Democratic Republic of the Congo", "DR Congo")

# Converting 'cause_of_death' to type factor
gbd_results_cwafrica_fig2$cause_of_death <- as.factor(gbd_results_cwafrica_fig2$cause_of_death)

# Rearranging 'cause of death' levels
levels(gbd_results_cwafrica_fig2$cause_of_death) <- c( "HIV/AIDS and sexually transmitted infections", "Neglected tropical diseases and malaria", "PM2.5 relative to WHO guideline", "Unsafe water, sanitation, and handwashing")

# wrapping x-axis labels text 
levels(gbd_results_cwafrica_fig2$cause_of_death) <- str_wrap(levels(gbd_results_cwafrica_fig2$cause_of_death), 30)

# getting country wise population
country_wise_population <- gadm2_aqli_2021_cw_african %>%
  gadm_level_summary(c("country"), c(2021), 10) %>%
  filter(country %in% c("Cameroon", "Nigeria", "Angola", "Ghana", 
                  "Democratic Republic of the Congo")) %>%
  arrange(desc(population))

# reorder within each location as per the total life years lost column
gbd_results_cwafrica_fig2 <- gbd_results_cwafrica_fig2 %>% 
  mutate(cause_of_death = reorder_within(cause_of_death, lyl, country))

# clean the "cause of death" column and save the cwafrica factsheet figure 2
cwafrica_fs_fig2_dataset <- gbd_results_cwafrica_fig2 %>%
  mutate(cause_of_death = str_remove(cause_of_death, "___.+"))


# plot 
plt <- cwafrica_fs_fig2_dataset %>%
  ggplot(mapping = aes(x = reorder_within(cause_of_death, lyl, country), y = lyl)) + 
  geom_col(mapping = aes(fill = cause_of_death), width = 0.5, color = "white") +
  scale_x_reordered() +
  facet_wrap(~factor(country, levels = c("Angola", "Cameroon", "DR Congo", 
                  "Ghana", "Nigeria")), scales = "free_x", ncol = 5) +
  scale_fill_manual(values = c("#D3D9E0", "#7BC1D9", "#8F3931",
                               "#808A94")) +
  labs(x = "Cause of Death", y = "Years Lost", title = expression("Life Expectancy Impacts of" ~ PM[2.5] ~ "& Other Health Threats"), 
       subtitle = "Five Most Populous Countries in Central and West Africa", fill = "Cause of Death") +
  themes_aqli_base +
   theme(axis.text.x = element_blank(), legend.position = "bottom", axis.ticks = element_blank(), 
         strip.text = element_text(size = 12)) 


#> cw africa figure 3----------------------------------


# central and west africa figure 3 dataset
 cwafrica_fs_fig3_dataset <- gadm2_aqli_2021_cw_african %>%
  group_by(name_1_country) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2021_pop_weighted = pop_weights*pm2021) %>%
  summarise(tot_pop = sum(population, na.rm = TRUE), avg_pm2.5_pop_weighted = sum(pm2021_pop_weighted, na.rm = TRUE), 
            gain_in_le = (avg_pm2.5_pop_weighted - 5) * 0.098, 
            gain_in_le = ifelse(gain_in_le < 0, 0, gain_in_le)) %>%
  ungroup() %>%
  slice_max(tot_pop, n = 10) 
 

cwafrica_fs_fig3_dataset <- cwafrica_fs_fig3_dataset %>%
  add_aqli_color_scale_buckets(scale_type = "lyl", col_name = "gain_in_le") 

cwafrica_fs_fig3 <- cwafrica_fs_fig3_dataset %>%
  mutate(name_1_country = str_replace(name_1_country, "Democratic Republic of the Congo", "DR Congo")) %>%
  aqli_bar(scale_type = "lyl", x_var = "name_1_country", y_var = "gain_in_le", title = expression("Potential Gain in Life Expectancy from Reducing 2021 " ~ PM[2.5] ~ "concentrations to the WHO Guideline"), subtitle = "(Ten most populous regions of Central and West Africa)", x_label = "Region", y_label = "Average Life Expectancy Gains (Years)", legend_title = "Life expectancy gains (Years)", caption = "")


#> cw africa figure 4 trendlines -------------------------------------------------

#> Figure 4: Average PM2.5 concentration in India from 1998 to 2021------------------------


# creating India's region wise average PM2.5 data from 1998 to 2021 
cwafrica_fs_fig4_data <- gadm2_aqli_2021_cw_african %>%
  group_by(region) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         mutate(across(starts_with("pm"), ~.x*pop_weights, .names = "{col}_weighted"))) %>%
  summarise(across(ends_with("weighted"), sum)) %>%
  pivot_longer(cols = pm1998_weighted:pm2021_weighted, names_to = "years", 
               values_to = "pop_weighted_avg_pm2.5") %>%
  mutate(years = as.integer(unlist(str_extract(years, "\\d+")))) %>%
  select(years, region, pop_weighted_avg_pm2.5)


# India factsheet figure 4
# fig4 caption removed: "* We define the Indo-Gangetic plain region as containing the following seven states and union territories: Bihar, Chandigarh, NCT of Delhi, Haryana, Punjab, Uttar Pradesh, and West Bengal."

cwafrica_fs_fig4_data$region <- factor(cwafrica_fs_fig4_data$region, levels = c("Central African", "West African"))

cwafrica_fs_fig4 <- cwafrica_fs_fig4_data %>%
  ggplot() +
  geom_line(mapping = ggplot2::aes(x = years, y = pop_weighted_avg_pm2.5, color = region), lwd = 1.3) +
    geom_hline(mapping = aes(yintercept = 5), lwd = 0.8, linetype = "dotted") +
  scale_y_continuous(breaks = seq(0, 40, 10), limits = c(0, 40)) +
  scale_x_continuous(breaks = c(seq(1998, 2019, 3), 2021)) +
    scale_color_manual(values = c("Central African" = "darkred", "West African" = "orange")) +
  ggthemes::theme_tufte() +
  labs(x = "Year", 
       y = expression("Annual Average   " ~ PM[2.5] ~ " concentrations (in µg/m³)"), 
       title = expression("Annual Average" ~ PM[2.5] ~ "concentrations in India, 1998 to 2021")) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 11),
        legend.title = element_blank(),
        axis.title.y = element_text(size = 13, margin = margin(r = 0.6, unit = "cm")), 
        axis.title.x = element_text(size = 13, margin = margin(t = 0.6, b = 0.6, unit = "cm")), 
        axis.line = element_line(), 
        legend.box.background = element_rect(color = "black"),
        plot.title = element_text(hjust = 0.5, size = 16, margin = margin(b = 0.7, unit = "cm")), 
        plot.subtitle = element_text(hjust = 0.5, size = 8, face = "italic"), 
        plot.caption = element_text(size = 7, margin = margin(t = 0.8, unit = "cm"), hjust = 0, face = "italic"),
        axis.text = element_text(size = 12), 
        plot.background = element_rect(color = "white"), 
        axis.ticks = element_blank()) +
     geom_text(x = 2000.5, y = 6, label = expression("WHO" ~ PM[2.5] ~ "Guideline (last updated: 2021): 5 µg/m³"), size = 4.5) 


  
#> appendix table cw africa (top 50 most polluted regions in cw africa)-------------------------
gadm2_aqli_2021_cw_african %>%
  slice_max(pm2021, n = 50) %>%
  mutate(le_gain_30_perc_red = round((pm2021*0.3)*0.098, 1), 
         pop_hund_thou = round(population/100000, 1), 
         pm2021 = round(pm2021, 1), 
         llpp_who_2021 = round(llpp_who_2021, 1)) %>%
  select(country, name_2, pop_hund_thou, pm2021, llpp_who_2021, le_gain_30_perc_red) %>%
   rename(`Population (in 100,000s)` = pop_hund_thou, 
           "Country" = country, 
           "Region" = name_2, 
           "PM2.5 concentration 2021 (in µg/m³)" = pm2021, 
           "Life Expectancy Gains from reducing PM2.5 from 2021 Concentrations to the WHO Guideline of 5 µg/m³" = llpp_who_2021, 
           "Life Expectancy Gains from reducing PM2.5 from 2021 Concentrations by 30 percent" = le_gain_30_perc_red) %>% 
  write_csv("./[June092023]cwafrica_fs_appendix_table.csv")





```


# sanity checks section (deactivated, just for testing purposes)

```{r eval=FALSE, include=FALSE}

# Nepal

nepal_fs_fig1_data <- gadm2_aqli_2021 %>%
  filter(country == "Nepal") %>%
  left_join(gadm2_aqli_2021_shp, by = c("objectid_gadm2" = "objidgadm2")) %>%
  mutate(lyl_bucket = ifelse((llpp_who_2021 >= 0) & (llpp_who_2021 < 0.1), "0 - < 0.1", NA), 
         lyl_bucket = ifelse((llpp_who_2021 >= 0.1) & (llpp_who_2021 <= 0.5), "0.1 - 0.5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 0.5) & (llpp_who_2021 <= 1), "> 0.5 - 1", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 1) & (llpp_who_2021 <= 2), "> 1 - 2", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 2) & (llpp_who_2021 <= 3), "> 2 - 3", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 3) & (llpp_who_2021 <= 4), "> 3 - 4", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 4) & (llpp_who_2021 <= 5), "> 4 - 5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 5) & (llpp_who_2021 < 6), "> 5 - < 6", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 >= 6), ">= 6", lyl_bucket)) %>%
  mutate(order_lyl_bucket = ifelse(lyl_bucket == "0 - < 0.1", 1, NA), 
         order_lyl_bucket = ifelse(lyl_bucket == "0.1 - 0.5", 2, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 0.5 - 1", 3, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 1 - 2", 4, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 2 - 3", 5, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 3 - 4", 6, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 4 - 5", 7, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 5 - < 6", 8, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == ">= 6", 9, order_lyl_bucket)) %>%
  select(-geometry, geometry) %>%
  st_as_sf()

south_korea_fs_fig1_data %>%
  sf::write_sf("./september.2023/factsheets/southkorea/fig1/sk_fs_fig1.shp")


# south korea figure 1
nepal_fs_fig1 <- nepal_fs_fig1_data %>%
  ggplot() +
  geom_sf(mapping = aes(fill = forcats::fct_reorder(lyl_bucket, order_lyl_bucket)), color = "transparent") +
  geom_sf(data = gadm1_aqli_2021_shp %>% filter(name0 == "Nepal"), color = "black", fill = "transparent", lwd = 0.5) +
  ggthemes::theme_map() + 
 scale_fill_manual(values = c("0 - < 0.1 years" = "#FFFFFF", 
                               "0.1 - 0.5" = "#FFE6B3", 
                               "> 0.5 - 1" = "#FFD25D", 
                               "> 1 - 2" = "#FFBA00", 
                               "> 2 - 3" = "#FF9600", 
                               "> 3 - 4" = "#FF6908", 
                               "> 4 - 5" = "#E63D23", 
                               "> 5 - < 6" = "#BD251C", 
                               ">= 6" = "#8C130E")) +
  ggthemes::theme_map() +
  labs(fill = "Life years lost  ", title = expression("Potential Gains in Life Expectancy through Permanently Reducing" ~ PM[2.5] ~ "from 2021 concentrations to the WHO guideline")) + 
 theme(legend.position = "bottom", 
        legend.justification = c(0.5, 3), 
        legend.background = element_rect(color = "black"), 
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15), 
        plot.title = element_text(hjust = 0.5, size = 15), 
       # legend.key = element_rect(color = "black"),
       legend.box.margin = margin(b = 1, unit = "cm"),
        plot.subtitle = element_text(hjust = 0.5, size = 7), 
        plot.caption = element_text(hjust = 0.7, size = 9, face = "italic"), 
       legend.key = element_rect(color = "black"), 
       legend.box.spacing = unit(0, "cm"), 
        legend.direction = "horizontal") +
  guides(fill = guide_legend(nrow = 1))


# nepal fs figure 3

# top 5 most deadly diseases
nepal_fs_fig3_dataset <- gbd_results_master_2021 %>%
  filter(country == "Nepal")

colnames(nepal_fs_fig3_dataset)[3] <- c("llpp_who_2021")

nepal_fs_fig3_dataset <- nepal_fs_fig3_dataset %>%
    mutate(lyl_bucket = ifelse((llpp_who_2021 >= 0) & (llpp_who_2021 < 0.1), "0 - < 0.1 years", NA), 
         lyl_bucket = ifelse((llpp_who_2021 >= 0.1) & (llpp_who_2021 <= 0.5), "0.1 - 0.5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 0.5) & (llpp_who_2021 <= 1), "> 0.5 - 1", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 1) & (llpp_who_2021 <= 2), "> 1 - 2", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 2) & (llpp_who_2021 <= 3), "> 2 - 3", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 3) & (llpp_who_2021 <= 4), "> 3 - 4", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 4) & (llpp_who_2021 <= 5), "> 4 - 5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 5) & (llpp_who_2021 < 6), "> 5 - < 6 years", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 >= 6), ">= 6 years", lyl_bucket)) %>%
  mutate(order_lyl_bucket = ifelse(lyl_bucket == "0 - < 0.1 years", 1, NA), 
         order_lyl_bucket = ifelse(lyl_bucket == "0.1 - 0.5", 2, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 0.5 - 1", 3, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 1 - 2", 4, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 2 - 3", 5, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 3 - 4", 6, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 4 - 5", 7, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 5 - 6 years", 8, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == ">= 6 years", 9, order_lyl_bucket))


# getting disease (ordered from most deadly to least deadly) in a vector (to be used later in the
# scale_x_discrete function).

# cause_of_death_ordered <- india_fs_fig3_dataset %>%
#   arrange(llpp_who_2021) %>%
#   pull(cause_of_death) %>%
#   unlist()


# figure 3 plot
nepal_fs_fig3 <- nepal_fs_fig3_dataset %>%
  ggplot() + 
  geom_col(mapping = aes(x = forcats::fct_reorder(cause_of_death, llpp_who_2021), y = llpp_who_2021, fill = forcats::fct_reorder(lyl_bucket, order_lyl_bucket)), width = 0.5, color = "black") +
  labs(x = "Cause of Death", y = "Life Years Lost", fill = "Life years lost", 
       caption = str_wrap("* Source: Global Burden of Disease (https://vizhub.healthdata.org/gbd-results/) causes and risks data and WHO Life Tables (https://apps.who.int/gho/data/node.main.LIFECOUNTRY?lang=en) were used combined with the Life table method to arrive at these results. 'PM2.5 relative to WHO Guideline' bar displays life years lost relative to the WHO guideline as calculated by latest AQLI (2021) data, which will be published in September 2023.", width = 130), title = expression("Life Expectancy Impact of" ~ PM[2.5] ~ "and Unassociated Causes/Risks of Deaths in India")) +
  coord_flip() + 
  ggthemes::theme_tufte() +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 14),
        axis.title.y = element_text(size = 16, margin = margin(r = 0.6, unit = "cm")),
        axis.title.x = element_text(size = 16, margin = margin(t = 0.6, b = 0.6, unit = "cm")),
        plot.caption = element_text(hjust = 0, size = 8, margin = margin(t = 0.8, unit = "cm")),
        plot.title = element_text(hjust = 0.5, size = 20, margin = margin(b = 0.8, unit = "cm")),
        plot.subtitle = element_text(hjust = 0.5, size = 10, margin = margin(b = 0.8, unit = "cm")),
        legend.box.background = element_rect(color = "black"),
        plot.background = element_rect(color = "white"),
        axis.line = element_line(), 
        legend.text = element_text(size = 11), 
        legend.title = element_text(size = 14),
        panel.grid.major.y = element_blank()) +
  scale_y_continuous(breaks = seq(0, 6, 0.5)) +
 # scale_x_discrete(limits = cause_of_death_ordered[seq(1, length(cause_of_death_ordered), by = 2)]) +
  scale_fill_manual(values = c("0 - < 0.1 years" = "#FFFFFF", 
                               "0.1 - 0.5" = "#FFE6B3", 
                               "> 0.5 - 1" = "#FFD25D", 
                               "> 1 - 2" = "#FFBA00", 
                               "> 2 - 3" = "#FF9600", 
                               "> 3 - 4" = "#FF6908", 
                               "> 4 - 5" = "#E63D23", 
                               "> 5 - < 6 years" = "#BD251C", 
                               ">= 6 years" = "#8C130E")) + 
  guides(fill = guide_legend(nrow = 1))


# outer terai region lyl calculation
outer_terai <- c("Kapilbastu", "Rupandehi", "Nawalparasi West", "Nawalparasi East", "Parsa", "Bara", "Rautahat", "Sarlahi", "Mahottari", "Dhanusha", "Siraha", "Saptari", "Sunsari", "Morang", "Jhapa")

south_nepal <- c("Kapilbastu", "Rupandehi", "Nawalparasi West", "Parsa", "Bara", "Rautahat", "Sarlahi", "Mahottari", "Siraha", "Saptari")

gadm2_aqli_2021 %>%
  filter(country == "Nepal", name_2 %in% outer_terai) %>%
  gadm_level_summary(c("country"), c(2021), 10)


# 

#> What percent of the world’s population live in places, at the country-level, (district level) where they are losing at least 2 years (3 years) of life)?

threshold <- 2

# global total population
tot_pop_world <- gadm2_aqli_2021 %>%
  summarise(tot_pop = sum(population, na.rm = TRUE))


# district wise
tot_pop_atleast_eq_thresh <- gadm2_aqli_2021 %>%
  filter(llpp_who_2021 >= threshold) %>%
  summarise(tot_pop = sum(population, na.rm = TRUE))


perc_pop_losing_atleast_x_lyl_level <- function(threshold, level = "district"){
  tot_pop_world <- gadm2_aqli_2021 %>%
  summarise(tot_pop = sum(population, na.rm = TRUE)) 
  
  if(level == "district"){
 
 tot_pop_atleast_eq_thresh <- gadm2_aqli_2021 %>%
  filter(llpp_who_2021 >= threshold) %>%
  summarise(tot_pop = sum(population, na.rm = TRUE))
 
 return(round((tot_pop_atleast_eq_thresh/tot_pop_world)*100, 4))

  } else if (level == "state"){
     tot_pop_atleast_eq_thresh <- gadm1_aqli_2021 %>%
  filter(llpp_who_2021 >= threshold) %>%
  summarise(tot_pop = sum(population, na.rm = TRUE))
     
      return(round((tot_pop_atleast_eq_thresh/tot_pop_world)*100, 4))
     
  } else if (level == "country"){
     tot_pop_atleast_eq_thresh <- gadm0_aqli_2021 %>%
  filter(llpp_who_2021 >= threshold) %>%
  summarise(tot_pop = sum(population, na.rm = TRUE))
     
      return(round((tot_pop_atleast_eq_thresh/tot_pop_world)*100, 4))
  } 
}
  

df_lyl <- tibble(min_lyl = seq(0, 13, 0.2), district_level_calc_perc_pop_above_lyl = NA, 
                 state_level_calc_perc_pop_above_lyl = NA, 
                 country_level_calc_perc_pop_above_lyl = NA)

for(i in 1:nrow(df_lyl)){
  df_lyl$district_level_calc_perc_pop_above_lyl[i] <- perc_pop_losing_atleast_x_lyl_level(df_lyl$min_lyl[i], "district")
  df_lyl$state_level_calc_perc_pop_above_lyl[i] <- perc_pop_losing_atleast_x_lyl_level(df_lyl$min_lyl[i], "state")
  df_lyl$country_level_calc_perc_pop_above_lyl[i] <- perc_pop_losing_atleast_x_lyl_level(df_lyl$min_lyl[i], "country")
}



df_lyl <- df_lyl %>%
  mutate(district_level_calc_perc_pop_above_lyl = perc_pop_losing_atleast_x_lyl_level(df_lyl, "district"), 
         state_level_calc_perc_pop_above_lyl = perc_pop_losing_atleast_x_lyl_level(df_lyl, "state"), 
         country_level_calc_perc_pop_above_lyl = perc_pop_losing_atleast_x_lyl_level(df_lyl, "country"))


df_lyl_long <- df_lyl %>%
  pivot_longer(district_level_calc_perc_pop_above_lyl:country_level_calc_perc_pop_above_lyl, names_to = "summary_level", values_to = "percentage_pop_atleast_x_lyl")

df_lyl_long <- df_lyl_long %>%
  mutate(summary_level = ifelse(summary_level == "district_level_calc_perc_pop_above_lyl", "district", summary_level), 
         summary_level = ifelse(summary_level == "state_level_calc_perc_pop_above_lyl", "state", summary_level), 
         summary_level = ifelse(summary_level == "country_level_calc_perc_pop_above_lyl", "country", summary_level))

df_lyl_long$min_lyl <- as.numeric(unlist(df_lyl_long$min_lyl))
df_lyl_long$percentage_pop_atleast_x_lyl <- as.numeric(unlist(df_lyl_long$percentage_pop_atleast_x_lyl))
df_lyl_long$summary_level <- as.factor(df_lyl_long$summary_level)


plt <- df_lyl_long %>%
  ggplot() +
  geom_smooth(mapping = aes(x = min_lyl, y = percentage_pop_atleast_x_lyl, color = summary_level), lwd = 2, se =FALSE) +
  labs(x = "Minimum life years lost", y = "% population corresp to min lyl", color = "Summary level", 
       title = "Percent population losing atleast x life years at different summary levels") +
  scale_x_continuous(breaks = seq(0, 12, 2)) +
  scale_color_viridis(discrete = TRUE) +
  themes_aqli_base 


# pakistan factsheet map (world view adjusted)

# pakistan figure1 dataset
pak_fs_fig1_data <- gadm2_aqli_2021 %>%
  filter(country == "Pakistan") %>%
  left_join(gadm2_aqli_2021_shp, by = c("objectid_gadm2" = "obidgadm2")) %>%
  mutate(lyl_bucket = ifelse((llpp_who_2021 >= 0) & (llpp_who_2021 < 0.1), "0 - < 0.1", NA), 
         lyl_bucket = ifelse((llpp_who_2021 >= 0.1) & (llpp_who_2021 <= 0.5), "0.1 - 0.5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 0.5) & (llpp_who_2021 <= 1), "> 0.5 - 1", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 1) & (llpp_who_2021 <= 2), "> 1 - 2", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 2) & (llpp_who_2021 <= 3), "> 2 - 3", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 3) & (llpp_who_2021 <= 4), "> 3 - 4", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 4) & (llpp_who_2021 <= 5), "> 4 - 5", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 > 5) & (llpp_who_2021 < 6), "> 5 - < 6", lyl_bucket), 
         lyl_bucket = ifelse((llpp_who_2021 >= 6), ">= 6", lyl_bucket)) %>%
  mutate(order_lyl_bucket = ifelse(lyl_bucket == "0 - < 0.1", 1, NA), 
         order_lyl_bucket = ifelse(lyl_bucket == "0.1 - 0.5", 2, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 0.5 - 1", 3, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 1 - 2", 4, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 2 - 3", 5, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 3 - 4", 6, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 4 - 5", 7, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == "> 5 - < 6", 8, order_lyl_bucket), 
         order_lyl_bucket = ifelse(lyl_bucket == ">= 6", 9, order_lyl_bucket)) %>%
  select(-geometry, geometry) %>%
  st_as_sf()

# china_fs_fig1_data %>%
#   sf::write_sf("./september.2023/factsheets/southkorea/fig1/sk_fs_fig1.shp")


# china figure 1
pak_fs_fig1 <- pak_fs_fig1_data %>%
  ggplot() +
  geom_sf(mapping = aes(fill = forcats::fct_reorder(lyl_bucket, order_lyl_bucket)), color = "lightgrey", lwd = 0.05) +
 geom_sf(data = gadm1_aqli_2021_shp %>% filter(name0 == "Pakistan"), color = "black", fill = "transparent", lwd = 0.5) +
  ggthemes::theme_map() + 
 scale_fill_manual(values = c("0 - < 0.1" = "#FFFFFF", 
                               "0.1 - 0.5" = "#FFE6B3", 
                               "> 0.5 - 1" = "#FFD25D", 
                               "> 1 - 2" = "#FFBA00", 
                               "> 2 - 3" = "#FF9600", 
                               "> 3 - 4" = "#FF6908", 
                               "> 4 - 5" = "#E63D23", 
                               "> 5 - < 6" = "#BD251C", 
                               ">= 6" = "#8C130E")) +
  ggthemes::theme_map() +
  labs(fill = "Gain in years of life expectancy  ", title = expression("Potential Gains in Life Expectancy (China) through Permanently Reducing" ~ PM[2.5] ~ "from 2021 concentrations to the WHO guideline")) + 
 theme(legend.position = "bottom", 
        legend.justification = c(0.5, 3), 
        legend.background = element_rect(color = "black"), 
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 15), 
        plot.title = element_text(hjust = 0.5, size = 15), 
       # legend.key = element_rect(color = "black"),
       legend.box.margin = margin(b = 1, unit = "cm"),
        plot.subtitle = element_text(hjust = 0.5, size = 7), 
        plot.caption = element_text(hjust = 0.7, size = 9, face = "italic"), 
       legend.key = element_rect(color = "black"), 
       legend.box.spacing = unit(0, "cm"), 
        legend.direction = "horizontal", 
       plot.background = element_rect(fill = "white", color = "white")) +
  guides(fill = guide_legend(nrow = 1))



# Pakistan factsheet

# filtering color level data for Pakistan and adding a "region" column
color_2020_pak <- gadm2_aqli_2021 %>%
  filter(country == "Pakistan") %>%
  mutate(region = ifelse(name_1  %in% c("Islamabad", "Punjab"), 
                         name_1, "All other regions (excluding Islanabad and Punjab)"))


# region wise pakistan data for 1998 to 2020
region_wise_pak_average_ts <- color_2020_pak %>%
  group_by(region) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         mutate(across(starts_with("pm"), ~.x*pop_weights, .names = "{col}_weighted"))) %>%
  summarise(across(ends_with("weighted"), sum)) %>%
  pivot_longer(cols = pm1998_weighted:pm2021_weighted, names_to = "years", 
               values_to = "pop_weighted_avg_pm2.5") %>%
  mutate(years = as.integer(unlist(str_extract(years, "\\d+")))) %>%
  select(years, region, pop_weighted_avg_pm2.5)

# national average pakistan data for 1998 to 2020
 national_average_pak_ts <- color_2020_pak %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         mutate(across(starts_with("pm"), ~.x*pop_weights, .names = "{col}_weighted"))) %>%
  summarise(across(ends_with("weighted"), sum)) %>%
  pivot_longer(cols = pm1998_weighted:pm2020_weighted, names_to = "years", 
               values_to = "pop_weighted_avg_pm2.5") %>%
  mutate(years = as.integer(unlist(str_extract(years, "\\d+"))), 
         region = "National Average") %>% 
  select(years, region, pop_weighted_avg_pm2.5)

# pakistan factsheet figure 4 final dataset 
pak_fs_fig4_dataset <- rbind(region_wise_pak_average_ts, national_average_pak_ts)

# save figure 4 dataset
pak_fs_fig4_dataset %>%
  write_csv("./june.2022/factsheets/pakistan/graph.wise.datasets/pak_fs_fig_4_dataset.csv")
  
# pakistan factsheet figure 4
pak_fs_fig4 <- ggplot(pak_fs_fig4_dataset) +
  geom_line(mapping = ggplot2::aes(x = years, y = pop_weighted_avg_pm2.5, color = region), lwd = 1.1) +
    geom_hline(mapping = aes(yintercept = 5), lwd = 0.8) +
  scale_y_continuous(breaks = seq(0, 60, 5), limits = c(0, 60)) +
  scale_x_continuous(breaks = seq(1998, 2020, 2)) +
    scale_color_manual(values = c("National Average" = "darkgrey", "Islamabad" = "darkred", "Punjab" = "orange", "All other regions (excluding Islamabad and Punjab)" = "tan")) +
  ggthemes::theme_clean() +
  labs(x = "Years", 
       y = expression(paste("Average PM2.5 concentration ( ", mu, "g", "/", m^3, " )")), 
       caption = "*FCT refers to the Federal Capital Territory region in Pakistan") +
  theme(legend.position = "bottom", legend.title = element_blank(), 
        legend.text = element_text(size = 7), 
        axis.title.y = element_text(size = 9), 
        axis.title.x = element_text(size = 9), 
        plot.caption = element_text(hjust = 0, size = 5), plot.caption.position = "plot") +
    geom_text(x = 2002.7, y = 8, label = expression(paste("WHO PM2.5 Guideline (last updated: 2021): 5 ", mu, "g","/", m^3, "")))



```
