---
title: "Factsheets 2022"
author: "Aarsh"
date: '2022-07-26'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# get packages, global variables and datasets
```{r load-lib-helper-files-data-global-var}
  
# libraries
library(readr)
library(dplyr)
library(stringr)
library(magrittr)
library(ggplot2)
library(readr)
library(tidytext)
library(tidyr)
library(tidyverse)
library(sf)
library(usethis)
library(devtools)
library(data.table)

# read in the helper file
source("./appPublic/aqli.data.explorer.helper.script.R")


# global variables
who_guideline <- 5
le_constant <- 0.098


```


# set paths

```{r}

# set path for where risks and causes master files are stored
gbd_risks_causes_path <- "./september.2023/gbd.calculation/GBDComparisons/GBD/"

# WHO raw life tables path
who_raw_life_tables_path <- "./september.2023/gbd.calculation/GBDComparisons/WHO life tables/"

# WHO clean life tables path
who_cleaned_life_tables_path <- "./september.2023/gbd.calculation/GBDComparisons/WHO life tables/"

# sex ratio country wise file path

sr_country_wise_file_path <- "./september.2023/gbd.calculation/GBDComparisons/"

# results (without country names) folder path
results_without_country_names_folder_path <- "./september.2023/gbd.calculation/GBDComparisons/results/"

# results (with country names) folder path
results_with_country_names_folder_path <- "./september.2023/gbd.calculation/GBDComparisons/results_used_in_dashboard/"


```

# sex ratio file cleaning

```{r}
# read in sex ration country wise file
sr_country_wise <- read_csv(str_c(sr_country_wise_file_path, "sr_country_wise.csv"))

# removing unnecessary top rows
sr_country_wise <- sr_country_wise[4:nrow(sr_country_wise), ]

# assigning the first row as the column names for the dataset and then remove the first row 
colnames(sr_country_wise) <- sr_country_wise[1, ]
sr_country_wise <- sr_country_wise[2:nrow(sr_country_wise), ]

# keep only the country column and the latest year's (which has data in it, have a look at the dataset for this and then set below, can also automate this but as of now not automated) sex ration column
sr_country_wise_cleaned <- sr_country_wise %>%
  select(`Country Name`, `2020`) %>%
  rename(region = `Country Name`, 
         sr = `2020`)

# places where sex ratio is not available, assume it to be 1
sr_country_wise_cleaned[is.na(sr_country_wise_cleaned$sr), ]$sr <- 1

```


# GBD master files cleaning

```{r}


# risks and causes master files specific paths
gbd_risks_master_path <- str_c(gbd_risks_causes_path, "gbdrates_master_risks_raw.csv")
gbd_causes_master_path <- str_c(gbd_risks_causes_path, "gbdrates_master_causes_raw.csv")

# read in master risks and causes files
gbd_risks_master <- read_csv(gbd_risks_master_path)
gbd_causes_master <- read_csv(gbd_causes_master_path)

# clean causes master file and write as csv
gbd_causes_master_region_cleaned <- gbd_causes_master %>%
  mutate(age = str_remove(age, " years"), 
         age = str_replace(age, "\\+", " plus"), 
         age = str_replace(age, "-", " to "))  

gbd_causes_master_region_cleaned %>%
write_csv("./september.2023/gbd.calculation/GBDComparisons/GBD/gbdrates_master_causes_cleaned.csv")

# clean risks master file and write as csv
gbd_risks_master_region_cleaned <- gbd_risks_master %>%
  mutate(age = str_remove(age, "years"), 
         age = str_replace(age, "\\+", " plus"), 
         age = str_replace(age, "-", " to ")) 

gbd_risks_master_region_cleaned %>%
 write_csv("./september.2023/gbd.calculation/GBDComparisons/GBD/gbdrates_master_risks_cleaned.csv")


```

# creating region wise causes and risks file from master files and writing it to the GBD folder

```{r}

# split the master risks dataset by location
gbd_risks_master_region_cleaned_region_wise <- gbd_risks_master_region_cleaned %>%
  group_split(location)

# write location wise CSVs
for (i in 1 : length(gbd_risks_master_region_cleaned_region_wise)){
  risks_data_tmp <- gbd_risks_master_region_cleaned_region_wise[[i]]
  location_tmp <- risks_data_tmp$location[i]
  risks_data_tmp %>%
    write_csv(str_c(gbd_risks_causes_path, "gbdrates_", location_tmp, "_", "risks.csv"))
}


# split the master causes dataset by loaction
gbd_causes_master_region_cleaned_region_wise <- gbd_causes_master_region_cleaned %>%
  group_split(location)


# write location wise CSVs
for (i in 1 : length(gbd_causes_master_region_cleaned_region_wise)){
  causes_data_tmp <- gbd_causes_master_region_cleaned_region_wise[[i]]
  location_tmp <- causes_data_tmp$location[i]
  causes_data_tmp %>%
    write_csv(str_c(gbd_risks_causes_path, "gbdrates_", location_tmp, "_", "causes.csv"))
}

```

# Writing the cleaned version of the WHO raw life tables

```{r}
# read in WHO raw life tables 1 by 1 in a loop, clean and write the cleaned versions

who_life_tables_list <- list.files(who_raw_life_tables_path)

for (i in 1:length(who_life_tables_list)){
  
  who_raw_life_table_tmp <- read_csv(str_c(who_raw_life_tables_path, who_life_tables_list[i]))

  location_tmp <- str_remove(who_life_tables_list[i], "_raw.csv")
  
  who_raw_life_table_tmp <- who_raw_life_table_tmp[, 1:5]
  
  colnames(who_raw_life_table_tmp) <- who_raw_life_table_tmp[1, ]
  
  who_raw_life_table_tmp <- who_raw_life_table_tmp[2:nrow(who_raw_life_table_tmp), c(1, 2, 4, 5)]
  
  colnames(who_raw_life_table_tmp) <- c("indicator", "agegroup", "male", "female")
  
  cleaned_life_table_tmp <- who_raw_life_table_tmp  %>%
  mutate(lt_var = str_extract(indicator, ".+ (\\-)"), 
         lt_var = str_remove(lt_var, " -"),
    agegroup = str_replace(agegroup, "&lt;", "<"))
    

  cleaned_life_table_tmp %>%
    write_csv(str_c(who_cleaned_life_tables_path, location_tmp, ".csv"))
  
  print(str_c("Iteration # ", i, "/", length(who_life_tables_list), " complete"))

}

```


# Reading in the country names of WHO life tables (after their country names have been matched with the names of the GBD risks and causes files) and creating a tibble that contains country wise sex ratio, which will then be read by the STATA lifetable_analysis script to produce all country wise results in a loop.

```{r}

# Sanity check----------------------

# test if the first n (figure out n) characters uniquely identify a cause of death (this is needed  because the STATA script creates variable names by selecting a couple characters of the causes of death. If enough characters are not used in assigning this, than multiple diseases will have the same variable names and there is a risk of them getting assigned the same life years lost. For example: If only first 5 characters are used for creating a variable name than "Child and Maternal Malnutrition" and "Childhood abuse", both will have the abbreviation "Child" and hence that's not right. Therefore, the 'n' selected should be long enough that it uniquely identifies the variable, but also not too long to break the STATA variable name length limit). I am reading in India results file, assuming that it would mostly have all causes of deaths.

# read in a gbd causes dataset (India/Somalia read in here, assuming that it has the most cod, can do it a different way directly listing cod and risks in a column from GBD)
causes_names_test <- read_csv(str_c(gbd_risks_causes_path, "gbdrates_Somalia_causes.csv")) 
causes_names_test <- causes_names_test %>%
  select(cause) %>%
  rename(cod = cause)


# read in a gbd risks dataset (India read in here, assuming that it has the most cod, can do it a different way directly listing cod and risks in a column from GBD) and rename risks to causes
risks_names_test <-  read_csv(str_c(gbd_risks_causes_path, "gbdrates_Somalia_risks.csv")) 
risks_names_test <-  risks_names_test %>%
  rename(cod = rei) %>%
  select(cod)

# append all risks and causes together in a single column and keep only unique causes
test_dataset <- rbind(causes_names_test, risks_names_test)
test_dataset <- test_dataset %>%
  count(cod) %>%
  select(cod)

# select first n (in this case I think 12 is enough to uniquely identify cod)
foo <- test_dataset %>%
  mutate(unique_var_name = str_sub(cod, 1,  12))

# check if its actually unique
foo1 <- count(foo, unique_var_name)

foo1 %>%
  filter(n > 1)

#-------------------------------------

# list all file names in the WHO life tables directory and create a tibble out of it
who_life_tables_file_names <- list.files(who_cleaned_life_tables_path)

who_life_tables_file_names_tibble <- who_life_tables_file_names %>%
  as_tibble() %>%
  rename(country_file_name = value)

# remove the _raw files
who_life_tables_file_names_tibble_non_raw <- who_life_tables_file_names_tibble %>%
  filter(!(str_detect(country_file_name, "raw"))) %>%
  mutate(region = str_remove(country_file_name, ".csv")) %>%
  select(region)

# join the country wise cleaned tibble with country wise sex ratio
country_wise_sr_who_life_table_country_names_join <- who_life_tables_file_names_tibble_non_raw %>% 
  left_join(sr_country_wise_cleaned, by = "region")

country_wise_sr_who_life_table_country_names_join %>%
  write_csv(str_c(sr_country_wise_file_path, "sr_country_wise_cleaned.csv"))

# For those countries for which we did not get a sex ratio, assign it manually (and for those countries that do not have a sex ratio available, assign sex ratio to be 1)


#> read in the processed results, fill in the coutry names in a new column called country, row bind all datasets to create a master GBD dataset. Keep only the country, cause of death and aggregate life years lost columns and the replace "Ambient particulate matter pollution" with "PM2.5 relative to WHO guideline" and add in the AQLI country level annual average PM2.5 there.

results_file_names <- list.files(results_without_country_names_folder_path)

# creating an empty list that will contain all processed datasets
results_with_file_names_list <- list()

# country level AQLI summary
country_level_aqli <- gadm_level_summary(gadm2_aqli_2021, "country", c(2021), 10)

# clean each country's processed results (add a country and AQLI lyl column) 
for(i in 1:length(results_file_names)){
  tmp_file <- readxl::read_excel(str_c(results_without_country_names_folder_path, results_file_names[i]))
  tmp_country <- str_remove(results_file_names[i], "estimated_life_expectancy_differences_")
  tmp_country <- str_remove(tmp_country, ".xlsx")
  tmp_file <- tmp_file %>%
    select(`Cause of death`, `Estimated life expectancy difference vs. actual: aggregate`) %>%
    rename(cause_of_death = `Cause of death`, lyl = `Estimated life expectancy difference vs. actual: aggregate`) %>%
    mutate(country = tmp_country) %>%
    select(country, cause_of_death, lyl) 
  
 tmp_country_level_aqli <- country_level_aqli %>%
    filter(country == tmp_country) 
 
 # if the exact country name is NA, fill in NA in AQLI lyl number (and then will later fill this in manually: should improve this)
 if(nrow(tmp_country_level_aqli) == 0){
   tmp_file <- tmp_file %>%
     mutate(cause_of_death = ifelse(cause_of_death == "Ambient particulate matter pollution", "PM2.5 relative to WHO guideline", cause_of_death),
       lyl = ifelse(cause_of_death == "PM2.5 relative to WHO guideline", NA, lyl))
 } else if (nrow(tmp_country_level_aqli) > 0) {
   tmp_aqli_lyl_num <- tmp_country_level_aqli %>%
     select(llpp_who_2021) %>%
     unlist()
   
    tmp_file <- tmp_file %>%
     mutate(cause_of_death = ifelse(cause_of_death == "Ambient particulate matter pollution", "PM2.5 relative to WHO guideline", cause_of_death),
       lyl = ifelse(cause_of_death == "PM2.5 relative to WHO guideline", tmp_aqli_lyl_num, lyl))
 }
 
 # store the final prepped data for a given country into the list
 results_with_file_names_list[[i]] <- tmp_file

  print(str_c("Iteration # ", i, "/", length(results_file_names), " completed"))
}

# create a data frame out of the list
final_master_results <- rbindlist(results_with_file_names_list)

# coerce it to a tibble format
final_master_results <- as_tibble(final_master_results)

# check how many lyl have an NA (will be filled in manually: should come up with a better alternate solution)
final_master_results %>% filter(cause_of_death == "PM2.5 relative to WHO guideline") %>% select(country, lyl) %>% filter(is.na(lyl)) %>% View()

final_master_results %>%
    write_csv(str_c(results_with_country_names_folder_path, "gbd_results_master", ".csv"))



```

