---
title: "NCAP report series: Part 1"
author: "Aarsh"
date: '2023-03-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# get packages, global variables, datasets
```{r load-lib-helper-files-data-global-var}

# libraries
library(readr)
library(dplyr)
library(stringr)
library(magrittr)
library(ggplot2)
library(readr)
library(tidytext)
library(tidyr)
library(tidyverse)
library(sf)
library(usethis)
library(devtools)
#library(data.table)
library(stringr)
library(forcats) # for reordering bar graphs within a plot
library(cowplot) # for inset plots

# load and document to keep things updated
# devtools::load_all()
# devtools::document()

# read in latest color file
color_2021 <- read_csv("./september.2023/master.dataset/gadm2_aqli2021_vit.csv")

#> read in color map shape file for 2021
colormap_shp_2021 <- st_read("./september.2023/master.dataset/shapefiles/master_global_allyears_gadm2_with_geom_Jan192023.shp")

# GBD final results file (resulting file that we get after applying the steps in the GBD technical appendix)
# gbd_results <- read_csv("./june.2022/other.important.calculations.data/estimated_life_expectancy_differences_master_table_final.csv")

# Countries with Ambient Air Quality Standards
# aaqs_places <- read_csv("./june.2022/other.important.calculations.data/countries_list_with_aaqs.csv")
 
# read in "county for aqli pm2.5 dataset" (US technical appendix output file)
# county_for_aqli_pm2.5_dataset <- read_csv("./june.2022/other.important.calculations.data/county_pm25_foraqli_stats.csv")


# investments in NCAP data (not using)
# ncap_budget_tracker <- read_csv("./june.2022/clean.air.countdown.campaign.2023/input/ncap_budget_tracker_data.csv")

# NCAP city, districts, state list 
ncap_regions_list <- read_csv("./september.2023/ncap.report.series.part.1/ncap_cities_district_state_long_updated_March_2023.csv")

# India city level population data (124 NCAP cities + 7 non-attainment cities) and drop other_info column
ncap_city_level_population_census_2011 <- read_csv("./september.2023/ncap.report.series.part.1/india_city_pop_data_census_wikipedia.csv")

ncap_city_level_population_census_2011 <- ncap_city_level_population_census_2011 %>%
  select(-other_info)

# AQLI district level data classified as Rural/Urban using CSE's classification
# aqli_2020_district_level_rural_urban_subset <- read_csv("./june.2022/clean.air.countdown.india.campaign.2023/input.data.shared.between.graphs/aqli_2020_district_level_rural_urban_subset.csv")

# india state
india_state <- st_read("./june.2022/clean.air.countdown.india.campaign.2023/input.data.shared.between.graphs/india_state.shp")


# global variables
who_guideline <- 5
le_constant <- 0.098

# global operations
`%notin%` <- Negate(`%in%`)

# NCAP target (as of March 05, 2023): 40 percent reduction by 2026 relative to 2017
rel_to_year <- 2017
red_perc <- 40

# India national annual average PM2.5 standard (as in March, 2023)
nat_pm2.5_stan_india <- 40


```


# Create a dataset that contains all AQLI data columns but only for the NCAP districts and then assign the district pollution and life years lost values to the non-attainment cities contained in those districts. Also, replace the 2017 pollution column of the color_2020 dataset

```{r get_data_ready}

# sanity check (check to see how many city names match district names)
sum(ncap_regions_list$city == ncap_regions_list$district, na.rm =TRUE) # 67/124 54.03%

#> some cleaning given the 2021 dataset

# replace the "&" with "and" in Jammu and Kashmir and "Chandigarhs" district state from "Punjab" to "Chandigarh"
ncap_regions_list <- ncap_regions_list %>%
  mutate(state_ut = ifelse(state_ut == "Jammu & Kashmir", "Jammu and Kashmir", state_ut), 
         state_ut = ifelse((state_ut == "Punjab") & (district == "Chandigarh"), "Chandigarh", state_ut))

# create a gadm2 level 2 dataset with only those states that are part of NCAP.
color_2021_ncap_states <- color_2021 %>%
  filter(country == "India", name_1 %in% unique(ncap_regions_list$state_ut)) 

# join the ncap regions data with the states dataset to attach pm and lyl information to all non-attainment cities and do some basic cleaning to get it finalized.
ncap_analysis_dataset <- ncap_regions_list %>%
  left_join(color_2021_ncap_states, by = c("state_ut" = "name_1", "district" = "name_2")) %>%
  select(state_ut, district, city, everything()) %>%
  rename(name_1 = state_ut, name_2 = district) %>%
  select(-c(district_option2, other_1, comments)) %>%
  select(objectid_gadm2, iso_alpha3, country, name_1, name_2, everything()) 

```

# "[NO LONGER IN USE]" Getting city level population data from Census 2011

```{r eval=FALSE, include=FALSE}
census_2011 <- readxl::read_xlsx("./A-1_NO_OF_VILLAGES_TOWNS_HOUSEHOLDS_POPULATION_AND_AREA.xlsx")

# cleaning column names, making them: lowercase, spaces relaced with underscores, brackets removed
colnames(census_2011) <- census_2011[1, ]
colnames(census_2011) <- str_to_lower(colnames(census_2011))
colnames(census_2011) <- str_replace_all(colnames(census_2011), "/", "")
colnames(census_2011) <- str_replace_all(colnames(census_2011), "\\(", "") 
colnames(census_2011) <- str_replace_all(colnames(census_2011), "\\)", "") 
colnames(census_2011) <- str_replace_all(colnames(census_2011), "-", "") 
colnames(census_2011) <- str_replace_all(colnames(census_2011), "\\.", "")
colnames(census_2011) <- str_replace_all(colnames(census_2011),  " ", "_") 


# adding subcolumn names
colnames(census_2011)[7] <- "number_of_villages_inhabited" 
colnames(census_2011)[8] <- "number_of_villages_uninhabited"
colnames(census_2011)[11] <- "population_persons"
colnames(census_2011)[12] <- "population_males"
colnames(census_2011)[13] <- "population_females"
colnames(census_2011)[6] <- "total_rural_urban"
colnames(census_2011)[14] <- "area_in_sq_km"

# removing redundant column names
census_2011 <- census_2011[4:nrow(census_2011), ]
census_2011 <- as_tibble(census_2011)


# census 2011 subset
census_2011_subset <- census_2011 %>%
  filter(india_state_union_territory_district_subdistrict == "SUB-DISTRICT", 
         total_rural_urban == "Total")

# checking how many of ncap analysis dataset city names are available in census 2011 (after converting to lower case)

ncap_df_city_names <- unique(str_to_lower(ncap_analysis_dataset$city))
census_2011_names <- unique(str_to_lower(census_2011$name))
census_2011_subset_names <- unique(str_to_lower(census_2011_subset$name)) 

sum(ncap_df_city_names %in% census_2011_names) # 85
sum(ncap_df_city_names %in% census_2011_subset_names) # 61
```



# "[NO LONGER IN USE]" Getting city level population data from the Wikipedia article that links back to a Census 2011 data table (as the raw data from Census did not have all the NCAP cities population information in it)

```{r eval=FALSE, include=FALSE}

census_2011_wiki <- read_csv("./india_city_pop_data_census_wikipedia.csv")
census_2011_wiki$city <- str_to_lower(census_2011_wiki$city)
# 
# census_2011_wiki_names <- unique(str_to_lower(census_2011_wiki$city))
# 
# indices_city_pop_not_avail <- which(ncap_df_city_names %notin% census_2011_wiki_names)
# city_pop_not_avail_rows <- ncap_analysis_dataset[indices_city_pop_not_avail, ] 
# 
# sum(unique(str_to_lower(city_pop_not_avail_rows$city)) %in% census_2011_subset_names)

# adding a new city pop column after converting city names to lower case
ncap_analysis_dataset$city <- str_to_lower(ncap_analysis_dataset$city)

ncap_analysis_dataset <- ncap_analysis_dataset %>%
  left_join(census_2011_wiki %>% select(city, pop_2011), by = "city")

ncap_analysis_dataset %>%
  select(country, name_1, name_2, city, pop_2011) %>%
  write_csv("city_wise_pop_43_missing.csv")

```








# Reading in the city level population data, compiled partly (~80/131) using the city level Census 2011 population referred to in this article: https://en.wikipedia.org/wiki/List_of_cities_in_India_by_population and the rest using manual Google searches. It happens to be the case that a lot of cities have the same name as their district or tehsil and also the census raw data naming scheme/categorization seemed ambiguous (as in, for a lot of cases we were unsure if the population number refers to the city or the district). This is why, for a chunk of these, instead of cleaning the census 2011 data ourselves, we decided to refer to someone else's cleaned compilation who might have dug in a little bit more into this, compared to us. 

```{r city_level_pop_data}

# convert the city column of the ncap analysis data to lower case (to join with city level population)
ncap_analysis_dataset$city <- str_to_lower(ncap_analysis_dataset$city)

# join ncap analysis data with city level population data, using the city column
final_analysis_dataset <- ncap_analysis_dataset %>%
  left_join(ncap_city_level_population_census_2011, by = c("country", "name_1", "name_2", "city"))

# rename the aqli population column to pop_district and rename the census 2011 pop_2011_city column to  "population"

final_analysis_dataset <- final_analysis_dataset %>%
  rename(population_district = population, 
         population = pop_2011_city)

# convert the city column to title case, reorder columns and the data is ready for plotting.
final_analysis_dataset <- final_analysis_dataset %>%
  mutate(city = str_to_title(city)) %>%
  select(objectid_gadm2:city, population, population_district, everything())

# Replace Nct Of Delhi in final analysis dataset with NCT of Delhi
final_analysis_dataset <- final_analysis_dataset %>%
  mutate(city = ifelse(city == "Nct Of Delhi", "NCT of Delhi", city))


```



# Fig 1: PM2.5 concentrations in 10 largest non-attainment cities of India

```{r fig1}

# plot 1 data
plot1_data <- final_analysis_dataset %>%
  slice_max(population, n = 10) %>%
  mutate(pm2017_40_perc_red = pm2017*((100 - red_perc)/100)) %>%
  select(country, name_1, name_2, city, population, pm2017, pm2017_40_perc_red) %>%
  slice_max(population, n = 10) %>%
  pivot_longer(cols = pm2017:pm2017_40_perc_red, names_to = "pol_type", values_to = "value") 

plot1_data_with_empty_rows <- plot1_data %>%
  add_row(country = "", name_1 = "", name_2 = "", 
          city = "", population = 0, pol_type = "pm2017", value = 0) %>%
  add_row(country = "", name_1 = "", name_2 = "", 
          city = "", population = 0, pol_type = "pm2017_40_perc_red", value = 0)

# plot 1
plt1 <- plot1_data_with_empty_rows %>%
  ggplot() + 
  geom_col(mapping = aes(x = forcats::fct_reorder(city, population, .desc = TRUE), y = value, fill = pol_type), position = "dodge", alpha = 0.7) +
  geom_hline(mapping = aes(yintercept = who_guideline), color = "red", linetype = "dashed", size = 0.8) +
   geom_hline(mapping = aes(yintercept = nat_pm2.5_stan_india), color = "red", linetype = "dashed", size = 0.8) +
  labs(x = "City", y = expression("Annual average" ~ PM[2.5] ~ "(in µg/m³)"), fill = "", 
       title = expression("2017" ~ PM[2.5] ~ "Concentrations in top 10 most populated Non-Attainment cities in India"), caption = "*[Placeholder for footnote]") +
  scale_fill_manual(values = c("pm2017" = "darkblue", "pm2017_40_perc_red" = "cornflowerblue"), labels = c("2017", "After 40% reduction under NCAP")) +
  guides(colour = guide_legend(override.aes = list(shape = 12))) +
  scale_y_continuous(breaks = seq(0, 100, 10)) +
   ggthemes::theme_tufte() +
   theme(axis.text.x = element_text(hjust = 1, angle = 55, size = 9),
         axis.text.y = element_text(size = 9),
         axis.title = element_text(size = 10),
        legend.position = "bottom", 
        legend.text = element_text(size = 7), 
        panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), 
plot.title = element_text(hjust = 0.5, size = 12), 
legend.background = element_rect(fill = "white", colour = "black"), 
axis.ticks = element_blank(), 
plot.caption = element_text(hjust = 0, margin = margin(0, 10, 0, 0), face = "italic")) +
  annotate("text", x = "", y = 43, label = "NAAQS", hjust = 0, size = 2.5) +
  annotate("text", x = "", y = 8, label = expression("WHO"), hjust = 0, size = 2.5)
  
  
  

```

# Fig 2: Life Expectancy Gain from achieving 40% reduction, under NCAP in 10 largest non-attainment cities of India

```{r fig2}

# plot2 data

plot2_data <- final_analysis_dataset %>%
    slice_max(population, n = 10) %>%
  mutate(pm2017_40_perc = pm2017*((red_perc)/100), 
         le_gain_from_40_perc_red = pm2017_40_perc*0.098) %>%
  select(country, name_1, name_2, city, population, pm2017, pm2017_40_perc, le_gain_from_40_perc_red) %>%
  slice_max(population, n = 10) 

plt2 <- plot2_data %>%
  ggplot() + 
    geom_col(mapping = aes(x = forcats::fct_reorder(city, population, .desc = TRUE), y = le_gain_from_40_perc_red), fill = "darkred", width = 0.6) +
  labs(x = "City", y = "Life Expectancy Gains (in years)", title = expression("Life Expectancy Gains by reducing" ~ PM[2.5] ~ "by 40% relative to 2017"), 
       subtitle = "(Top 10 most populated non-attainment cities in India)", 
       caption = "*[Placeholder for footnote]") +
  ggthemes::theme_clean() +
  theme(axis.text.x = element_text(hjust = 1, angle = 55, size = 9),
         axis.text.y = element_text(size = 9),
         axis.title = element_text(size = 10),
        legend.position = "bottom", 
        legend.text = element_text(size = 7), 
        panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), 
plot.title = element_text(hjust = 0.5, size = 13), 
plot.subtitle = element_text(hjust = 0.5, size = 9), 
legend.background = element_rect(fill = "white", colour = "black"), 
axis.ticks = element_blank(), 
plot.caption = element_text(hjust = 0, margin = margin(0, 10, 0, 0), face = "italic"))
  
```


# Figure 3, Map: Change in PM2.5 from achieving the 40% reduction target under NCAP

```{r fig3}

# separately getting Gilgit Baltistan and Azad Kashmir
azad_kashmir_gilgit_baltistan <- color_2021 %>%
  filter(country == "Pakistan", name_1 %in% c("Azad Kashmir", "Gilgit Baltistan"))

# plot 3 data
plot3_data <- color_2021 %>%
  filter(country == "India") %>%
  bind_rows(azad_kashmir_gilgit_baltistan) %>%
  mutate(pm2017_40_perc_red = pm2017*((100 - red_perc)/100), 
         pm2017_40_perc = -(pm2017*(red_perc/100))) %>%
  left_join(colormap_shp_2021, by = c("objectid_gadm2" = "objidgadm2")) %>%
  mutate(pol_bucket = ifelse(((pm2017_40_perc > -5) & (pm2017_40_perc <= 0)), ">(-5) to 0", pm2017_40_perc), 
         pol_bucket = ifelse(((pm2017_40_perc > -10) & (pm2017_40_perc <= -5)), ">(-10) to -5", pol_bucket), 
         pol_bucket = ifelse(((pm2017_40_perc > -15) & (pm2017_40_perc <= -10)), ">(-15) to -10", pol_bucket), 
           pol_bucket = ifelse(((pm2017_40_perc > -30) & (pm2017_40_perc <= -15)), ">(-30) to -15", pol_bucket), 
           pol_bucket = ifelse(((pm2017_40_perc > -40) & (pm2017_40_perc <= -30)), ">(-40) to -30", pol_bucket), 
         pol_bucket = ifelse(((pm2017_40_perc > -50) & (pm2017_40_perc <= -40)), ">(-50) to -40", pol_bucket)) %>%
  mutate(pol_bucket_order = ifelse(pol_bucket == ">(-5) to 0", 1, 0), 
         pol_bucket_order = ifelse(pol_bucket == ">(-10) to -5", 2, pol_bucket_order), 
         pol_bucket_order = ifelse(pol_bucket == ">(-15) to -10", 3, pol_bucket_order), 
         pol_bucket_order = ifelse(pol_bucket == ">(-30) to -15", 4, pol_bucket_order), 
         pol_bucket_order = ifelse(pol_bucket == ">(-40) to -30", 5, pol_bucket_order), 
         pol_bucket_order = ifelse(pol_bucket == ">(-50) to -40", 5, pol_bucket_order)) %>%
  select(-geometry, geometry) %>%
  st_as_sf()

# plot 3  
plt3 <- plot3_data %>%
  ggplot() +
  geom_sf(mapping = aes(fill = forcats::fct_reorder(pol_bucket, pol_bucket_order)), color = "transparent") +
    geom_sf(data = india_state, color = "black", fill = "transparent") +
  ggthemes::theme_map() + 
  scale_fill_manual(values = c(">(-5) to 0" = "#CBE8F3", ">(-10) to -5" = "#B4CDD9", ">(-15) to -10" = "#8FA1AD", ">(-30) to -15" = "#707F8C", ">(-40) to -30" = "#576572", ">(-50) to -40" = "#414F5D")) +
  theme(legend.position = "bottom", 
        legend.justification = c(0.5, 3), 
        legend.background = element_rect(color = "black"), 
        legend.text = element_text(size = 7), 
        legend.title = element_text(size = 7), 
        plot.title = element_text(hjust = 0.5)) +
  labs(fill = expression("Reduction in" ~ PM[2.5] ~ "(µg/m³)"), 
       title = expression("Change in" ~ PM[2.5] ~ "from achieving 40%" ~ PM[2.5] ~ "reduction target under NCAP")) +
  guides(fill = guide_legend(nrow = 1))
  
  
  




```


# Figure 4, Map: Life Expectancy Gain from achieving the 40% reduction target under NCAP

```{r}

# plot 4 data
plot4_data <- color_2021 %>%
  filter(country == "India") %>%
  bind_rows(azad_kashmir_gilgit_baltistan) %>%
  mutate(pm2017_40_perc = (pm2017*(red_perc/100)),
         le_gain_pm2017_40_perc_red = pm2017_40_perc*0.098) %>%
  left_join(colormap_shp_2021, by = c("objectid_gadm2" = "objidgadm2")) %>%
  mutate(lyl_bucket = ifelse((le_gain_pm2017_40_perc_red < 0.1), "0 - 0.1 years", le_gain_pm2017_40_perc_red), 
         lyl_bucket = ifelse((le_gain_pm2017_40_perc_red >= 0.1) & (le_gain_pm2017_40_perc_red < 0.5), "0.1 - < 0.5", lyl_bucket), 
         lyl_bucket = ifelse((le_gain_pm2017_40_perc_red >= 0.5) & (le_gain_pm2017_40_perc_red < 1), "0.5 - < 1", lyl_bucket), 
         lyl_bucket = ifelse((le_gain_pm2017_40_perc_red >= 1) & (le_gain_pm2017_40_perc_red < 2), "1 - < 2", lyl_bucket), 
         lyl_bucket = ifelse((le_gain_pm2017_40_perc_red >= 2) & (le_gain_pm2017_40_perc_red < 3), "2 - < 3", lyl_bucket), 
         lyl_bucket = ifelse((le_gain_pm2017_40_perc_red >= 3) & (le_gain_pm2017_40_perc_red < 4), "3 - < 4", lyl_bucket), 
         lyl_bucket = ifelse((le_gain_pm2017_40_perc_red >= 4), ">= 4 years", lyl_bucket)) %>%
  mutate(lyl_bucket_order = ifelse(lyl_bucket == "0 - 0.1 years", 1, 0), 
         lyl_bucket_order = ifelse(lyl_bucket == "0.1 - < 0.5", 2, lyl_bucket_order), 
         lyl_bucket_order = ifelse(lyl_bucket == "0.5 - < 1", 3, lyl_bucket_order), 
         lyl_bucket_order = ifelse(lyl_bucket == "1 - < 2", 4, lyl_bucket_order), 
         lyl_bucket_order = ifelse(lyl_bucket == "2 - < 3", 5, lyl_bucket_order), 
         lyl_bucket_order = ifelse(lyl_bucket == "3 - < 4", 6, lyl_bucket_order), 
         lyl_bucket_order = ifelse(lyl_bucket == ">= 4 years", 7, lyl_bucket_order)) %>%
  select(-geometry, geometry) %>%
  st_as_sf()

# plot 4
plt4 <- plot4_data %>%
  ggplot() +
  geom_sf(mapping = aes(fill = forcats::fct_reorder(lyl_bucket, lyl_bucket_order)), color = "transparent") +
    geom_sf(data = india_state, color = "black", fill = "transparent") +
  ggthemes::theme_map() + 
scale_fill_manual(values = c("0 - 0.1 years" = "#FFFFFF", "0.1 - < 0.5" = "#FFE6B3", "0.5 - < 1" = "#FFD25D", 
                                "1 - < 2" = "#FFBA00", "2 - < 3" = "#FF9600", "3 - < 4" = "#FF6908", 
                                ">= 4 years" = "#E63D23")) +
  theme(legend.position = "bottom", 
        legend.justification = c(0.5, 3), 
        legend.background = element_rect(color = "black"), 
        legend.text = element_text(size = 7), 
        legend.title = element_text(size = 7), 
        plot.title = element_text(hjust = 0.5), 
        legend.key = element_rect(color = "black")) +
  labs(fill = str_wrap("Gain in Life Expectancy (in years)", 10), 
       title = expression("Life Expectancy Gains from Achieving 40% reduction" ~ PM[2.5] ~ "reduction target under NCAP"))+
  guides(fill = guide_legend(nrow = 1))


```


# Generating the appendix table for NCAP report part 1

```{r appendix_table}

# ncap analysis data
appendix_table <- final_analysis_dataset %>%
  select(-population_district) %>%
  mutate(pm2017_red_40_perc = pm2017*((100 - red_perc)/100), 
         pm2017_40_perc = pm2017*(red_perc/100), 
         le_gain_pm2017_40_perc_red = pm2017_40_perc * le_constant, 
         le_gain_pm2017_40_perc_red = round(le_gain_pm2017_40_perc_red, 1),
         population_millions = round(population/1000000, 1)) %>%
  select(city, name_1, population_millions, pm2017, pm2017_red_40_perc, le_gain_pm2017_40_perc_red) %>%
  rename(City = city, State = name_1, `City (Census 2011) Population (in millions)` = population_millions, `2017 PM2.5 Concentrations (µg/m³)` = pm2017, `2017 PM2.5 Concentrations (after 40% reduction)` = pm2017_red_40_perc, `Life Expectancy Gains (in years) from a 40% reduction` = le_gain_pm2017_40_perc_red)

appendix_table %>% 
  write_csv("./september.2023/ncap.report.series.part.1/output_files/ncap.2023.report.part1.appendix_table.csv")

```

# other calculations

```{r other_calculations}

#> gains in life years that an average Indian would experience if the NCAP 40 percent reduction target by 2026 (relative to 2017)

# calculate india average pollution in 2017 and then take 40% of it to calculate the life years lost for that reduction
final_analysis_dataset %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2017_pop_weighted = pm2017*pop_weights) %>%
  summarise(avg_pm2017_pop = sum(pm2017_pop_weighted, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(avg_pm2017_pop_40_perc = ((red_perc/100)*(avg_pm2017_pop)), 
         lyg_40_perc_red = round((avg_pm2017_pop_40_perc*le_constant), 1))

#> average gains in life expectancy by reducing pollution in top 5 most polluted non-attainment cities

# top 5 most polluted non-attainment cities 
top_5_most_pol_non_att_cities <- final_analysis_dataset %>%
  slice_max(pm2017, n = 5) %>%
  pull(city) 

final_analysis_dataset %>%
  filter(city %in% top_5_most_pol_non_att_cities) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2017_pop_weighted = pop_weights*pm2017) %>%
  summarise(avg_pm2017_pop = sum(pm2017_pop_weighted, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(avg_pm2017_pop_40_perc = ((red_perc/100)*(avg_pm2017_pop)), 
         lyg_40_perc_red = round(avg_pm2017_pop_40_perc*le_constant, 1))

#> India's average PM2.5 percentage increase since 1998
color_2021 %>%
  filter(country == "India") %>%
  group_by(country) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2017_pop_weighted = pop_weights*pm2017, 
         pm1998_pop_weighted = pop_weights*pm1998) %>%
    summarise(avg_pm2017_pop = sum(pm2017_pop_weighted, na.rm = TRUE), 
              avg_pm1998_pop = sum(pm1998_pop_weighted, na.rm = TRUE)) %>%
    ungroup() %>%
  mutate(perc_inc_1998_2017 = (((avg_pm2017_pop - avg_pm1998_pop)/avg_pm1998_pop)*100))

#> Delhi's annual average PM2.5 value in 2017 and how many life years are lost by not being compliant with WHO and National PM2.5 standard

color_2021 %>%
  filter(country == "India", name_1 == "NCT of Delhi", name_2 == "NCT of Delhi") %>%
  select(pm2017) %>%
  mutate(lyl_rel_who = (pm2017 - who_guideline)*0.098, 
         lyl_rel_nat = (pm2017 - 40)*0.098)
  
#> country ranking in pollution given 2020 data

color_2021 %>%
  group_by(country) %>%
   mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2020_pop_weighted = pop_weights*pm2020, 
         pm1998_pop_weighted = pop_weights*pm1998) %>%
  summarise(avg_pm2020_pop = sum(pm2020_pop_weighted, na.rm = TRUE), 
            avg_pm1998_pop = sum(pm1998_pop_weighted, na.rm = TRUE)) %>%
  slice_max(avg_pm2020_pop, n = 10)
  

# India's average pollution increase from 1998 to 2020
color_2021 %>%
  filter(country == "India") %>%
  group_by(country) %>%
   mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2020_pop_weighted = pop_weights*pm2020, 
         pm1998_pop_weighted = pop_weights*pm1998) %>%
  summarise(avg_pm2020_pop = sum(pm2020_pop_weighted, na.rm = TRUE), 
            avg_pm1998_pop = sum(pm1998_pop_weighted, na.rm = TRUE))

# Top 5 most polluted states in ncap

final_analysis_dataset %>%
  slice_max(pm2017, n = 15) %>%
  pull(name_1) %>%
  unique()

# state level pm2.5
final_analysis_dataset %>%
  group_by(country, name_1) %>%
   mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2020_pop_weighted = pop_weights*pm2020, 
         pm1998_pop_weighted = pop_weights*pm1998) %>%
  summarise(avg_pm2020_pop = sum(pm2020_pop_weighted, na.rm = TRUE), 
            avg_pm1998_pop = sum(pm1998_pop_weighted, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(name_1 %in% c("NCT of Delhi", "Uttar Pradesh", "Bihar", "Haryana")) %>%
  mutate(times = avg_pm2020_pop/avg_pm1998_pop)
  
# top most polluted districts in 2020
final_analysis_dataset %>%
  slice_max(pm2020) %>%
  pull(pm2020)

# total population of all NCAP cities
sum(final_analysis_dataset$population, na.rm = TRUE)/1210854977

#> assuming that each of the district's (that include the NCAP cities) 2017 pollution is reduced by 40 percent and then recalculating India's pollution average.

ncap_districts <- final_analysis_dataset %>%
  count(country, name_1, name_2) 

# india's pollution average without any reductions in 2017. # 53.34 micrograms per cubic meter in 2017
color_2021 %>%
  filter(country == "India") %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2017_pop_weighted = pop_weights*pm2017, 
         pm1998_pop_weighted = pop_weights*pm1998) %>%
  summarise(avg_pm2017_pop = sum(pm2017_pop_weighted, na.rm = TRUE), 
            avg_pm1998_pop = sum(pm1998_pop_weighted, na.rm = TRUE))

# india's pollution average with 40% reduction in each of NCAP's district (as a proxy for 40% reduction in each of 131 ncap cities). # 46.56 micrograms per cubic meter in 2017 if reduction is achieved. 46.56434

color_2021 %>%
  filter(country == "India") %>%
  mutate(pm2017_red = ifelse((name_1 %in% ncap_districts$name_1) & (name_2 %in% ncap_districts$name_2), pm2017*0.6, pm2017)) %>%
    mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2017_pop_weighted_red = pop_weights*pm2017_red, 
         pm1998_pop_weighted = pop_weights*pm1998) %>%
  summarise(avg_pm2017_pop_red = sum(pm2017_pop_weighted_red, na.rm = TRUE), 
            avg_pm1998_pop = sum(pm1998_pop_weighted, na.rm = TRUE))

# Range of gains as a result of reduction in pm2.5 by 40 percent
final_analysis_dataset %>%
  mutate(pm2017_40_perc = pm2017 * 0.4, 
         lyg_40_perc = pm2017_40_perc*0.098) %>%
  group_by(country, name_1) %>%
  summarise(min_lyg = min(lyg_40_perc, na.rm = TRUE), 
            max_lyg = max(lyg_40_perc, na.rm = TRUE)) %>% 
  slice_max(max_lyg, n = 10)

# india average pollution in 2020
color_2021 %>%
  filter(country == "India") %>%
   mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2020_pop_weighted = pop_weights*pm2020, 
         pm1998_pop_weighted = pop_weights*pm1998) %>%
  summarise(avg_pm2020_pop = sum(pm2020_pop_weighted, na.rm = TRUE), 
            avg_pm1998_pop = sum(pm1998_pop_weighted, na.rm = TRUE))
  

#> hex plot

# hex plot data
hex_plt_data <- final_analysis_dataset %>%
  mutate(pm2017_40_perc_red = pm2017*0.6) %>%
  select(country, name_1, name_2, population, pm2017, pm2017_40_perc_red) %>%
  pivot_longer(pm2017:pm2017_40_perc_red, values_to = "pollution_value", names_to = "pol_type") %>%
  mutate(pol_type = ifelse(pol_type == "pm2017", "2017 PM2.5", "40% reduction in 2017 PM2.5"))

# hex plot: plot
hex_plot <- hex_plt_data %>%
  ggplot() +
  geom_hex(mapping = aes(x = population, y = pollution_value), color = "white", alpha = 0.7) +
  theme_bw() +
  scale_x_log10() + 
  viridis::scale_fill_viridis(option = "viridis", direction = -1) + 
geom_hline(mapping = aes(yintercept = 40)) +
  # geom_hline(mapping = aes(yintercept = 5)) +
  geom_vline(mapping = aes(xintercept = 50000)) + 
  scale_y_continuous(breaks = seq(0, 120, 10)) +
  facet_wrap(~pol_type) +
  labs(x = "Population", y = "PM2.5 pollution (in µg/m³)", 
       fill = "City Count", 
       title = "Population v/s Pollution in 131 non-attainment cities",
       subtitle = "(2017 PM2.5 levels v/s if NCAP targets are met)") +
  theme(legend.title.align = 0.5, 
        plot.title = element_text(hjust = 0.5), 
        plot.subtitle = element_text(size = 8, hjust = 0.5)) 


#> creating a heatmap snapshot of the 131 non-attainment cities--------

# snapshot heatmap data
snapshot_map_data <- final_analysis_dataset %>%
  select(-pm2021) %>%
  select(country, name_1, name_2, population, city, starts_with("pm")) %>%
  pivot_longer(cols = pm1998:pm2020, names_to = "years", 
               values_to = "pop_weighted_avg_pm2.5") %>%
  mutate(years = as.integer(unlist(str_extract(years, "\\d+"))), 
         region = "City Average") 


# cities to plot in the snapshot heatmap
city_to_plot <- final_analysis_dataset %>%
  filter(pm2017 > 40) %>%
  pull(city) 

# snapshot heatmap
snapshot_heatmap <- snapshot_map_data %>%
 filter(city %in% city_to_plot, years <= 2017) %>%
  ggplot() +
  geom_raster(mapping = aes(x = years, y = forcats::fct_reorder(city, pop_weighted_avg_pm2.5), fill = pop_weighted_avg_pm2.5), color = "black", size = 2) +
  scale_fill_viridis_b(option = "rocket", direction = -1) +
  scale_x_continuous(breaks = seq(1998, 2017, 2)) +
  labs(x = "Years", y = "City", 
       title = "PM2.5 pollution from 1998 to 2017 (NCAP baseline)", 
       subtitle = "(Cities in which PM2.5 pollution was above the national standard in 2017)", 
       fill = "PM2.5 (in µg/m³)") +
  theme(plot.title = element_text(hjust = 0.5, size = 12), 
        plot.subtitle = element_text(hjust = 0.5, size = 8))
  
  
  
```
