---
title: "NCAP report series: Part 1"
author: "Aarsh"
date: '2023-03-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# get packages, global variables, datasets
```{r load-lib-helper-files-data-global-var}

# libraries
library(readr)
library(dplyr)
library(stringr)
library(magrittr)
library(ggplot2)
library(readr)
library(tidytext)
library(tidyr)
library(tidyverse)
library(sf)
library(usethis)
library(devtools)
#library(data.table)
library(stringr)
library(forcats) # for reordering bar graphs within a plot
library(cowplot) # for inset plots

# load and document to keep things updated
devtools::load_all()
devtools::document()

# read in latest color file
color_2021 <- read_csv("./september.2023/master.dataset/gadm2_aqli2021_vit.csv")

#> read in color map shape file for 2021
colormap_shp_2021 <- st_read("./september.2023/master.dataset/shapefiles/master_global_allyears_gadm2_with_geom_Jan192023.shp")

# GBD final results file (resulting file that we get after applying the steps in the GBD technical appendix)
# gbd_results <- read_csv("./june.2022/other.important.calculations.data/estimated_life_expectancy_differences_master_table_final.csv")

# Countries with Ambient Air Quality Standards
# aaqs_places <- read_csv("./june.2022/other.important.calculations.data/countries_list_with_aaqs.csv")
 
# read in "county for aqli pm2.5 dataset" (US technical appendix output file)
# county_for_aqli_pm2.5_dataset <- read_csv("./june.2022/other.important.calculations.data/county_pm25_foraqli_stats.csv")


# investments in NCAP data (not using)
# ncap_budget_tracker <- read_csv("./june.2022/clean.air.countdown.campaign.2023/input/ncap_budget_tracker_data.csv")

# NCAP city, districts, state list 
ncap_regions_list <- read_csv("./june.2022/clean.air.countdown.india.campaign.2023/input.data.shared.between.graphs/ncap_cities_district_state_long.csv")

# AQLI district level data classified as Rural/Urban using CSE's classification
# aqli_2020_district_level_rural_urban_subset <- read_csv("./june.2022/clean.air.countdown.india.campaign.2023/input.data.shared.between.graphs/aqli_2020_district_level_rural_urban_subset.csv")

# india state
india_state <- st_read("./june.2022/clean.air.countdown.india.campaign.2023/input.data.shared.between.graphs/india_state.shp")


# global variables
who_guideline <- 5
le_constant <- 0.098

# global operations
`%notin%` <- Negate(`%in%`)

# NCAP target (as of March 05, 2023): 40 percent reduction by 2026 relative to 2017
rel_to_year <- 2017
red_perc <- 40

# India national annual average PM2.5 standard (as in March, 2023)
nat_pm2.5_stan_india <- 40


```


# Create a dataset that contains all AQLI data columns but only for the NCAP districts and then assign the district pollution and life years lost values to the non-attainment cities contained in those districts. Also, replace the 2017 pollution column of the color_2020 dataset

```{r get_data_ready}

# sanity check (check to see how many city names match district names)
sum(ncap_regions_list$city == ncap_regions_list$district, na.rm =TRUE) # 67/124 54.03%

#> some cleaning given the 2021 dataset

# replace the "&" with "and" in Jammu and Kashmir and "Chandigarhs" district state from "Punjab" to "Chandigarh"
ncap_regions_list <- ncap_regions_list %>%
  mutate(state_ut = ifelse(state_ut == "Jammu & Kashmir", "Jammu and Kashmir", state_ut), 
         state_ut = ifelse((state_ut == "Punjab") & (district == "Chandigarh"), "Chandigarh", state_ut))

# create a gadm2 level 2 dataset with only those states that are part of NCAP.
color_2021_ncap_states <- color_2021 %>%
  filter(country == "India", name_1 %in% unique(ncap_regions_list$state_ut)) 

# join the ncap regions data with the states dataset to attach pm and lyl information to all non-attainment cities and do some basic cleaning to get it finalized.
ncap_analysis_dataset <- ncap_regions_list %>%
  left_join(color_2021_ncap_states, by = c("state_ut" = "name_1", "district" = "name_2")) %>%
  select(state_ut, district, city, everything()) %>%
  rename(name_1 = state_ut, name_2 = district) %>%
  select(-c(district_option2, other_1, comments)) %>%
  select(objectid_gadm2, iso_alpha3, country, name_1, name_2, everything()) 

```

# Getting city level population data from Census 2011

```{r}
census_2011 <- readxl::read_xlsx("./A-1_NO_OF_VILLAGES_TOWNS_HOUSEHOLDS_POPULATION_AND_AREA.xlsx")

# cleaning column names, making them: lowercase, spaces relaced with underscores, brackets removed
colnames(census_2011) <- census_2011[1, ]
colnames(census_2011) <- str_to_lower(colnames(census_2011))
colnames(census_2011) <- str_replace_all(colnames(census_2011), "/", "")
colnames(census_2011) <- str_replace_all(colnames(census_2011), "\\(", "") 
colnames(census_2011) <- str_replace_all(colnames(census_2011), "\\)", "") 
colnames(census_2011) <- str_replace_all(colnames(census_2011), "-", "") 
colnames(census_2011) <- str_replace_all(colnames(census_2011), "\\.", "")
colnames(census_2011) <- str_replace_all(colnames(census_2011),  " ", "_") 


# adding subcolumn names
colnames(census_2011)[7] <- "number_of_villages_inhabited" 
colnames(census_2011)[8] <- "number_of_villages_uninhabited"
colnames(census_2011)[11] <- "population_persons"
colnames(census_2011)[12] <- "population_males"
colnames(census_2011)[13] <- "population_females"
colnames(census_2011)[6] <- "total_rural_urban"
colnames(census_2011)[14] <- "area_in_sq_km"

# removing redundant column names
census_2011 <- census_2011[4:nrow(census_2011), ]

# census 2011 subset
census_2011_subset <- census_2011 %>%
  as_tibble() %>%
  filter(india_state_union_territory_district_subdistrict == "SUB-DISTRICT", 
         total_rural_urban == "Total")


```


# Fig 1: PM2.5 concentrations in 10 largest non-attainment cities of India

```{r fig1}

# plot 1 data
plot1_data <- ncap_analysis_dataset %>%
  slice_max(population, n = 10) %>%
  mutate(pm2017_40_perc_red = pm2017*((100 - red_perc)/100)) %>%
  select(country, name_1, name_2, city, population, pm2017, pm2017_40_perc_red) %>%
  slice_max(population, n = 10) %>%
  pivot_longer(cols = pm2017:pm2017_40_perc_red, names_to = "pol_type", values_to = "value") 

plot1_data_with_empty_rows <- plot1_data %>%
  add_row(country = "", name_1 = "", name_2 = "", 
          city = "", population = 0, pol_type = "pm2017", value = 0) %>%
  add_row(country = "", name_1 = "", name_2 = "", 
          city = "", population = 0, pol_type = "pm2017_40_perc_red", value = 0)

# plot 1
foo <- plot1_data_with_empty_rows %>%
  ggplot() + 
  geom_col(mapping = aes(x = forcats::fct_reorder(city, population, .desc = TRUE), y = value, fill = pol_type), position = "dodge", alpha = 0.7) +
  geom_hline(mapping = aes(yintercept = who_guideline), color = "red", linetype = "dashed", size = 0.8) +
   geom_hline(mapping = aes(yintercept = nat_pm2.5_stan_india), color = "red", linetype = "dashed", size = 0.8) +
  labs(x = "City", y = expression("Annual average" ~ PM[2.5] ~ "(in µg/m³)"), fill = "", 
       title = expression("2017" ~ PM[2.5] ~ "Concentrations in top 10 most populated Non-Attainment cities in India"), caption = "*[Placeholder for footnote]") +
  scale_fill_manual(values = c("pm2017" = "darkblue", "pm2017_40_perc_red" = "cornflowerblue"), labels = c("2017", "After 40% reduction under NCAP")) +
  guides(colour = guide_legend(override.aes = list(shape = 12))) +
  scale_y_continuous(breaks = seq(0, 100, 10)) +
   ggthemes::theme_tufte() +
   theme(axis.text.x = element_text(hjust = 1, angle = 55, size = 9),
         axis.text.y = element_text(size = 9),
         axis.title = element_text(size = 10),
        legend.position = "bottom", 
        legend.text = element_text(size = 7), 
        panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), 
plot.title = element_text(hjust = 0.5, size = 12), 
legend.background = element_rect(fill = "white", colour = "black"), 
axis.ticks = element_blank(), 
plot.caption = element_text(hjust = 0, margin = margin(0, 10, 0, 0), face = "italic")) +
  annotate("text", x = "", y = 43, label = "NAAQS", hjust = 0, size = 2.5) +
  annotate("text", x = "", y = 8, label = expression("WHO"), hjust = 0, size = 2.5)
  
  
  

```

# Fig 2: Life Expectancy Gain from achieving 40% reduction, under NCAP in 10 largest non-attainment cities of India

```{r fig2}

# plot2 data

plot2_data <- ncap_analysis_dataset %>%
    slice_max(population, n = 10) %>%
  mutate(pm2017_40_perc = pm2017*((red_perc)/100), 
         le_gain_from_40_perc_red = pm2017_40_perc*0.098) %>%
  select(country, name_1, name_2, city, population, pm2017, pm2017_40_perc, le_gain_from_40_perc_red) %>%
  slice_max(population, n = 10) 

foo <- plot2_data %>%
  ggplot() + 
    geom_col(mapping = aes(x = forcats::fct_reorder(city, population, .desc = TRUE), y = le_gain_from_40_perc_red), fill = "darkred", width = 0.6) +
  labs(x = "City", y = "Life Expectancy Gains (in years)", title = expression("Life Expectancy Gains by reducing" ~ PM[2.5] ~ "by 40% relative to 2017"), 
       subtitle = "(Top 10 most populated non-attainment cities in India)", 
       caption = "*[Placeholder for footnote]") +
  ggthemes::theme_clean() +
  theme(axis.text.x = element_text(hjust = 1, angle = 55, size = 9),
         axis.text.y = element_text(size = 9),
         axis.title = element_text(size = 10),
        legend.position = "bottom", 
        legend.text = element_text(size = 7), 
        panel.border = element_blank(), panel.grid.major = element_blank(),
panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), 
plot.title = element_text(hjust = 0.5, size = 13), 
plot.subtitle = element_text(hjust = 0.5, size = 9), 
legend.background = element_rect(fill = "white", colour = "black"), 
axis.ticks = element_blank(), 
plot.caption = element_text(hjust = 0, margin = margin(0, 10, 0, 0), face = "italic"))
  
```


# Figure 3, Map: Change in PM2.5 from achieving the 40% reduction target under NCAP

```{r fig3}

# separately getting Gilgit Baltistan and Azad Kashmir
azad_kashmir_gilgit_baltistan <- color_2021 %>%
  filter(country == "Pakistan", name_1 %in% c("Azad Kashmir", "Gilgit Baltistan"))

# plot 3 data
plot3_data <- color_2021 %>%
  filter(country == "India") %>%
  bind_rows(azad_kashmir_gilgit_baltistan) %>%
  mutate(pm2017_40_perc_red = pm2017*((100 - red_perc)/100), 
         pm2017_40_perc = -(pm2017*(red_perc/100))) %>%
  left_join(colormap_shp_2021, by = c("objectid_gadm2" = "objidgadm2")) %>%
  mutate(pol_bucket = ifelse(((pm2017_40_perc > -5) & (pm2017_40_perc <= 0)), ">(-5) to 0", pm2017_40_perc), 
         pol_bucket = ifelse(((pm2017_40_perc > -10) & (pm2017_40_perc <= -5)), ">(-10) to -5", pol_bucket), 
         pol_bucket = ifelse(((pm2017_40_perc > -15) & (pm2017_40_perc <= -10)), ">(-15) to -10", pol_bucket), 
           pol_bucket = ifelse(((pm2017_40_perc > -30) & (pm2017_40_perc <= -15)), ">(-30) to -15", pol_bucket), 
           pol_bucket = ifelse(((pm2017_40_perc > -40) & (pm2017_40_perc <= -30)), ">(-40) to -30", pol_bucket), 
         pol_bucket = ifelse(((pm2017_40_perc > -50) & (pm2017_40_perc <= -40)), ">(-50) to -40", pol_bucket)) %>%
  mutate(pol_bucket_order = ifelse(pol_bucket == ">(-5) to 0", 1, 0), 
         pol_bucket_order = ifelse(pol_bucket == ">(-10) to -5", 2, pol_bucket_order), 
         pol_bucket_order = ifelse(pol_bucket == ">(-15) to -10", 3, pol_bucket_order), 
         pol_bucket_order = ifelse(pol_bucket == ">(-30) to -15", 4, pol_bucket_order), 
         pol_bucket_order = ifelse(pol_bucket == ">(-40) to -30", 5, pol_bucket_order), 
         pol_bucket_order = ifelse(pol_bucket == ">(-50) to -40", 5, pol_bucket_order)) %>%
  select(-geometry, geometry) %>%
  st_as_sf()

# plot 3  
foo <- plot3_data %>%
  ggplot() +
  geom_sf(mapping = aes(fill = forcats::fct_reorder(pol_bucket, pol_bucket_order)), color = "transparent") +
    geom_sf(data = india_state, color = "black", fill = "transparent") +
  ggthemes::theme_map() + 
  scale_fill_manual(values = c(">(-5) to 0" = "#CBE8F3", ">(-10) to -5" = "#B4CDD9", ">(-15) to -10" = "#8FA1AD", ">(-30) to -15" = "#707F8C", ">(-40) to -30" = "#576572", ">(-50) to -40" = "#414F5D")) +
  theme(legend.position = "bottom", 
        legend.justification = c(0.5, 3), 
        legend.background = element_rect(color = "black"), 
        legend.text = element_text(size = 7), 
        legend.title = element_text(size = 7), 
        plot.title = element_text(hjust = 0.5)) +
  labs(fill = expression("Reduction in" ~ PM[2.5] ~ "(µg/m³)"), 
       title = expression("Change in" ~ PM[2.5] ~ "from achieving 40%" ~ PM[2.5] ~ "reduction target under NCAP")) +
  guides(fill = guide_legend(nrow = 1))
  
  
  




```


# Figure 4, Map: Life Expectancy Gain from achieving the 40% reduction target under NCAP

```{r}

# plot 4 data
plot4_data <- color_2021 %>%
  filter(country == "India") %>%
  bind_rows(azad_kashmir_gilgit_baltistan) %>%
  mutate(pm2017_40_perc = (pm2017*(red_perc/100)),
         le_gain_pm2017_40_perc_red = pm2017_40_perc*0.098) %>%
  left_join(colormap_shp_2021, by = c("objectid_gadm2" = "objidgadm2")) %>%
  mutate(lyl_bucket = ifelse((le_gain_pm2017_40_perc_red < 0.1), "0 - 0.1 years", le_gain_pm2017_40_perc_red), 
         lyl_bucket = ifelse((le_gain_pm2017_40_perc_red >= 0.1) & (le_gain_pm2017_40_perc_red < 0.5), "0.1 - < 0.5", lyl_bucket), 
         lyl_bucket = ifelse((le_gain_pm2017_40_perc_red >= 0.5) & (le_gain_pm2017_40_perc_red < 1), "0.5 - < 1", lyl_bucket), 
         lyl_bucket = ifelse((le_gain_pm2017_40_perc_red >= 1) & (le_gain_pm2017_40_perc_red < 2), "1 - < 2", lyl_bucket), 
         lyl_bucket = ifelse((le_gain_pm2017_40_perc_red >= 2) & (le_gain_pm2017_40_perc_red < 3), "2 - < 3", lyl_bucket), 
         lyl_bucket = ifelse((le_gain_pm2017_40_perc_red >= 3) & (le_gain_pm2017_40_perc_red < 4), "3 - < 4", lyl_bucket), 
         lyl_bucket = ifelse((le_gain_pm2017_40_perc_red >= 4), ">= 4 years", lyl_bucket)) %>%
  mutate(lyl_bucket_order = ifelse(lyl_bucket == "0 - 0.1 years", 1, 0), 
         lyl_bucket_order = ifelse(lyl_bucket == "0.1 - < 0.5", 2, lyl_bucket_order), 
         lyl_bucket_order = ifelse(lyl_bucket == "0.5 - < 1", 3, lyl_bucket_order), 
         lyl_bucket_order = ifelse(lyl_bucket == "1 - < 2", 4, lyl_bucket_order), 
         lyl_bucket_order = ifelse(lyl_bucket == "2 - < 3", 5, lyl_bucket_order), 
         lyl_bucket_order = ifelse(lyl_bucket == "3 - < 4", 6, lyl_bucket_order), 
         lyl_bucket_order = ifelse(lyl_bucket == ">= 4 years", 7, lyl_bucket_order)) %>%
  select(-geometry, geometry) %>%
  st_as_sf()

# plot 4
foo <- plot4_data %>%
  ggplot() +
  geom_sf(mapping = aes(fill = forcats::fct_reorder(lyl_bucket, lyl_bucket_order)), color = "transparent") +
    geom_sf(data = india_state, color = "black", fill = "transparent") +
  ggthemes::theme_map() + 
scale_fill_manual(values = c("0 - 0.1 years" = "#FFFFFF", "0.1 - < 0.5" = "#FFE6B3", "0.5 - < 1" = "#FFD25D", 
                                "1 - < 2" = "#FFBA00", "2 - < 3" = "#FF9600", "3 - < 4" = "#FF6908", 
                                ">= 4 years" = "#E63D23")) +
  theme(legend.position = "bottom", 
        legend.justification = c(0.5, 3), 
        legend.background = element_rect(color = "black"), 
        legend.text = element_text(size = 7), 
        legend.title = element_text(size = 7), 
        plot.title = element_text(hjust = 0.5), 
        legend.key = element_rect(color = "black")) +
  labs(fill = str_wrap("Gain in Life Expectancy (in years)", 10), 
       title = expression("Life Expectancy Gains from Achieving 40% reduction" ~ PM[2.5] ~ "reduction target under NCAP"))+
  guides(fill = guide_legend(nrow = 1))


```


# Generating the appendix table for NCAP report part 1

```{r appendix_table}

# ncap analysis data
appendix_table <- ncap_analysis_dataset %>%
  mutate(pm2017_red_40_perc = pm2017*((100 - red_perc)/100), 
         pm2017_40_perc = pm2017*(red_perc/100), 
         le_gain_pm2017_40_perc_red = pm2017_40_perc * le_constant, 
         le_gain_pm2017_40_perc_red = round(le_gain_pm2017_40_perc_red, 1),
         population_millions = round(population/1000000, 1)) %>%
  select(city, name_1, population_millions, pm2017, pm2017_red_40_perc, le_gain_pm2017_40_perc_red) %>%
  rename(City = city, State = name_1, `Population (millions)` = population_millions, `2017 PM2.5 Concentrations (µg/m³)` = pm2017, `2017 PM2.5 Concentrations (after 40% reduction)` = pm2017_red_40_perc, `Life Expectancy Gains (in years) from a 40% reduction` = le_gain_pm2017_40_perc_red)

appendix_table %>% 
  write_csv("./ncap.2023.report.part1.appendix_table.csv")

```
