---
title: "annual.report.june.2022"
author: "Aarsh"
date: '2022-07-12'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# get packages, global variables, datasets
```{r load-lib-helper-files-data-global-var}

# libraries
library(readr)
library(dplyr)
library(stringr)
library(magrittr)
library(ggplot2)
library(readr)
library(tidytext)
library(tidyr)
library(tidyverse)
library(sf)
library(usethis)
library(devtools)

# load and document to keep things updated
devtools::load_all()
devtools::document()

# read in latest color file
color_2020 <- read_csv("./master.dataset/color_2020.csv")

# read in old available color datasets
color_2019 <- read_csv("./master.dataset/color_2019.csv")
color_2016 <- read_csv("./master.dataset/color_2016.csv")

# GBD final results file (resulting file that we get after applying the steps in the GBD technical appendix)
gbd_results <- read_csv("./other.important.calculations.data/estimated_life_expectancy_differences_master_table_final.csv")

# global variables
who_guideline <- 5
le_constant <- 0.098

```



# Annual Report Figures

# Figure 1: Impact of the Revised WHO Guideline on Polluted Regions in the World

```{r ar-fig-1}

# dividing color dataset into 3 pollution groups: [0, 5], (5, 10], (10, inf)

ar_fig1_dataset <- color_2020 %>%
  mutate(region = ifelse(pm2020 >= 0 & pm2020 <= 5, "In Compliance with new WHO Guideline", pm2020), 
         region = ifelse((pm2020 > 5 & pm2020 <= 10), "Newly out of compliance", region),
         region = ifelse((pm2020 > 10), "Already out of compliance", region))


# read in color map shape file
colormap_shp <- st_read("./other.important.calculations.data/color.map.shape.file/colormap.shp")

# ar figure 1 dataset
ar_fig1_dataset <- ar_fig1_dataset %>%
  left_join(colormap_shp, by = c("country" = "NAME_0", "name_1" = "NAME_1", "name_2" = "NAME_2")) %>%
  select(country, name_1, name_2, pm2020, region, geometry)

# save annual report figure 1 dataset
ar_fig1_dataset %>%
  st_write("./annual.report/figure1.data.graph/ar_fig1_dataset.shp")

# plot ar figure 1
ar_fig1 <- ar_fig1_dataset %>%
  st_as_sf() %>%
  ggplot() +
  geom_sf(mapping = aes(fill = region))

# image saved separately by taking a screenshot from high definiton monitor

```

# Figure 2: Global Trends in PM2.5 Concentrations, 2000-2020

```{r ar-fig-2}

# calculating population weights and replacing the "pm" columns with their population weighted counterparts
color_2020_pop_weighted <- color_2020 %>%
  dplyr::mutate(global_pop_weights = population/sum(population)) %>%
  dplyr::mutate(across(starts_with("pm"), ~.x*global_pop_weights, .names = "{col}_weighted"))

# get year wise population weighted average in wide format
color_2020_summary <- color_2020_pop_weighted %>%
  dplyr::summarise(across(ends_with("weighted"), sum))

# convert the above data to long format
global_trends_figure_2020_dataset <- tibble(years = 1998:2020, global_avg_pm2.5_2020_dataset = as.numeric(unlist(color_2020_summary)))

# add le gains column
global_trends_figure_2020_dataset <- global_trends_figure_2020_dataset %>%
  dplyr::mutate(le_gains_red_pm2.5_to_who = round((global_avg_pm2.5_2020_dataset -  who_guideline)*le_constant, 1), 
         le_gains_red_pm2.5_to_who = ifelse(le_gains_red_pm2.5_to_who < 0, 0, le_gains_red_pm2.5_to_who))

# filter data and only keep the following year range: 2000 to 2020
global_trends_figure_2020_dataset_filtered <- global_trends_figure_2020_dataset %>%
  dplyr::filter(years > 1999)

# plot the figure
ar_fig2 <- ggplot() +
  geom_line(global_trends_figure_2020_dataset_filtered, mapping = aes(x = years, y = global_avg_pm2.5_2020_dataset), color = "darkblue", lwd = 1.3) +
  theme_classic() +
  labs(x = "Years", y = expression(paste("Average PM2.5 concentration ( ", mu, "gm", "/", m^3, " )"))) + 
  scale_x_continuous(breaks = seq(2000, 2020, 2)) +
  scale_y_continuous(breaks = seq(0, 35, 5), limits = c(0, 35)) +
  theme(legend.position = "none", legend.title = element_blank(), axis.title.x =  element_text(size = 8), axis.title.y = element_text(size = 10))

# save figure
ggsave("./annual.report/figure2.data.graph/ar_fig2.png", ar_fig2)

# save figure data
write_csv(global_trends_figure_2020_dataset_filtered, "./annual.report/figure2.data.graph/ar_fig2_data.csv")

```

# Figure 3: Potential Gain in Life Expectancy from Permanently Reducing PM2.5 from 2020 Concentrations to the WHO Guideline in the 10 Most Populated Countries in the World

```{r ar-fig-3}

# Creating the master dataset with all relevant columns to plot the 3 figures in the panel (will be arranged in a single row)
ar_fig3_dataset <- color_2020 %>%
  group_by(country) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2020_pop_weighted = pm2020*pop_weights, 
         pm2019_pop_weighted = pm2019*pop_weights,
         le_gain_pre = ((pm2020 - who_guideline)*le_constant), 
         le_gain_pre = ifelse(le_gain_pre < 0, 0, le_gain_pre), 
         lifeyears_gained = le_gain_pre * population) %>%
  summarise(tot_population = sum(population, na.rm = TRUE),
            tot_population_millions = round((tot_population/1000000), 2),
            avg_pm2.5_2019 = sum(pm2019_pop_weighted, na.rm = TRUE),
            avg_pm2.5_2020 = sum(pm2020_pop_weighted, na.rm = TRUE), 
            avg_life_exp_gain = round((avg_pm2.5_2020 - who_guideline)*le_constant, 2),
            avg_life_exp_gain = ifelse(avg_life_exp_gain < 0, 0, avg_life_exp_gain),
            total_person_years_gained = sum(lifeyears_gained, na.rm = TRUE), 
            total_person_years_gained_billions = round(total_person_years_gained/1000000000, 2)) %>%
  ungroup() %>%
  slice_max(tot_population, n = 10) 

# population plot of the top 10 most populous
plt_top10_most_populous_population <- ar_fig3_dataset %>%
  ggplot() +
  geom_col(mapping = aes(x = forcats::fct_reorder(country, tot_population_millions), y = tot_population_millions), fill = "darkred", width = 0.7) +
  labs(x = "Country", y = "Population (in millions)") +
  scale_y_continuous(breaks = seq(0, 1500, 200)) +
  coord_flip() +
  ggthemes::theme_tufte()

# Average Life Expectancy Gains experienced by a person in top 10 most populous countries if PM2.5 is reduced to the WHO guideline
plt_top10_most_populous_le_gains_per_person <- ar_fig3_dataset %>%
  ggplot() +
  geom_col(mapping = aes(x = forcats::fct_reorder(country, tot_population_millions), y = avg_life_exp_gain), fill = "darkred", width = 0.7) +
  labs(x = "", y = "Average Life Expectancy Gains (Years)") +
  scale_y_continuous(breaks = seq(0, 8, 1)) +
   ggthemes::theme_tufte() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  coord_flip() 

# Total person years gained on the country level for the top 10 most populous regions in the world (if PM2.5 is reduced to the WHO guideline)

plt_top10_most_populous_tot_person_years_gained <- ar_fig3_dataset %>%
  ggplot() +
  geom_col(mapping = aes(x = forcats::fct_reorder(country, tot_population_millions), y = total_person_years_gained_billions), fill = "darkred", width = 0.7) +
  labs(x = "", y = "Total Person Years Gained (Billion Years)") +
  scale_y_continuous(breaks = seq(0, 8, 1)) +
   ggthemes::theme_tufte() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  coord_flip() 

# Combine all 3 graphs in a panel
ar_fig3 <- gridExtra::grid.arrange(plt_top10_most_populous_population , plt_top10_most_populous_le_gains_per_person, plt_top10_most_populous_tot_person_years_gained, nrow = 1)

# save the plot in the appropriate folder
ggsave("./annual.report/figure3.data.graph/ar_fig3.png", ar_fig3)

# save the dataset that was reduced to generate the figure
write_csv(ar_fig3_dataset , "./annual.report/figure3.data.graph/ar_fig3_data.csv")
```


# Figure 4: Life Expectancy Impact of PM2.5 and Unassociated Causes/Risks of Death, Global

```{r ar-fig-4}

# prepare the gbd figure dataset keeping relevant causes of death and filtering out others
ar_fig4_dataset <- gbd_results %>%
  filter(location == "global", 
         cause_of_death %in% c("PM2.5 relative to WHO", "Smoking", 
                               "Alcohol use", "Unsafe water, sanitation, and handwashing",
                               "Road injuries", "HIV/AIDS", 
                               "Malaria", "Conflict and terrorism")) %>%
  select(cause_of_death, est_le_diff_vs_actual_aggregate) 

# plot the gbd figure
ar_fig4 <- ggplot(gbd_fig_4_dataset, mapping = aes(x = fct_reorder(cause_of_death, est_le_diff_vs_actual_aggregate, .desc = TRUE), y = est_le_diff_vs_actual_aggregate)) + 
  geom_col(mapping = aes(fill = est_le_diff_vs_actual_aggregate), width = 0.5) +
  theme_classic() + 
  labs(x = "", y = "Life Years Lost") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=0.90), 
        legend.position = "none") +
  ggtitle(str_c("Life Expectany Impact of Mortality Causes & Risks:", "Global", sep = " ")) +
   scale_fill_gradient(low = "darkgoldenrod1", high = "darkorange")
 
# save the plot in the appropriate folder
ggsave("./annual.report/figure4.data.graph/ar_fig4.png", ar_fig4)

# save the dataset that was reduced to generate the figure
write_csv(ar_fig4_dataset , "./annual.report/figure4.data.graph/ar_fig4_data.csv")

 
```


# Figure 5: Global Trends in PM2.5 Concentrations, 2000-2020

```{r ar-fig-5}
# AR Figure 5
 
# South Asia definition
south_asia_def <- c("Afghanistan", "Bangladesh", 
                    "Bhutan", "India", 
                    "Maldives", "Nepal", 
                    "Pakistan", "Sri Lanka")

# create an identifer for South Asia  
color_2020 <- color_2020 %>%
  mutate(region = ifelse(country %in% south_asia_def, "South Asia", country), 
         region = ifelse(region %in% "China", "China", region), 
         region = ifelse(region %in% c("China", "South Asia"), region, "Rest of the World"))


# convert data from wode to long
color_2020_long <- color_2020 %>%
    group_by(region) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         mutate(across(starts_with("pm"), ~.x*pop_weights, .names = "{col}_weighted"))) %>%
  summarise(across(ends_with("weighted"), sum)) %>%
  pivot_longer(cols = pm1998_weighted:pm2020_weighted, names_to = "years", 
               values_to = "pop_weighted_avg_pm2.5") %>%
  mutate(years = as.integer(unlist(str_extract(years, "\\d+")))) 

# keep only years post 1999
color_2020_long_2000_2020 <- color_2020_long %>%
  filter(years > 1999) 

# Save AR figure 5 dataset
color_2020_long_2000_2020 %>%
  write_csv("./annual.report/figure5.data.graph/ar_fig5_data.csv")
  

# AR figure 5 plot
ar_fig5 <- ggplot(color_2020_long_2000_2020) +  geom_line(mapping = ggplot2::aes(x = years, y = pop_weighted_avg_pm2.5, color = region, lwd = 1.5), lwd = 2) +
  labs(x = "Years", y = "Average PM2.5 concentrations (in µg/m³)") +
  ggthemes::theme_fivethirtyeight() +
  scale_color_manual(values = c("chartreuse4", "blue4", "darkorange2")) +
   theme(legend.position = "bottom", legend.title = element_blank(), 
        legend.text = element_text(size = 7), 
        axis.title.y = element_text(size = 9), 
        axis.title.x = element_text(size = 9)) +
  scale_y_continuous(breaks = seq(0, 60, 10)) +
  scale_x_continuous(breaks = seq(2000, 2020, 1))

# save the plot in the appropriate folder
ggsave("./annual.report/figure5.data.graph/ar_fig5.png", ar_fig5)


```


# Figure 6: Distribution of Person-Years Gained if PM2.5 is Reduced to WHO Guideline Around the World

```{r ar-fig6}

# oecd countries
oecd_countries <- c("Australia", "Austria", "Belgium",
"Canada", "Chile", "Colombia", "Costa Rica", "Czech Republic", "Denmark",
"Estonia", "Finland", "France", "Germany", "Greece", "Hungary",
"Iceland", "Ireland", "Israel", "Italy", "Japan", "North Korea", "South Korea",
"Latvia", "Lithuania", "Luxembourg", "Mexico", "Netherlands",
"New Zealand", "Norway", "Poland", "Portugal", "Slovakia", "Slovenia",
"Spain", "Sweden", "Switzerland", "Turkey", "United Kingdom", "United States")

# add a le_gain and le_gain_grp column in the color_2020 dataset, this is to group regions into LE gains buckets
ar_fig6_dataset <- color_2020 %>%
  mutate(le_gain = (pm2020 - who_guideline)*le_constant, 
         le_gain = ifelse(le_gain < 0, 0, le_gain), 
         le_gain_grp = ifelse(le_gain == 0, "0", le_gain), 
         le_gain_grp = ifelse(le_gain > 0 & le_gain <= 1, ">0-1", le_gain_grp), 
         le_gain_grp = ifelse(le_gain > 1 & le_gain <= 2, ">1-2", le_gain_grp), 
         le_gain_grp = ifelse(le_gain > 2 & le_gain <= 3, ">2-3", le_gain_grp), 
         le_gain_grp = ifelse(le_gain > 3 & le_gain <= 4, ">3-4", le_gain_grp),
         le_gain_grp = ifelse(le_gain > 4 & le_gain <= 5, ">4-5", le_gain_grp),
         le_gain_grp = ifelse(le_gain > 5 & le_gain <= 6, ">5-6", le_gain_grp), 
         le_gain_grp = ifelse(le_gain > 6 & le_gain <= 7, ">6-7", le_gain_grp), 
         le_gain_grp = ifelse(le_gain > 7 & le_gain <= 8, ">7-8", le_gain_grp),
         le_gain_grp = ifelse(le_gain > 8 & le_gain <= 9, ">8-9", le_gain_grp),
         le_gain_grp = ifelse(le_gain > 9, ">9", le_gain_grp)) %>%
  mutate(country_grp = ifelse(country %in% oecd_countries, "OECD countries", country), 
         country_grp = ifelse(country_grp %in% c("OECD countries", "India", "China", "Bangladesh", "Pakistan", "Indonesia"), 
                              country_grp, "Rest of the World")) %>%
  select(country, name_1, name_2, population, pm2020, le_gain_grp, country_grp) 

# add a ordering column that will help us arrange the le_gain buckets in the final figure
ar_fig6_dataset <- ar_fig6_dataset %>%
  group_by(le_gain_grp, country_grp) %>%
  summarise(tot_population = sum(population, na.rm = TRUE), 
            tot_population_millions = tot_population/1000000) %>%
  ungroup() %>%
  mutate(le_gain_grp_order = ifelse(le_gain_grp == "0", 1, 0), 
         le_gain_grp_order = ifelse(le_gain_grp == ">0-1", 2, le_gain_grp_order), 
         le_gain_grp_order = ifelse(le_gain_grp == ">1-2", 3, le_gain_grp_order), 
         le_gain_grp_order = ifelse(le_gain_grp == ">2-3", 4, le_gain_grp_order), 
         le_gain_grp_order = ifelse(le_gain_grp == ">3-4", 5, le_gain_grp_order), 
         le_gain_grp_order = ifelse(le_gain_grp == ">4-5", 6, le_gain_grp_order), 
         le_gain_grp_order = ifelse(le_gain_grp == ">5-6", 7, le_gain_grp_order), 
         le_gain_grp_order = ifelse(le_gain_grp == ">6-7", 8, le_gain_grp_order), 
         le_gain_grp_order = ifelse(le_gain_grp == ">7-8", 9, le_gain_grp_order), 
         le_gain_grp_order = ifelse(le_gain_grp == ">8-9", 10, le_gain_grp_order), 
         le_gain_grp_order = ifelse(le_gain_grp == ">9", 11, le_gain_grp_order))

# changing the ordering column type to type numeric
ar_fig6_dataset$le_gain_grp_order <- as.numeric(ar_fig6_dataset$le_gain_grp_order)

# save annual report figure 6 dataset
ar_fig6_dataset %>%
  write_csv("./annual.report/figure6.data.graph/ar_fig6_dataset.csv")

  
# plot ar fig6 dataset
ar_fig6 <- ar_fig6_dataset %>%
    ggplot() +
  geom_bar(mapping = aes(x = fct_reorder(le_gain_grp, le_gain_grp_order), y = tot_population_millions, fill = country_grp), stat = "identity") +
    scale_fill_manual(values = c("chocolate3", "red", "black", "lightgreen", "lightblue", "orange", "grey"), name = "Region") +
    scale_y_continuous(breaks = seq(0, 2500, 500)) +
  labs(x = "Years of Life Gained if Pollution is reduced to WHO Guideline",
       y = "People (in Millions)") +
    theme_classic() 

# ar figure6 version 2
tmp <- ar_fig6_dataset %>%
    ggplot() +
  geom_bar(mapping = aes(x = fct_reorder(le_gain_grp, le_gain_grp_order), y = tot_population_millions, fill = country_grp), stat = "identity", position = position_dodge()) +
    scale_fill_manual(values = c("chocolate3", "red", "black", "lightgreen", "lightblue", "orange", "grey"), name = "Region") +
    scale_y_continuous(breaks = seq(0, 1500, 250)) +
  labs(x = "Years of Life Gained if Pollution is reduced to WHO Guideline",
       y = "People (in Millions)") +
    theme_classic() 


# save annual report figure 6
ggsave("./annual.report/figure6.data.graph/ar_fig6.png")

```


# Figure 7: Year-Over-Year Change in PM2.5 Levels in 2020, the First Year of the Pandemic

```{r ar-fig7}

# Prepare AR Figure 7 data set
ar_fig7_dataset <- color_2020 %>%
  group_by(country) %>%
  mutate(pop_weights = population/sum(population, na.rm =TRUE), 
         pm2019_pop_weighted = pop_weights*pm2019,
         pm2020_pop_weighted = pop_weights*pm2020) %>%
  summarise(avg_pm2.5_pop_weighted_2019 = sum(pm2019_pop_weighted, na.rm = TRUE), 
            avg_pm2.5_pop_weighted_2020 = sum(pm2020_pop_weighted, na.rm = TRUE),
            change_pm2.5_2019_to_2020 = avg_pm2.5_pop_weighted_2020 - avg_pm2.5_pop_weighted_2019,
            percent_change_pm2.5_2019_to_2020 = (change_pm2.5_2019_to_2020/avg_pm2.5_pop_weighted_2019)*100,
            total_population = sum(population, na.rm = TRUE)) %>%
  mutate(direction_label = ifelse(change_pm2.5_2019_to_2020 < 0, "Negative", "Positive")) %>%
  slice_max(total_population, n = 20) 
  
# Save AR figure 7 dataset as a csv
ar_fig7_dataset %>% 
  write_csv("./annual.report/figure7.data.graph/ar_fig7_dataset.csv")

# AR figure 7 plot
ar_fig7 <- ar_fig7_dataset %>%  
  ggplot(mapping = aes(x = fct_reorder(country, percent_change_pm2.5_2019_to_2020), y = percent_change_pm2.5_2019_to_2020, fill = direction_label)) + geom_bar(stat = "identity") +
  geom_hline(mapping = aes(yintercept = 0)) +
  scale_y_continuous(limits = c(-25, 25), breaks = seq(-25, 25, 5)) +
  labs(x = "country", y = "Year over Year change in PM2.5 (%)")+
  ggthemes::theme_clean() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5), legend.position = "none") 


# Save AR figure 7
ggsave("./annual.report/figure7.data.graph/ar_fig7.png", ar_fig7)
```
# Figure 8: Potential Gain in Years of Life Expectancy Through Permanently Reducing PM2.5 from 2020 Concentrations to the WHO Guideline, South Asia

```{r ar-fig8}

# This is a screenshot from the AQLI Index page.
```

# Figure 9: Potential Gain in Years of Life Expectancy Through Permanently Reducing PM2.5 from 2020 Concentrations to the WHO Guideline, in 10 Most Populated Regions in Southeast Asia

```{r ar-fig9}

# South East Asia Definition
se_asia_vec <- c("Brunei", "Myanmar", "Cambodia", "Timor-Leste", "Indonesia", "Laos", "Malaysia", "Philippines", "Singapore", "Thailand", "Vietnam")

# filter dataset for South East Asia and add a new column that combines gadm_1 and gadm_0 columns
ar_fig9_dataset <- color_2020 %>%
  filter(country %in% se_asia_vec) %>%
  mutate(name_1_country = str_c(name_1, "(", country, ")", sep = "")) %>%
  group_by(name_1_country) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2020_pop_weighted = pop_weights*pm2020) %>%
  summarise(tot_pop = sum(population, na.rm = TRUE), avg_pm2.5_pop_weighted = sum(pm2020_pop_weighted, na.rm = TRUE), 
            gain_in_le = (avg_pm2.5_pop_weighted - who_guideline) * le_constant, 
            gain_in_le = ifelse(gain_in_le < 0, 0, gain_in_le)) %>%
  ungroup() %>%
  mutate(name_1_country = str_replace(name_1_country, "Jawa Barat", "West Java"),
         name_1_country = str_replace(name_1_country, "Jawa Timur", "East Java"), 
         name_1_country = str_replace(name_1_country, "Jawa Tengah", "Central Java")
         ) %>%
  slice_max(tot_pop, n = 10) 

# Save AR figure 9 dataset as a csv
ar_fig9_dataset %>% 
  write_csv("./annual.report/figure9.data.graph/ar_fig9_dataset.csv")

# AR figure 9
ar_fig9 <- ar_fig9_dataset %>%
  ggplot(mapping = aes(x = forcats::fct_reorder(name_1_country, gain_in_le), y = gain_in_le, fill = gain_in_le)) +
  geom_col() + 
  labs(y = "Gain In Life Expectancy (Years)", x = "Region") +
  coord_flip() + 
  scale_fill_gradient(low = "darkgoldenrod1", high = "darkorange2") +
  ggthemes::theme_tufte() +
  theme(legend.position = "none")

# save AR figure 9
ggsave("./annual.report/figure9.data.graph/ar_fig9.png", ar_fig9)

```

# Figure 10: Potential Gain in Years of Life Expectancy Through Permanently Reducing PM2.5 from 2020 Concentrations to the WHO Guideline, Southeast Asia

```{r ar-fig10}

# This is a screenshot from the AQLI Index page.

```

# Figure 11: Potential Gain in Life Expectancy from Permanently Reducing PM2.5 from 2020 Concentrations to the WHO Guideline, Central and West Africa

```{r ar-fig11}
# This is a screenshot from the AQLI Index page.
```

# Figure 12: Life Expectancy Impacts of Particulate Pollution and Other Health Threats in the Five Most Populous Countries in Central and West Africa

```{r ar-fig12}

# GBD results filtered for relevant cause of death and countries 
gbd_results_fig12 <- gbd_results %>%
  filter(cause_of_death %in% c("PM2.5 relative to WHO",
                               "Malaria", "Unsafe water, sanitation, and handwashing", 
                               "HIV/AIDS"), location %in% c("nigeria", "democratic republic of the congo", "Angola", "ghana", "cameroon"))

# making the 'location' column of type factor
gbd_results_fig12$location <- factor(gbd_results_fig12$location, 
                                        levels = c("nigeria", "democratic republic of the congo", "Angola", "ghana", "cameroon"))

# Rename Democratic Republic of the Congo to DRC
gbd_results_fig12$location <-  str_replace(gbd_results_fig12$location, 
            "democratic republic of the congo", "DRC")

# Converting 'cause_of_death' to type factor
gbd_results_fig12$cause_of_death <- as.factor(gbd_results_fig12$cause_of_death)

# Rearranging 'cause of death' levels
levels(gbd_results_fig12$cause_of_death) <- c("HIV/AIDS", "Malaria", "PM2.5 relative to WHO", "Unsafe Water, Sanitation and Handwashing")

# wrapping x-axis labels text 
levels(gbd_results_fig12$cause_of_death) <- str_wrap(levels(gbd_results_fig12$cause_of_death), 15)

# getting country wise population
country_wise_population <- color_2020 %>%
  group_by(country) %>%
  summarise(population_total = sum(population, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(country %in% c("Cameroon", "Nigeria", "Angola", "Ghana", 
                  "Democratic Republic of the Congo")) %>%
  arrange(desc(population_total))

# reorder within each location as per the total life years lost column
gbd_results_fig12 <- gbd_results_fig12 %>% 
  mutate(cause_of_death = reorder_within(cause_of_death, est_le_diff_vs_actual_aggregate, location))

# clean the "cause of death" column
gbd_results_fig12 <- gbd_results_fig12 %>%
  mutate(`Cause of Death` = str_remove(cause_of_death, "___.+"))
  

# plot AR figure 12
ar_fig12 <- ggplot(gbd_results_fig12, mapping = aes(x = cause_of_death, y = est_le_diff_vs_actual_aggregate)) + 
  geom_col(mapping = aes(fill = `Cause of Death`), width = 0.7, color = "white") +
  facet_wrap(~factor(location, levels = c("nigeria", "DRC", "Angola", 
                  "ghana", "cameroon")), scales = "free_x", ncol = 5) +
  scale_fill_manual(values = c("darkcyan", "deepskyblue", "red",
                               "aquamarine")) +
  labs(x = "", y = "Years Lost", title = "Life Expectancy Impacts of PM2.5 & Other Health Threats", 
       subtitle = "Five Most Populous Countries in Central and West Africa") +
  ggthemes::theme_fivethirtyeight() +
  theme(axis.text.x = element_blank(), legend.position = "bottom", axis.ticks = element_blank()) 

# save annual report figure 12
ggsave("./annual.report/figure12.data.graph/ar_fig12.png", ar_fig12)


# save AR figure 12 data set
gbd_results_fig12 %>%
  select(location, cause_of_death, est_le_diff_vs_actual_aggregate) %>%
  rename(life_years_lost = est_le_diff_vs_actual_aggregate) %>%
  mutate(life_years_lost = round(life_years_lost, 1)) %>%
  write_csv("./annual.report/figure12.data.graph/ar_fig12_dataset.csv")

```

# Figure 13: Potential Gain in Years of Life Expectancy Through Permanently Reducing PM2.5 from 2020 Concentrations to the WHO Guideline, in 15 Most Populated Regions in Latin America

```{r ar-fig13}

# Latin America definition
latin_america_countries_vec <- c("Mexico", "Guatemala", "Honduras", 
                 "El Salvador", "Nicaragua", 
                 "Costa Rica", "Panama", 
                 "Colombia", "Venezuela", 
                 "Ecuador", "Peru", 
                 "Bolivia", "Brazil", 
                 "Paraguay", "Chile",
                 "Argentina", "Uruguay", 
                 "Cuba", "Haiti", "Dominican Republic", 
                 "Puerto Rico")

# AR Figure 13 dataset
ar_fig13_dataset <- color_2020 %>%
  filter(country %in% latin_america_countries_vec) %>%
  mutate(name_1_country = str_c(name_1, "(", country, ")", sep = "")) 

# Save AR Figure 13 dataset
ar_fig13_dataset %>%
  write_csv("./annual.report/figure13.data.graph/ar_fig13_dataset.csv")


# Plot AR figure 13 dataset
ar_fig13 <- ar_fig13_dataset %>%
  group_by(name_1_country) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         pm2020_pop_weighted = pop_weights*pm2020) %>%
  summarise(tot_pop = sum(population, na.rm = TRUE), avg_pm2.5_pop_weighted = sum(pm2020_pop_weighted, na.rm = TRUE), 
            gain_in_le = (avg_pm2.5_pop_weighted - who_guideline) * le_constant, 
            gain_in_le = ifelse(gain_in_le < 0, 0, gain_in_le)) %>%
  ungroup() %>%
  slice_max(tot_pop, n = 15) %>%
  ggplot(mapping = aes(x = forcats::fct_reorder(name_1_country, tot_pop), y = gain_in_le, fill = gain_in_le)) +
  geom_col() + 
  labs(y = "Gain In Life Expectancy (Years)", x = "Region") +
  coord_flip() + 
  ggthemes::theme_clean() +
  scale_fill_gradient(low = "lightgoldenrod", high = "darkorange2") +
  theme(legend.position = "none", 
        axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12))

# save annual report figure 13
ggsave("./annual.report/figure13.data.graph/ar_fig13.png", ar_fig13)

```


# Figure 14: Potential Gain in Life Expectancy from Permanently Reducing PM2.5 from 2020 Concentrations to the WHO Guideline, Latin America

```{r ar-fig14}

```


# Figure 15: PM2.5 Concentrations in Major Regions in Mainland China Over Time

```{r ar-fig15}

# filter master dataset for China
color_2020_china <- color_2020 %>%
  filter(country %in% "China")

# BTH region (name_1)
bth_region <- c("Beijing", "Tianjin", "Hebei")

# YRD (name_1)
yrd_region <- c("Shanghai", "Jiangsu", "Zhejiang")

# PRD (name_1, name_2)

prd_region_name_2_guandong_regions <- c("Dongguan", "Foshan", "Guangzhou", 
                                        "Huizhou", "Jiangmen", "Shenzhen", 
                                        "Zhaoqing", "Zhongshan", "Zhuhai")

prd_region_name_1 <- c("Hong Kong", "Macao")

# prd combined
prd <- c(prd_region_name_2_guandong_regions, prd_region_name_1)


# add region column in the China dataset
color_2020_china  <- color_2020_china %>%
  mutate(region = ifelse(name_1 %in% bth_region, "BTH", "others"), 
         region = ifelse(name_1 %in% yrd_region,  "YRD", region), 
         region = ifelse(name_1 %in% prd_region_name_1, "PRD", region), 
         region = ifelse(name_2 %in% prd_region_name_2_guandong_regions, "PRD", region))
  

# AR figure 15 (part-1): trend lines figure region wise (BTH, YRD, PRD) dataset
trendlines_china_region_wise_df <- color_2020_china %>%
  group_by(region) %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         mutate(across(starts_with("pm"), ~.x*pop_weights, .names = "{col}_weighted"))) %>%
  summarise(across(ends_with("weighted"), sum)) %>%
  pivot_longer(cols = pm1998_weighted:pm2020_weighted, names_to = "years", 
               values_to = "pop_weighted_avg_pm2.5") %>%
  mutate(years = as.integer(unlist(str_extract(years, "\\d+")))) %>%
  select(years, region, pop_weighted_avg_pm2.5) %>%
  filter(region %in% c("YRD", "BTH", "PRD"))

# AR figure 15 (part-2): trend line national average dataset
trendline_national_avg_df <- color_2020_china %>%
  mutate(pop_weights = population/sum(population, na.rm = TRUE), 
         mutate(across(starts_with("pm"), ~.x*pop_weights, .names = "{col}_weighted"))) %>%
  summarise(across(ends_with("weighted"), sum)) %>%
  pivot_longer(cols = pm1998_weighted:pm2020_weighted, names_to = "years", 
               values_to = "pop_weighted_avg_pm2.5") %>%
  mutate(years = as.integer(unlist(str_extract(years, "\\d+"))), 
         region = "China") %>% 
  select(years, region, pop_weighted_avg_pm2.5)

# AR Figure 15 part 1 and part 2 datasets combined
ar_fig15_dataset <- rbind(trendlines_china_region_wise_df, trendline_national_avg_df)

# Save AR Figure 15 data set as a csv
ar_fig15_dataset %>%  
  write_csv("./annual.report/figure15.data.graph/ar_fig15_dataset.csv")

# plot AR Figure 15
ar_fig15 <- ggplot(ar_fig15_dataset) +
  geom_line(mapping = ggplot2::aes(x = years, y = pop_weighted_avg_pm2.5, color = region), lwd = 1.1) +
    geom_hline(mapping = aes(yintercept = 5), lwd = 0.5, linetype = "dashed") +
    geom_vline(mapping = aes(xintercept = 2014), lwd = 0.5, linetype = "dashed") +
  scale_y_continuous(breaks = seq(0, 100, 10), limits = c(0, 100)) +
  scale_x_continuous(breaks = seq(1998, 2020, 2)) +
    scale_color_manual(values = c("China" = "darkgrey", "BTH" = "darkred", "YRD" = "darkorange", "PRD" = "burlywood1")) +
  ggthemes::theme_clean() +
  labs(x = "Years", 
       y = expression(paste("Average PM2.5 concentration ( ", mu, "g", "/", m^3, " )"))) +
  theme(legend.position = "bottom", legend.title = element_blank(), 
        legend.text = element_text(size = 6.5), 
        axis.title.y = element_text(size = 9), 
        axis.title.x = element_text(size = 9)) + 
    geom_text(x = 2003.8, y = 10, label = expression(paste("WHO PM2.5 Guideline (last updated: 2021): 5 ", mu, "g","/", m^3, ""))) + 
    geom_text(x = 2016.5, y = 93, label = str_wrap("China announces war on pollution", width = 18))
  

# Save AR figure 15
ggsave("./annual.report/figure15.data.graph/ar_fig15.png", ar_fig15)
```


# Figure 16: Potential Gain in Life Expectancy from Reducing PM2.5 to the WHO Guideline in Mainland China, 2013 vs. 2020

```{r ar-fig16}

```


# Figure 17: Change in Life Expectancy Due to Change in PM2.5 in United States, 1970-2020

```{r ar-fig17}

```

# Figure 18: Change in Life Expectancy Due to Change in PM2.5 Concentrations in Europe, 1998-2020

```{r ar-fig18}

```

# Figure 19: Potential Gain in Life Expectancy From Permanently Reducing PM2.5 From 2020 Concentrations to the WHO Guideline, Europe

```{r ar-fig19}

```


# Figure 20: Global Population Weighted PM2.5 concentrations over time

```{r ar-fig20}

# add population weighted pm column in the 2020 color dataset
color_2020_fig20 <- color_2020 %>%
  mutate(global_pop_weights = population/sum(population)) %>%
  mutate(across(starts_with("pm"), ~.x*global_pop_weights, .names = "{col}_weighted"))

# add population weighted pm column in the 2019 color dataset
color_2019_fig20 <- color_2019 %>%
  mutate(global_pop_weights = population/sum(population)) %>%
  mutate(across(starts_with("pm"), ~.x*global_pop_weights, .names = "{col}_weighted"))

# add population weighted pm column in the 2016 color dataset
color_2016_fig20 <- color_2016 %>%
  mutate(global_pop_weights = population/sum(population)) %>%
  mutate(across(starts_with("pm"), ~.x*global_pop_weights, .names = "{col}_weighted"))

# calculate average PM2.5 from population weighted column in the 2019 color dataset
color_2019_summary <- color_2019_fig20 %>%
  summarise(across(ends_with("weighted"), sum))

# calculate average PM2.5 from population weighted column in the 2020 color dataset
color_2020_summary <- color_2020_fig20 %>%
  summarise(across(ends_with("weighted"), sum))

# calculate average PM2.5 from population weighted column in the 2016 color dataset
color_2016_summary <- color_2016_fig20 %>%
  summarise(across(ends_with("weighted"), sum))

# Creating a list of datasets storing global trendlines data for each of color_2020, color_2019 and color_2016 datasets
global_trends_figure_list <- list(tibble(years = 1998:2019, global_avg_pm2.5 = as.numeric(unlist(color_2019_summary))), tibble(years = 1998:2020, global_avg_pm2.5 = as.numeric(unlist(color_2020_summary))), tibble(years = 2016:1998, global_avg_pm2.5 = as.numeric(unlist(color_2016_summary))))

# creating a final wide dataset from the individual datasets stored in the above list
final_dataset <- left_join(left_join(global_trends_figure_list[[2]], global_trends_figure_list[[1]], by = "years"), global_trends_figure_list[[3]], by = "years")

# renaming columns in the final wide dataset
final_dataset <- final_dataset %>%
  rename(`2020 dataset` = global_avg_pm2.5.x , `2019 dataset` = global_avg_pm2.5.y, 
         `2016 dataset` = global_avg_pm2.5)

# converting dataset into a long format so that we can easily plot it using ggplot
final_dataset <- tidyr::pivot_longer(final_dataset, `2020 dataset`:`2016 dataset`,  names_to = "Reference Dataset", values_to = "PM2.5")

# adding a le_gains column
ar_fig20_dataset <- final_dataset %>%
  mutate(le_gains = (PM2.5 - 5)*0.098, 
         le_gains = ifelse(le_gains < 0, 0, le_gains))

# Save AR Figure 20 data set as a csv
ar_fig20_dataset %>%  
  write_csv("./annual.report/figure15.data.graph/ar_fig20_dataset.csv")

# plotting Annual Report Figure 20
ar_fig20 <- ggplot(ar_fig20_dataset) +
  geom_line(mapping = aes(x = years, y = PM2.5, group = `Reference Dataset`, color = `Reference Dataset`), size = 1.75) +
  ggthemes::theme_stata() +
  labs(x = "Years", y = expression(paste("Average PM2.5 concentration ( ", mu, "g", "/", m^3, " )"))) + 
  scale_x_continuous(breaks = seq(1998, 2020, 2)) +
  scale_y_continuous(breaks = seq(0, 40, 10), limits = c(0, 40)) +
  scale_color_manual(values = c("lightgrey", "darkgrey", "darkred")) +
  theme(legend.position = "bottom", legend.title = element_blank(), axis.title.x =  element_text(size = 8), axis.title.y = element_text(size = 10)) +
  ggtitle("Global Population Weighted PM2.5 Concentration Over Time")


# Save AR figure 20
ggsave("./annual.report/figure20.data.graph/ar_fig20.png", ar_fig20)

```

